### 2019-08-26 paper.Rmd
Adding material from other repo.

---
title: |
  A slice tour
type: Short Technical Note
author:
  - name: Ursula Laa
    affil: a, b
    email: ursula.laa@monash.edu
  - name: Dianne Cook
    affil: b
    email: dicook@monash.edu
affiliation:
  - num: a
    address: |
      School of Physics and Astronomy, Monash University
  - num: b
    address: |
      Department of Econometrics and Business Statistics, Monash University
bibliography: biblio.bib
abstract: |
  Projections and sectioning are complementary methods for the visualisation of
  high-dimensional data. In this note we describe slicing in the orthogonal
  space of projections obtained when running a tour, thus presenting the viewer
  with an interpolated sequence of projected sections. The method has been
  implemented in R as an extension to the tourr package, and can be used to
  explore concave structures in multivariate distributions.
keywords: |
  data visualisation; grand tour; sectioning
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
output: rticles::tf_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction

A good way to visualize high-dimensional data is to display a smooth sequence of interpolated lower dimensional projections. This is called a tour [@As85; @BCAH05] and allows to extrapolate the high-dimensional shape of the distribution. It is particularly useful for detecting clusters, outliers and non-linear dependence. A limitation of this approach is that concavities in a distribution may be hidden when looking at projections. A simple example are projections of samples on a hypersphere. No projection of such a distribution allows to distinguish a hollow sphere from a full one. On the other hand, the difference would be obvious when looking at a slice through the center of the sphere.

Slicing or sectioning can be thought of as complementary to projections when visualizing high-dimensional data, as discussed in detail in @prosection. To overcome the respective shortcomings of each method, it is suggested to use a composition of sections and projections, so-called "prosections". In practice sectioning is typically performed in interactive graphics displays in the form of brushing [XXX reference?], i.e. the user is selecting a region in a subspace of the high-dimensional dataspace that is highlighted in a projection of the data points.
More recently @JSSv081i05 suggested conditional visualization for statistical models by selecting sections in data space for detailed comparisons of a fitted model to the observed data. Similar to brushing, the section is selected interactively by the user, by choosing a point in low-dimensional displays showing the data in the space of conditioning predictors.

Here we propose a systematic approach to slicing that does not rely on the user choosing sections to display, but will present the viewer with an interpolated sequence showing different slices of parameter space. The method is based on the grand tour, and uses the orthogonal distance from the current plane to define the section corresponding to each view in an interpolated tour path. We demonstrate how this can be used to gain insights into the concave structure of a dataset, using the example of hollow hyperspheres defined in 3 or more dimensions.

# Method

The basis underlying our approach is the grand tour. We first briefly review the relevant concepts. Building on these ideas we introduce a method for displaying slices of parameter space that are defined for each projection matrix.

## Tour review

A tour is providing a continuous animation, showing a sequence of $d$-dimensional (typically $d=1$ or $2$) projections from $n$-dimensional parameter space. The tour combines basis selection with basis interpolation. In a grand tour the basis selection is random (each new basis is chosen from all possible projections), and typically geodesic interpolation is used to generate smooth transitions between viewing planes. Different basis selection methods as well as the geodesic interpolation are implemented in the `tourr` package [@tourr] in R [@rref], which also provides several display functions for viewing the tour.

The notation used in the following for the description of each projection is denoting $\mathbf{X}$ the $m\times n$ dimensional data matrix ($m$ observations in $n$ dimensions) and $\mathbf{A}$ an orthonormal $n \times d$ projection matrix. The $d$-dimensional projection of the data is thus given by $\mathbf{Y} = \mathbf{X}\cdot\mathbf{A}$, giving the $n\times d$ dimensional projected data matrix to be plotted in each frame of the tour display.

In the following we are interested in the orthogonal distance of scatter points from an anchored projection plane, thus $d=2$.

## Slicing in the orthogonal space

In addition to the projected data $\mathbf{Y}$ we calculate for each projection the orthogonal distance of the data points from the current projection plane centered at the origin, $v$. The orthogonal component is given by
\begin{equation}
x_i' = x_i - (x_i\cdot a_1) a_1 - (x_i\cdot a_2) a_2
\end{equation}
with $x_i, i=1,...,m$ an $n$-dimensional observation in $\mathbf{X}$ and $a_i, i=1,2$ denoting the columns of the projection matrix, $\mathbf{A}=(a_1, a_2)$.
Calculating the euclidean norm of $x'$ is straightforward thanks to the orthonormality of $\mathbf{A}$, and gives the components of the required distance vector.
\begin{equation}
\tilde{v}_i^2 = ||x_i'||^2 = x_i^2 - (x_i\cdot a_1)^2 - (x_i\cdot a_2)^2.
\end{equation}
We also see that this is trivially generalizing for orthonormal projection matrices with $d>2$.

The distance vector can then be used to display a slice tour by highlighting points for which the orthogonal distance is smaller than a selected cutoff value $h$. In general a larger cutoff value and/or much larger sample sizes are required as the original dimensionality ($n$) increases. For practical purposes we therefore specify a volume $\epsilon$ and then use as a cutoff $h = \epsilon^{1/n}$.

### Anchoring

The above definition implicitly assumes that we slice the distribution using the origin as a reference point. This is generally a good starting point assuming the data is centered before applying the slice tour. Alternative reference points can however provide additional insight into the concave structure of the distribution. We can also use the mean in each parameter as an anchor point if the data is not centered.

We therefore generalize the distance vector to measure the anchored orthogonal distance, i.e. distance with respect to the anchor $\alpha$ used as a reference point.

In general $\alpha$ can be any point in the $n$-dimensional parameter space, but we are only interested in the orthogonal component, thus we use $\alpha' = \alpha - (\alpha\cdot a_1) a_1 - (\alpha\cdot a_2 )a_2$ to define the generalized measure of orthogonal distance as
\begin{equation}
v_i^2 = ||x_i' - \alpha'||^2 = x_i'^2 + \alpha'^2 - 2 x_i'\cdot\alpha'
\end{equation}

where the cross term can be expressed as

\begin{equation}
x_i'\cdot\alpha' = x_i\cdot\alpha - (\alpha\cdot a_1) (x_i\cdot a_1) - (\alpha\cdot a_2) (x_i\cdot a_2).
\end{equation}

Using the generalized distance measure with a cutoff volume $\epsilon$ then corresponds to moving a slice of fixed thickness to be centered around the anchor point $\alpha$ in the space orthogonal to the current projection plane. In three dimensions this simply corresponds to moving up or down along the orthogonal direction.

# Implementation

The slice tour has been implemented in R [@rref] as a new display method `display_slice` in the `tourr` package [@tourr]. In addition to usual parameters the user can choose the volume parameter $\epsilon$ by setting `eps` and, if required, the anchor point $\alpha$ is set by the `anchor` argument. By default `eps=0.001` and `anchor=NULL`, resulting in slicing through the origin. In addition the user can select the marker symbols for point inside and outside the slice. By default points in the slice are highlighted as `pch_slice  = 20` and `pch_other = 46`, i.e. plotting a bullet for points inside the slice, and a dot for points outside.

# Example

To demonstrate the slice tour we sample points on the surface of a sphere with radius $r=1$ in $n=3$ and $n=5$ dimensions. We use the `geozoo` package [@geozoo] to draw 2000 (5000) points on the 3d (5d) sphere.

For both examples we start by slicing through the origin with the default parameters, in particular $\epsilon = 0.001$, i.e. $h = 0.1$ in 3d and $h = 0.25$ in 5d. Slicing through the origin results in sections where the points inside the slice are approximately on a circle with the full radius. Snapshots of the slice tour animation are shown in Figure \ref{fig:centered}, animated gifs are available at XXX. Notice that the larger number of points and increased thickness parameter $h$ do not make up for the reduced number of points inside the slice through the 5d hypersphere. (projected points are on average closer to the center as we increase the dimension of the hypersphere, is there some convergence similar to central limit theorem at play?)

\begin{figure*}[ht]
\centerline{\includegraphics[width=0.3\textwidth]{gifs/png/sphere-3-centered-055.png}\hspace{1cm}
\includegraphics[width=0.3\textwidth]{gifs/png/sphere-5-centered-055.png}}
\caption{Slice through a 3d (left) and 5d (right) hollow sphere, anchored at the origin. Points inside the slice are highlighted with a bullet marker and are found on a circle with the radius of the hypersphere, $r$.}
\label{fig:centered}
\end{figure*}

It is also instructive to look at the sliced spheres with an off-center anchor point. In this case the different views obtained in the slice tour reveal more of the concave structure of the distribution. Figure \ref{fig:3anchored} shows example views from a slice tour on a 3d sphere with $r=1$ and anchor point $(0.7, 0.7, 0.7)$. Notice that the anchor point was chosen to fall outside the sphere. Depending on the viewing angle selected by the tour the slice contains points on the circle with radius $r$ (when the anchor point has a negligible orthogonal component to the viewing plane); a circle with radius $<r$ as the viewing angle is tilted away from the axis connecting the anchor point to the origin; a full circle with small radius as the angle increases; and finally we see an empty slice when the projection plane is orthogonal to the anchor point axis.

\begin{figure*}[ht]
\centerline{\includegraphics[width=0.3\textwidth]{gifs/png/sphere-3-anchored-033.png}
\includegraphics[width=0.3\textwidth]{gifs/png/sphere-3-anchored-036.png}
\includegraphics[width=0.3\textwidth]{gifs/png/sphere-3-anchored-041.png}}
\caption{Different slices through a 3d hollow sphere with $r=1$, with anchor point $(0.7, 0.7, 0.7)$, showing full circle with small radius (left), circle with radius $<r$ (middle) and circle with radius $r$ (right).}
\label{fig:3anchored}
\end{figure*}

The same behavior is observed when slicing the 5d hollow sphere with an off-center anchor point, however a large number of sample points is required to get a similarly clear display to that found for the 3d sphere. The animation showing the anchored slice tour of the 5d hollow sphere with 5000 sample points is available at XXX. In this example the anchor point was fixed to $(0.5,0.5,0.5,0.5,0.5)$.

# Summary and Outlook

In this note we introduce a new visualization method for systematic slicing of high-dimensional parameter spaces, based on interpolated projections obtained in a grand tour. The application to the hollow sphere demonstrates the potential of this new display to find and explore concave structures in distributions.

In applications with large number of dimensions and where only a few samples are available the method is limited. In this case thick slices may be needed to get a sufficient number of observations inside the slice, and as a result  small scale concavities cannot be resolved in the slice tour. If on the other hand a large number of samples are available one may consider displaying only points inside the slice to keep a clear display.

Instead of slicing in the full $n-d=n-2$ dimensional orthogonal space we can also use a subspace in which to slice. As pointed out in @prosection combining projections with sectioning in varying numbers of dimensions can reveal the codimension of an unknown distribution. One approach would be to interpolate between projection matrices with more than two columns, and use the first two for the display projection. The remaining directions then provide an orthogonal subspace that can be used for sectioning. A subtlety to consider is that geodesic interpolation is not ideal for such applications, as it ignores rotations between the directions in the projection matrix. As alternative interpolation methods are not currently available in the R implementation, we leave such extensions for future work.

We also expect interesting new developments by combining the slice tour and the guided tour into a weighted guided tour with index functions using the distance vector information in addition to the projected data points.

### 2019-08-26 gifs/mkgif.py

@@ -1,0 +1,20 @@
+import os
+
+
+p = "./png/"
+allNames = os.listdir(p)
+allNamesUnique = []
+for n in allNames:                                
+  n = n[0:-8]
+  if not n in allNamesUnique: allNamesUnique.append(n)
+
+import moviepy.editor as mpy
+
+for cn in allNamesUnique:                         
+  cNames = []        
+  for nn in allNames:                                 
+    if cn in nn: cNames.append(nn)
+  cNames = sorted(cNames)
+  cPaths = [p+cName for cName in cNames]
+  clip = mpy.ImageSequenceClip(cPaths, fps=6)
+  clip.write_gif(cn+".gif", fps=6)

### 2019-08-26 render-examples.R

@@ -1,0 +1,80 @@
+library(tourr)
+library(tidyverse)
+
+# Generate samples on a 3d and 5d hollow sphere using the geozoo package
+sphere3 <- geozoo::sphere.hollow(3, 2000)$points
+sphere5 <- geozoo::sphere.hollow(5, 5000)$points
+
+# Columns need to be named before launching the tour
+colnames(sphere3) <- c("x1", "x2", "x3")
+colnames(sphere5) <- c("x1", "x2", "x3", "x4", "x5")
+
+# Animate with the slice display using the default parameters
+max_f <- 100 # max frames for rendering the animation
+set.seed(2019)
+render(sphere3, grand_tour(), display_slice(), 'png', "gifs/png/sphere-3-centered-%03d.png", frames = max_f, rescale = FALSE)
+
+set.seed(2019)
+render(sphere5, grand_tour(), display_slice(), 'png', "gifs/png/sphere-5-centered-%03d.png", frames = max_f, rescale = FALSE)
+
+
+# Animate with off-center anchoring
+anchor3 <- rep(0.7, 3)
+anchor5 <- rep(0.5, 5)
+
+set.seed(2019)
+render(sphere3, grand_tour(), display_slice(anchor = anchor3), 'png', "gifs/png/sphere-3-anchored-%03d.png", frames = max_f, rescale = FALSE)
+
+# Animate with thicker slice to capture more points in each view
+set.seed(2019)
+render(sphere5, grand_tour(), display_slice(anchor = anchor5), 'png', "gifs/png/sphere-5-anchored-%03d.png", frames = max_f, rescale = FALSE)
+
+
+
+### playing with other exampels
+
+ex1 <- geozoo::boy.surface()$points
+colnames(ex1) <- c("x1", "x2", "x3")
+animate_slice(ex1)
+
+ex2 <- geozoo::cross.cap()$points
+colnames(ex2) <- c("x1", "x2", "x3")
+animate_slice(ex2)
+
+# nice example, only sometimes do we see "corners and edges"
+ex3 <- geozoo::cube.face(6)$points %>%
+  center()
+colnames(ex3) <- c("x1", "x2", "x3", "x4", "x5", "x6")
+animate_slice(ex3)
+
+ex4 <- geozoo::dini.surface()$points %>%
+  scale()
[…]

### 2019-08-26 paper.Rmd
additions to short technical note'

@@
 title: |
-  A slice tour
+  A slice tour for finding hollowness in high-dimensional data
 type: Short Technical Note
@@
 
+### Distance from the origin
+
 In addition to the projected data $\mathbf{Y}$ we calculate for each projection the orthogonal distance of the data points from the current projection plane centered at the origin, $v$.
@@
 For practical purposes we therefore specify a volume $\epsilon$ and then use as a cutoff $h = \epsilon^{1/n}$.
+
+XXX Example 2x2, row 1-col 1 has a slice projection  of a solid 4D spehere, row 1-col 2 has a slice projectionof a hollow 4D sphere; row 2-col 1 has a slice projection of a  solid 4D cube, row 1-col 2 has a slice projection of a hollow 4D cube
+
+### Slice thickness
 
@@
 
-# Example
+XXX R code
+
+# Examples
+
+XXX choose 6
 

### 2019-08-26 render-examples.R
working on more examples

@@ -78,3 +78,20 @@ ex9 <- geozoo::torus.flat()$points
 colnames(ex9) <- c("x1", "x2", "x3", "x4")
 anchorT <- c(0.5, 0.5, 0.5, 0.5)
 animate_slice(ex9, anchor=anchorT)
+
+# Classical pollen data: this doesn't work, structure is 2D in 5D so mostly missed
+library(animation)
+data(pollen)
+pollen <- as.matrix(pollen)
+animate_slice(pollen, axes = "bottomleft", eps=0.01, rescale=TRUE)
+
+# Could we look at nonlinear boundaries between classifiers
+library(RColorBrewer)
+wine_radial <- read_csv("data/wine-svm-radial.csv")
+wine_poly <- read_csv("data/wine-svm-poly.csv")
+clrs <- brewer.pal(3, "Dark2")
+col <- clrs[as.numeric(as.factor(wine_poly$type))]
+animate_slice(as.matrix(wine_poly[,1:5]), axes = "bottomleft", 
+              eps=0.001, col=col)
+
+

### 2019-08-27 paper.Rmd
Trying to work out thickness required.

@@
 
+The thickness of the slice should be chosen such that enough points are inside the slice be representative of the distribution.
+This will depend on the total number of points in the dataset, its original dimensionality and on the distribution.
+Here we estimate the number of points inside a slice of thickness $h$ for the rotation invariant example of an $n$ dimensional full hypersphere with radius $R$, assuming uniformly distributed points throughout the sphere.
+This means we calculate the relative volume inside the slice as a function of $h$ and $n$.
+
+The total volume of the full hypersphere is given as
+\begin{equation}
+V_n(R) = \frac{\pi^{n/2}}{\Gamma(\frac{n}{2} + 1)} R^n.
+\end{equation}
+To calculate the volume of the slice we subtract the volume of a hypersphere cap of hight $h'=r-h$, as given in @capVol,
+\begin{equation}
+V^{\phi}_n(R) = \frac{1}{2}V_n(R) I_{sin^2\phi}(\frac{n+1}{2},\frac{1}{2}),
+\end{equation}
+where $I$ is the regularized incomplete beta function and $\phi =\arccos(1-\frac{h'}{r}) = \arccos(\frac{h}{r})$ the colatitude angle (XXX got this off the paper, don't yet understand the angle..).
+The relative volume of the slice is thus given as (XXX better names needed!)
+\begin{equation}
+V_n^h(R) = \frac{V_n(R) - 2 V^{\phi}_n(R)}{V_n(R)} = 1 - I_{1-\frac{h^2}{r^2}}(\frac{n+1}{2},\frac{1}{2}),
+\end{equation}
+using $\sin(\arccos(x)) = \sqrt{1-x^2}$.
+
+In the case of a hollow hypersphere we are instead interested in the relative area inside the slice. We use the cap area derived in @capVol to calculate the area inside the slice as
+\begin{equation}
+A_n^h(R) = \frac{A_n(R) - 2 A^{\phi}_n(R)}{A_n(R)} = 1 - I_{1-\frac{h^2}{r^2}}(\frac{n-1}{2},\frac{1}{2}).
+\end{equation}
+
+XXXX not sure this makes sense, need to check against plots (not what I was expecting, need to think about this more)
+
 ### Anchoring

### 2019-08-27 render-examples.R

@@ -77,4 +77,10 @@ animate_slice(ex8)
 ex9 <- geozoo::torus.flat()$points
 colnames(ex9) <- c("x1", "x2", "x3", "x4")
 anchorT <- c(0.5, 0.5, 0.5, 0.5)
-animate_slice(ex9, anchor=anchorT)
+animate_slice(ex9, eps=0.01) #, anchor=anchorT)
+
+
+ex9 <- geozoo::torus(p=4)$points %>%
+  scale()
+colnames(ex9) <- c("x1", "x2", "x3", "x4")
+animate_slice(ex9, eps=0.01)

### 2019-08-27 paper.Rmd
Starting work on introduction.

@@
 
+Modern data is typically high-dimensional and thus data visualisation often relies on some form of dimension reduction.
+Interactive and dynamic displays provide information beyond what can be achieved in a static display.
+For example, the grand tour [@As85; @BCAH05] shows a smooth sequence of interpolated lower dimensional projections and allows to extrapolate the high-dimensional shape of the distribution.
+It is particularly useful for detecting clusters, outliers and non-linear dependence.
+Additionaly we may complement projections with section information.
+This is often possible in interactive graphics systems by means of brushing [@ggobiCSDA2003;@JSSv081i05],
+(XXX check what references to put)
+where the user is selecting a region in a subspace of the dataspace that is highlighted in a projection of the data points.
+The complementarity between projections and sectionings has previously been discussed in @prosection, dubbing the combination "prosections".
+These prosections were found to allow identification of lower-dimensional structure embedded in the high-dimensional space.
+One simple example is that prosections can reveal the difference between a full and hollow hypersphere, that is it reveals the intrinsically lower dimension of the hyper-surface embedded in the higher-dimensional space.
+
+However, all existing approaches to sectioning are interactive and rely on the user to redefine the section many times to explore the possible sections.
+While this is often preferred, it is impractical for detecting and exploring unknown concave structures in a dataset.
+We therefore propose a more systematic approach to sectioning based on interpolated projections obtained with a grand tour.
+The method defines a slice based on 2-d projections provided by the tour and thus presents the user with an interpolated sequence of slices through the data space.
+
+We first define the new slice tour in Section XXX, and outline the implementation in R in Section XXX.
+We then apply the method to examples from the geozoo package, showing how hollowness can be visualised and explored with a slice tour.
+Possible extensions are discussed in Section XXX.
+
+<!--
 A good way to visualize high-dimensional data is to display a smooth sequence of interpolated lower dimensional projections.
@@
 We demonstrate how this can be used to gain insights into the concave structure of a dataset, using the example of hollow hyperspheres defined in 3 or more dimensions.
+-->
 
@@
 XXXX not sure this makes sense, need to check against plots (not what I was expecting, need to think about this more)
+Best guess is that relation between $\phi$ and $h$ only holds for 3d sphere, how can I generalise it?
 

### 2019-08-28 relative_vol_area.R
Checks on relative volume.

@@ -18,9 +18,15 @@ relArea <- function(n, r){
   }
 }
 
+pbetaX <- function(n, r){
+  function(x){
+    pbeta(1-x^2, n/2, 1/2)
+  }
+}
+
 r <- 1
 
-ggplot(data.frame(x=c(0.01, 0.2)), aes(x=x)) +
+ggplot(data.frame(x=c(0.01, 0.5)), aes(x=x)) +
   stat_function(fun=relVol(3, r), geom="line", color="black") +
   stat_function(fun=relVol(6, r), geom="line", color="blue") +
   stat_function(fun=relVol(10, r), geom="line", color="green") +
@@ -30,4 +36,15 @@ ggplot(data.frame(x=c(0.01, 0.2)), aes(x=x)) +
   stat_function(fun=relArea(10, r), geom="line", color="green", linetype="dashed") +
   stat_function(fun=relArea(20, r), geom="line", color="red", linetype="dashed") +
   xlab("h") +
-  ylab("relative volume")
+  ylab("relative volume/area")
+
+ggplot(data.frame(x=c(0, 0.2)), aes(x=x)) +
+  stat_function(fun=pbetaX(3, r), geom="line", color="black") +
+  stat_function(fun=pbetaX(6, r), geom="line", color="blue") +
+  stat_function(fun=pbetaX(10, r), geom="line", color="green") +
+  stat_function(fun=pbetaX(20, r), geom="line", color="red") +
+  xlab("h") +
+  ylab("pbeta")
+
+
+

### 2019-08-28 render-examples.R
data added

@@ -83,7 +83,7 @@ animate_slice(ex9, anchor=anchorT)
 library(animation)
 data(pollen)
 pollen <- as.matrix(pollen)
-animate_slice(pollen, axes = "bottomleft", eps=0.01, rescale=TRUE)
+animate_slice(pollen, axes = "bottomleft", eps=0.0001, rescale=TRUE)
 
 # Could we look at nonlinear boundaries between classifiers
 library(RColorBrewer)
@@ -93,5 +93,5 @@ clrs <- brewer.pal(3, "Dark2")
 col <- clrs[as.numeric(as.factor(wine_poly$type))]
 animate_slice(as.matrix(wine_poly[,1:5]), axes = "bottomleft", 
               eps=0.001, col=col)
-
+animate_xy(as.matrix(wine_poly[,1:5]), axes = "bottomleft", col=col)
 

### 2019-08-28 paper.Rmd
Rewriting of the introduction.

--- 2019-08-27 paper.Rmd
@@
 
+Modern data is typically high-dimensional and thus data visualisation often relies on some form of dimension reduction.
+This entails looking at (non-)linear projections of the data.
+Interactive and dynamic displays often provide information beyond what can be achieved in a static display.
+For example, the grand tour [@As85; @BCAH05] shows a smooth sequence of interpolated lower dimensional projections and allows to extrapolate the high-dimensional shape of the distribution.
+It is particularly useful for detecting clusters, outliers and non-linear dependence.
+Additionaly, we may complement projections with section information.
+This is often possible in interactive graphics systems by means of brushing [@ggobiCSDA2003;@JSSv081i05], where the user is selecting a region in a subspace of the dataspace that is highlighted in a projection of the data points.
+
+The complementarity between projections and sections has previously been discussed in @prosection, dubbing the combination "prosections".
+These prosections allow the identification of lower-dimensional structure embedded in the high-dimensional space.
+One simple example is that prosections can reveal the difference between a full and hollow hypersphere.
+In addition to hollowness (concave structures) they can also provide a clear picture of internal structures that may be concealed in a projection.
+For example, @prosection uses prosections to reveal the internal structure of the four-point ultrametric locus [@locus].
+
+Since existing approaches to sectioning are interactive, they rely on the user to redefine the section many times to explore the data.
+While this level of control is often preferred, it is impractical for detecting and exploring unknown structures in a dataset.
+As a result smaller hidden features that are only visible in specific sections of the full distribution may easily be missed.
+In addition having a dynamic approach to sectioning, similar to the grand tour approach to projections, will provide the viewer with context allowing to extrapolate e.g.
+the shape of a concave structure.
+We therefore propose a more systematic approach to sectioning based on interpolated projections obtained with a grand tour.
+The method defines a slice based on 2-d projections provided by the tour and will presents the viewer with an interpolated sequence of slices when running a tour.
+
+The relevant definitions and an overview of the implementation in R are given in Section XXX, and in Section XXX.
+We then apply the newly defined slice tour to several examples in Section XXX.
+We consider different high-dimensional geometric shapes and XXX, showing how concave structures can be visualised and explored with a slice tour.
+Possible extensions are discussed in Section XXX.
+
+<!--
 A good way to visualize high-dimensional data is to display a smooth sequence of interpolated lower dimensional projections.

### 2019-08-28 relative_vol_area.R

@@ -28,13 +28,27 @@ r <- 1
 
 ggplot(data.frame(x=c(0.01, 0.5)), aes(x=x)) +
   stat_function(fun=relVol(3, r), geom="line", color="black") +
+  stat_function(fun=relVol(4, r), geom="line", color="purple") +
   stat_function(fun=relVol(6, r), geom="line", color="blue") +
   stat_function(fun=relVol(10, r), geom="line", color="green") +
   stat_function(fun=relVol(20, r), geom="line", color="red") +
   stat_function(fun=relArea(3, r), geom="line", color="black", linetype="dashed") +
+  stat_function(fun=relArea(4, r), geom="line", color="purple", linetype="dashed") +
   stat_function(fun=relArea(6, r), geom="line", color="blue", linetype="dashed") +
   stat_function(fun=relArea(10, r), geom="line", color="green", linetype="dashed") +
   stat_function(fun=relArea(20, r), geom="line", color="red", linetype="dashed") +
+  xlab("h") +
+  ylab("relative volume/area")
+
+ggplot(data.frame(x=c(0.01, 0.2)), aes(x=x)) +
+  stat_function(fun=relVol(3, r), geom="line", color="black") +
+  stat_function(fun=relVol(4, r), geom="line", color="blue") +
+  stat_function(fun=relVol(5, r), geom="line", color="green") +
+  stat_function(fun=relVol(6, r), geom="line", color="red") +
+  stat_function(fun=relArea(3, r), geom="line", color="black", linetype="dashed") +
+  stat_function(fun=relArea(4, r), geom="line", color="blue", linetype="dashed") +
+  stat_function(fun=relArea(5, r), geom="line", color="green", linetype="dashed") +
+  stat_function(fun=relArea(6, r), geom="line", color="red", linetype="dashed") +
   xlab("h") +
   ylab("relative volume/area")
 

### 2019-08-28 relative_vol_area.R
Adding h as function of relative volume

@@ -60,5 +60,20 @@ ggplot(data.frame(x=c(0, 0.2)), aes(x=x)) +
   xlab("h") +
   ylab("pbeta")
 
+library(zipfR)
+hPerRelVol <- function(n, r){
+    function(vrel){
+      x <- 1 - vrel
+      a <- (n+1) / 2
+      b <- 1/2
+      1 - Rbeta.inv(x, a, b)
+    }
+}
 
-
+ggplot(data.frame(x=c(0.01, 0.2)), aes(x=x)) +
+  stat_function(fun=hPerRelVol(3, r), geom="line", color="black") +
+  stat_function(fun=hPerRelVol(4, r), geom="line", color="blue") +
+  stat_function(fun=hPerRelVol(5, r), geom="line", color="green") +
+  stat_function(fun=hPerRelVol(6, r), geom="line", color="red") +
+  xlab("relative volume") +
+  ylab("h")

### 2019-08-28 render-examples.R

@@ -96,7 +96,10 @@ wine_radial <- read_csv("data/wine-svm-radial.csv")
 wine_poly <- read_csv("data/wine-svm-poly.csv")
 clrs <- brewer.pal(3, "Dark2")
 col <- clrs[as.numeric(as.factor(wine_poly$type))]
+wine_poly_scaled <- scale(as.matrix(wine_poly[,1:5]))
 animate_slice(as.matrix(wine_poly[,1:5]), axes = "bottomleft", 
               eps=0.001, col=col)
+animate_slice(wine_poly_scaled, axes = "bottomleft", 
+              eps=0.05, col=col)
 animate_xy(as.matrix(wine_poly[,1:5]), axes = "bottomleft", col=col)
 

### 2019-08-28 relative_vol_area.R
Small fix

@@ -66,7 +66,7 @@ hPerRelVol <- function(n, r){
       x <- 1 - vrel
       a <- (n+1) / 2
       b <- 1/2
-      1 - Rbeta.inv(x, a, b)
+      r * sqrt(1 - Rbeta.inv(x, a, b))
     }
 }
 

### 2019-08-29 paper.Rmd
Added simulation, trying to work out whats wrong.

@@
 
+XXXX maybe the problem is that each plane is 2d, but the cap assumes I am cutting with a hyperplane that is n-1, so it works as expected in 3d, beyond that i need to account for volume effects in the n-2 dimensions, but how?
+e.g.
+consider 1-d sliced of 3d sphere, in that case I get the volume of the slice as a pipe of radius h, so for n-d sphere I need the volume of n-1 d sphere of radius h but translated across the plane (but what does that really mean)?
+
 The thickness of the slice should be chosen such that enough points are inside the slice be representative of the distribution.
@@
 \end{equation}
-
-XXXX not sure this makes sense, need to check against plots (not what I was expecting, need to think about this more)
-Best guess is that relation between $\phi$ and $h$ only holds for 3d sphere, how can I generalise it?
 

### 2019-08-29 relative_vol_area.R

@@ -9,6 +9,15 @@ relVol <- function(n, r){
   }
 }
 
+#WRONG
+relVol <- function(n, r){
+  Vectorize(
+  function(h){
+    (h/r)*(1/sqrt(pi))*(gamma((n/2)+1) / gamma(((n-1)/2) + 1))
+  }
+  )
+}
+
 relArea <- function(n, r){
   function(h){
     x <- 1 - (h/r)^2
@@ -40,7 +49,7 @@ ggplot(data.frame(x=c(0.01, 0.5)), aes(x=x)) +
   xlab("h") +
   ylab("relative volume/area")
 
-ggplot(data.frame(x=c(0.01, 0.2)), aes(x=x)) +
+ggplot(data.frame(x=c(0.01, 0.5)), aes(x=x)) +
   stat_function(fun=relVol(3, r), geom="line", color="black") +
   stat_function(fun=relVol(4, r), geom="line", color="blue") +
   stat_function(fun=relVol(5, r), geom="line", color="green") +
@@ -77,3 +86,42 @@ ggplot(data.frame(x=c(0.01, 0.2)), aes(x=x)) +
   stat_function(fun=hPerRelVol(6, r), geom="line", color="red") +
   xlab("relative volume") +
   ylab("h")
+
+
+### as simulation
+
+inSliceFracFull <- function(n, nPt=100000){
+  Vectorize(
+  function(h){
+    d <- geozoo::sphere.solid.random(n, nPt)$points
+    anchor <- rep(0, n)
+    plane <- tourr::basis_init(n, 2)
+    distV <- tourr::anchored_orthogonal_distance(plane, d, anchor)
+    sum(distV < h) / nPt
+  }
+  )
+}
+
+inSliceFracHollow <- function(n, nPt=100000){
+  Vectorize(
+  function(h){
+    d <- geozoo::sphere.hollow(n, nPt)$points
+    anchor <- rep(0, n)
+    plane <- tourr::basis_init(n, 2)
+    distV <- tourr::anchored_orthogonal_distance(plane, d, anchor)
+    sum(distV < h) / nPt
+  }
+  )
+}
+
+ggplot(data.frame(x=c(0.01, 0.5)), aes(x=x)) +
+  stat_function(fun=inSliceFracFull(3), geom="line", color="black") +
+  stat_function(fun=inSliceFracFull(4), geom="line", color="blue") +
+  stat_function(fun=inSliceFracFull(5), geom="line", color="green") +
[…]

### 2019-08-29 paper.Rmd
commit before meeting

@@
 
-XXX R code
+```{r codeExampel, echo=TRUE, eval=FALSE}
+library(tourr)
+# use geozoo to generate points on a hollow 3d sphere
+sphere3 <- geozoo::sphere.hollow(3)$points
+colnames(sphere3) <- c("x1", "x2", "x3") # naming variables
+# slice tour animation with default settings
+animate_slice(sphere3)
+# trying an off-center anchor point, thicker slice, and
+# we use pch=26 to hide points outside the slice
+anchor3 <- rep(0.7, 3)
+animate_slice(sphere3, anchor = anchor3, eps = 0.005, pch_other=26)
+```
 

### 2019-08-29 relative_vol_area.R

@@ -9,11 +9,11 @@ relVol <- function(n, r){
   }
 }
 
-#WRONG
+#almost right
 relVol <- function(n, r){
   Vectorize(
   function(h){
-    (h/r)*(1/sqrt(pi))*(gamma((n/2)+1) / gamma(((n-1)/2) + 1))
+    ((2*h/r)^(n-2)) * (pi^(1-n/2)) * (gamma((n/2)+1))
   }
   )
 }
@@ -54,10 +54,6 @@ ggplot(data.frame(x=c(0.01, 0.5)), aes(x=x)) +
   stat_function(fun=relVol(4, r), geom="line", color="blue") +
   stat_function(fun=relVol(5, r), geom="line", color="green") +
   stat_function(fun=relVol(6, r), geom="line", color="red") +
-  stat_function(fun=relArea(3, r), geom="line", color="black", linetype="dashed") +
-  stat_function(fun=relArea(4, r), geom="line", color="blue", linetype="dashed") +
-  stat_function(fun=relArea(5, r), geom="line", color="green", linetype="dashed") +
-  stat_function(fun=relArea(6, r), geom="line", color="red", linetype="dashed") +
   xlab("h") +
   ylab("relative volume/area")
 

### 2019-08-29 paper.Rmd
intro

@@
 
+<!-- Background -->
 Modern data is typically high-dimensional and thus data visualisation often relies on some form of dimension reduction.
-This entails looking at (non-)linear projections of the data.
-Interactive and dynamic displays often provide information beyond what can be achieved in a static display.
-For example, the grand tour [@As85; @BCAH05] shows a smooth sequence of interpolated lower dimensional projections and allows to extrapolate the high-dimensional shape of the distribution.
+This entails looking at linear projections of the data, or nonlinear if one considers techniques like multidimensional scaling (MDS) or t-Distributed Stochastic Neighbor Embedding (t-SNE).
+For our purposes we focus on linear projections.
+Interactive and dynamic displays can provide information beyond what can be achieved in a static display.
+For example, the grand tour [@As85; @BCAH05] shows a smooth sequence of interpolated lower dimensional projections and allows the viewer to extrapolate from the low-dimensional shapes to the multidimensional distribution.
 It is particularly useful for detecting clusters, outliers and non-linear dependence.
+
+<!-- The disadvantage of projection: opacity, occlusion, even with pseudo-transparency from scatterplots;  motivation -->
 Additionaly, we may complement projections with section information.
@@
 
+<!-- Our approach: introduce sections; many different ways to slice, refer to linked brushing as slicing whcih mirrors prosection, but here we will explain/utilise a simple slicing on the projection where point in orthogonal space are faded.
+-->
 The complementarity between projections and sections has previously been discussed in @prosection, dubbing the combination "prosections".
@@
 For example, @prosection uses prosections to reveal the internal structure of the four-point ultrametric locus [@locus].
-
 Since existing approaches to sectioning are interactive, they rely on the user to redefine the section many times to explore the data.
@@
 
+<!-- paper structure --> 
 The relevant definitions and an overview of the implementation in R are given in Section XXX, and in Section XXX.

### 2019-08-29 relative_vol_area.R
playing with relative volumen, still not getting it right

@@ -13,7 +13,7 @@ relVol <- function(n, r){
 relVol <- function(n, r){
   Vectorize(
   function(h){
-    ((2*h/r)^(n-2)) * (pi^(1-n/2)) * (gamma((n/2)+1))
+    (((2*h)/r)^(n-2)) * (pi^((4-n)/2)) * (gamma((n/2)+1))
   }
   )
 }
@@ -121,3 +121,41 @@ ggplot(data.frame(x=c(0.01, 0.5)), aes(x=x)) +
   stat_function(fun=inSliceFracHollow(6), geom="line", color="red", linetype="dashed") +
   xlab("h") +
   ylab("relative volume/area")
+
+
+relVolCube <- function(p, r){
+  Vectorize(
+    function(h){
+      (h/(2*r))^(p-2)
+    }
+  )
+}
+
+ggplot(data.frame(x=c(0.01, 0.5)), aes(x=x)) +
+  stat_function(fun=relVolCube(3, r), geom="line", color="black") +
+  stat_function(fun=relVolCube(4, r), geom="line", color="blue") +
+  stat_function(fun=relVolCube(5, r), geom="line", color="green") +
+  stat_function(fun=relVolCube(6, r), geom="line", color="red") +
+  xlab("h") +
+  ylab("relative volume/area")
+
+
+hForVrel <- function(vrel, r){
+  Vectorize(
+    function(p){
+      (2 * r * vrel^(2-p))
+    }
+  )
+}
+
+#### completely off....h up to 1e+46 does not make sense...
+ggplot(data.frame(x=c(3, 20)), aes(x=x)) +
+  stat_function(fun=hForVrel(0.001, r), geom="line", color="black") +
+  stat_function(fun=hForVrel(0.01, r), geom="line", color="blue") +
+  stat_function(fun=hForVrel(0.05, r), geom="line", color="green") +
+  stat_function(fun=hForVrel(0.1, r), geom="line", color="red") +
+  xlab("p") +
+  ylab("h") +
+  scale_y_log10()
+
+

### 2019-08-29 relative_vol_area.R
Seems I get h/r as function of p right for the cube now

@@ -140,22 +140,20 @@ ggplot(data.frame(x=c(0.01, 0.5)), aes(x=x)) +
   ylab("relative volume/area")
 
 
-hForVrel <- function(vrel, r){
+hrelForVrel <- function(vrel, r){
   Vectorize(
     function(p){
-      (2 * r * vrel^(2-p))
+      (2 * r * vrel^(1/(p-2)))/(2*r)
     }
   )
 }
 
-#### completely off....h up to 1e+46 does not make sense...
-ggplot(data.frame(x=c(3, 20)), aes(x=x)) +
-  stat_function(fun=hForVrel(0.001, r), geom="line", color="black") +
-  stat_function(fun=hForVrel(0.01, r), geom="line", color="blue") +
-  stat_function(fun=hForVrel(0.05, r), geom="line", color="green") +
-  stat_function(fun=hForVrel(0.1, r), geom="line", color="red") +
+ggplot(data.frame(x=c(3, 10)), aes(x=x)) +
+  stat_function(fun=hrelForVrel(0.001, r), geom="line", color="black") +
+  stat_function(fun=hrelForVrel(0.01, r), geom="line", color="blue") +
+  stat_function(fun=hrelForVrel(0.05, r), geom="line", color="green") +
+  stat_function(fun=hrelForVrel(0.1, r), geom="line", color="red") +
   xlab("p") +
-  ylab("h") +
-  scale_y_log10()
+  ylab("h/r")
 
 

### 2019-08-29 paper.Rmd
Working on the introduction

@@
 <!-- The disadvantage of projection: opacity, occlusion, even with pseudo-transparency from scatterplots;  motivation -->
-Additionaly, we may complement projections with section information.
-This is often possible in interactive graphics systems by means of brushing [@ggobiCSDA2003;@JSSv081i05], where the user is selecting a region in a subspace of the dataspace that is highlighted in a projection of the data points.
+A major limitation of projections is their opacity, even if mitigated somewhat through the pseudo-transparancy present in sparse scatterplots.
+As a result certain features of distributions may easily be hidden in projections, especially in the case of concave structures.
+Considering the example of a simple geometric shape such as a hypershere, we understand that projections cannot distinguish between a full or a hollow sphere.
+Similarly we can think of small scale structures hidden in the center of a multivariate distribution.
+Again projections will not help in findig such features in a distribution.
+Other examples where the use of projections is limited include the visualisation of non-linear boundaries between groups, that will be obscured in each projection.
 
@@
 -->
-The complementarity between projections and sections has previously been discussed in @prosection, dubbing the combination "prosections".
-These prosections allow the identification of lower-dimensional structure embedded in the high-dimensional space.
-One simple example is that prosections can reveal the difference between a full and hollow hypersphere.
-In addition to hollowness (concave structures) they can also provide a clear picture of internal structures that may be concealed in a projection.
-For example, @prosection uses prosections to reveal the internal structure of the four-point ultrametric locus [@locus].
-Since existing approaches to sectioning are interactive, they rely on the user to redefine the section many times to explore the data.
-While this level of control is often preferred, it is impractical for detecting and exploring unknown structures in a dataset.
-As a result smaller hidden features that are only visible in specific sections of the full distribution may easily be missed.
-In addition having a dynamic approach to sectioning, similar to the grand tour approach to projections, will provide the viewer with context allowing to extrapolate e.g.
-the shape of a concave structure.
-We therefore propose a more systematic approach to sectioning based on interpolated projections obtained with a grand tour.
-The method defines a slice based on 2-d projections provided by the tour and will presents the viewer with an interpolated sequence of slices when running a tour.
+We can overcome these limitations when adding section information, see e.g.
+@prosection.
+For our example of the hypersphere, we can clearely identify a hollow sphere when looking at a slice throught the distribution, where all points will lie on a circle in a 2d slice.
+In practice there are many different ways to slice, typically facilitated through linked brushing (e.g.
+@ggobiCSDA2003, @JSSv081i05), where the user is selecting a region in a subspace of the dataspace that is highlighted in a projection of the data points.
+These systems are interactive, and rely on the user to redefine the section many times to fully explore the data in this way.
+In this note we introduce a simple slicing method for projections where points in the orthogonal space are faded.
+Combining it with the grand tour results in a new dynamic display that can be used to systematically search for features hidden in sections.
 
 <!-- paper structure --> 
-The relevant definitions and an overview of the implementation in R are given in Section XXX, and in Section XXX.
-We then apply the newly defined slice tour to several examples in Section XXX.
-We consider different high-dimensional geometric shapes and XXX, showing how concave structures can be visualised and explored with a slice tour.
+The method and its implementation in R are described in Section XXX, and in Section XXX.
+We then apply it to several examples in Section XXX.
+We consider different high-dimensional geometric shapes and XXX, showing how concave or occluded structures can be visualised and explored with a slice tour.
 Possible extensions are discussed in Section XXX.
-
-<!--
-A good way to visualize high-dimensional data is to display a smooth sequence of interpolated lower dimensional projections.
-This is called a tour [@As85; @BCAH05] and allows to extrapolate the high-dimensional shape of the distribution.
-It is particularly useful for detecting clusters, outliers and non-linear dependence.
-A limitation of this approach is that concavities in a distribution may be hidden when looking at projections.
-A simple example are projections of samples on a hypersphere.
-No projection of such a distribution allows to distinguish a hollow sphere from a full one.
-On the other hand, the difference would be obvious when looking at a slice through the center of the sphere.
-
-Slicing or sectioning can be thought of as complementary to projections when visualizing high-dimensional data, as discussed in detail in @prosection.
-To overcome the respective shortcomings of each method, it is suggested to use a composition of sections and projections, so-called "prosections".
-In practice sectioning is typically performed in interactive graphics displays in the form of brushing [XXX reference?], i.e.
-the user is selecting a region in a subspace of the high-dimensional dataspace that is highlighted in a projection of the data points.
-More recently @JSSv081i05 suggested conditional visualization for statistical models by selecting sections in data space for detailed comparisons of a fitted model to the observed data.
-Similar to brushing, the section is selected interactively by the user, by choosing a point in low-dimensional displays showing the data in the space of conditioning predictors.
-
-Here we propose a systematic approach to slicing that does not rely on the user choosing sections to display, but will present the viewer with an interpolated sequence showing different slices of parameter space.
-The method is based on the grand tour, and uses the orthogonal distance from the current plane to define the section corresponding to each view in an interpolated tour path.
-We demonstrate how this can be used to gain insights into the concave structure of a dataset, using the example of hollow hyperspheres defined in 3 or more dimensions.
--->
 

### 2019-08-29 relative_vol_area.R

@@ -148,7 +148,7 @@ hrelForVrel <- function(vrel, r){
   )
 }
 
-ggplot(data.frame(x=c(3, 10)), aes(x=x)) +
+ggplot(data.frame(x=c(3, 50)), aes(x=x)) +
   stat_function(fun=hrelForVrel(0.001, r), geom="line", color="black") +
   stat_function(fun=hrelForVrel(0.01, r), geom="line", color="blue") +
   stat_function(fun=hrelForVrel(0.05, r), geom="line", color="green") +

### 2019-08-30 paper.Rmd
Thickness revisited.

@@
 
-XXXX maybe the problem is that each plane is 2d, but the cap assumes I am cutting with a hyperplane that is n-1, so it works as expected in 3d, beyond that i need to account for volume effects in the n-2 dimensions, but how?
-e.g.
-consider 1-d sliced of 3d sphere, in that case I get the volume of the slice as a pipe of radius h, so for n-d sphere I need the volume of n-1 d sphere of radius h but translated across the plane (but what does that really mean)?
+Choosing the slice thickness is a compromise between what feature size can be revealed and how many points fall inside a thin slice.
+As dimensionality increases, the relative number of points that are inside a slice of fixed thickness will decrease.
+The exact relation depends heavily on the distribution of the data points.
+Here we present expressions for spectial cases to get an idea of this scaling.
 
-The thickness of the slice should be chosen such that enough points are inside the slice be representative of the distribution.
-This will depend on the total number of points in the dataset, its original dimensionality and on the distribution.
-Here we estimate the number of points inside a slice of thickness $h$ for the rotation invariant example of an $n$ dimensional full hypersphere with radius $R$, assuming uniformly distributed points throughout the sphere.
-This means we calculate the relative volume inside the slice as a function of $h$ and $n$.
+We first consider as a simple example a hypercube that is sliced along two of the paremeters (note that a cube is not rotation invariant).
+To estimate the relative number of points in a slice of thickness $2h$ we calculate the relative volume of the slice compared to the full hypercube:
+\begin{equation}
+V_{rel} = \frac{V_{slice}}{V_{cube}} = \frac{(2r)^2 * (2h)^{p-2}}{(2r)^p} = (\frac{h}{r})^{p-2},
+\end{equation}
+where the cube side lenght is $2r$.
+Thus if we want to keep the relative number of points (i.e.
+$V_{rel}$) constant for slices in different $p$ we calculate $h$ from a fixed volume parameter as $h = \epsilon^{1/(p-2)}$.
 
-The total volume of the full hypersphere is given as
+Considering now slicing through a multivariate normal, we can estimate the number of points in a slice by integrating the probability distribution with boundaries $\pm h$ in $p-2$.
+In this case the relative density scales with the error function as
 \begin{equation}
-V_n(R) = \frac{\pi^{n/2}}{\Gamma(\frac{n}{2} + 1)} R^n.
+\rho_{rel} = \pi^{(2-p)/2} \mathrm{erf}((h/\sqrt 2 \sigma)) ^{p-2}.
 \end{equation}
-To calculate the volume of the slice we subtract the volume of a hypersphere cap of hight $h'=r-h$, as given in @capVol,
-\begin{equation}
-V^{\phi}_n(R) = \frac{1}{2}V_n(R) I_{sin^2\phi}(\frac{n+1}{2},\frac{1}{2}),
-\end{equation}
-where $I$ is the regularized incomplete beta function and $\phi =\arccos(1-\frac{h'}{r}) = \arccos(\frac{h}{r})$ the colatitude angle (XXX got this off the paper, don't yet understand the angle..).
-The relative volume of the slice is thus given as (XXX better names needed!)
-\begin{equation}
-V_n^h(R) = \frac{V_n(R) - 2 V^{\phi}_n(R)}{V_n(R)} = 1 - I_{1-\frac{h^2}{r^2}}(\frac{n+1}{2},\frac{1}{2}),
-\end{equation}
-using $\sin(\arccos(x)) = \sqrt{1-x^2}$.
 
-In the case of a hollow hypersphere we are instead interested in the relative area inside the slice. We use the cap area derived in @capVol to calculate the area inside the slice as
-\begin{equation}
-A_n^h(R) = \frac{A_n(R) - 2 A^{\phi}_n(R)}{A_n(R)} = 1 - I_{1-\frac{h^2}{r^2}}(\frac{n-1}{2},\frac{1}{2}).
-\end{equation}
+XXX two plots: vrel/rhorel scaling with h for n for p = 3, 4, 5, 10, and h scaling with p for vrel = 0.001, 0.01, 0.05 and 0.1
 

### 2019-08-30 relative_vol_area.R
hypercube vs gaussian...

@@ -1,158 +1,63 @@
 
 
-relVol <- function(n, r){
-  function(h){
-    x <- 1 - (h/r)^2
-    a <- (n+1) / 2
-    b <- 1/2
-    1 - pbeta(x, a, b)
-  }
-}
-
-#almost right
-relVol <- function(n, r){
+relVolCube <- function(p){
   Vectorize(
-  function(h){
-    (((2*h)/r)^(n-2)) * (pi^((4-n)/2)) * (gamma((n/2)+1))
-  }
+    function(x){
+      (x)^(p-2)
+    }
   )
 }
 
-relArea <- function(n, r){
-  function(h){
-    x <- 1 - (h/r)^2
-    a <- (n-1) / 2
-    b <- 1/2
-    1 - pbeta(x, a, b)
-  }
-}
+erf <- function(x) 2 * pnorm(x * sqrt(2)) - 1
+erfinv <- function (x) qnorm((1 + x)/2)/sqrt(2)
 
-pbetaX <- function(n, r){
-  function(x){
-    pbeta(1-x^2, n/2, 1/2)
-  }
-}
-
-r <- 1
-
-ggplot(data.frame(x=c(0.01, 0.5)), aes(x=x)) +
-  stat_function(fun=relVol(3, r), geom="line", color="black") +
-  stat_function(fun=relVol(4, r), geom="line", color="purple") +
-  stat_function(fun=relVol(6, r), geom="line", color="blue") +
-  stat_function(fun=relVol(10, r), geom="line", color="green") +
-  stat_function(fun=relVol(20, r), geom="line", color="red") +
-  stat_function(fun=relArea(3, r), geom="line", color="black", linetype="dashed") +
-  stat_function(fun=relArea(4, r), geom="line", color="purple", linetype="dashed") +
-  stat_function(fun=relArea(6, r), geom="line", color="blue", linetype="dashed") +
-  stat_function(fun=relArea(10, r), geom="line", color="green", linetype="dashed") +
-  stat_function(fun=relArea(20, r), geom="line", color="red", linetype="dashed") +
-  xlab("h") +
-  ylab("relative volume/area")
-
-ggplot(data.frame(x=c(0.01, 0.5)), aes(x=x)) +
-  stat_function(fun=relVol(3, r), geom="line", color="black") +
-  stat_function(fun=relVol(4, r), geom="line", color="blue") +
-  stat_function(fun=relVol(5, r), geom="line", color="green") +
[…]

### 2019-08-30 paper.Rmd
Adding plots

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = FALSE, message = FALSE)
 ```
@@
 
-XXX Example 2x2, row 1-col 1 has a slice projection  of a solid 4D spehere, row 1-col 2 has a slice projectionof a hollow 4D sphere; row 2-col 1 has a slice projection of a  solid 4D cube, row 1-col 2 has a slice projection of a hollow 4D cube
+```{r renderFunction, echo=FALSE, eval=TRUE, warning=FALSE, fig.height=1.5, fig.width = 5.5, fig.cap="Projected slices through 4d geometric shapes.
+On the left a full (hollow) hypersphere (a, b), on the right a full (hollow) hypercube (c, d).
+Points inside the slice are shown as black bullets, points outside the slice are shown as grey dots.
+We can clearly distinguish the full from the hollow objects based on the slice display."}
+library(tidyverse)
+
+# slight modification of the function in tourr::display_slice
+render_data <- function(data, proj, h, anchor) {
+    # Render projected points
+    x <- data %*% proj
+    d <- tourr::anchored_orthogonal_distance(proj, data, anchor)
+    pch <- rep(46, nrow(x))
+    pch[d < h] <- 20
+    col <- rep("grey", nrow(x))
+    col[d < h] <- "black"
+    ggplot() +aes(x[,1], x[,2]) + geom_point(pch = pch, color=col) + theme_void() + coord_fixed()
+}
+
+anchor <- rep(0, 4)
+plane <- tourr::basis_init(4, 2)
+nPt <- 5000
+
+set.seed(3008)
+
+d1 <- geozoo::sphere.solid.random(4, nPt)$points
+d2 <- geozoo::sphere.hollow(4, nPt)$points
+d3 <- geozoo::cube.solid.random(4)$points[1:nPt,] %>% scale(scale = FALSE)
+d4 <- geozoo::cube.face(4)$points[1:nPt,] %>% scale(scale = FALSE)
+
+a <- render_data(d1, plane, 0.2, anchor) + labs(title="a")
+b <- render_data(d2, plane, 0.2, anchor) + labs(title="b")
+c <- render_data(d3, plane, 0.2, anchor) + labs(title="c")
+d <- render_data(d4, plane, 0.2, anchor) + labs(title="d")
+
+library(gridExtra)
+grid.arrange(a, b, c, d, nrow=1)
+
+```
 

### 2019-08-30 paper.Rmd
Notation fix

@@
 For our example of the hypersphere, we can clearely identify a hollow sphere when looking at a slice throught the distribution, where all points will lie on a circle in a 2d slice.
-In practice there are many different ways to slice, typically facilitated through linked brushing (e.g.
+Note that there are many ways to define a slice through high-dimensional data.
+In practice, slicing is typically facilitated through linked brushing (e.g.
 @ggobiCSDA2003, @JSSv081i05), where the user is selecting a region in a subspace of the dataspace that is highlighted in a projection of the data points.
-These systems are interactive, and rely on the user to redefine the section many times to fully explore the data in this way.
+These systems are interactive, and rely on the user to redefine the section many times to explore the data in this way.
 In this note we introduce a simple slicing method for projections where points in the orthogonal space are faded.
@@
 
-A tour is providing a continuous animation, showing a sequence of $d$-dimensional (typically $d=1$ or $2$) projections from $n$-dimensional parameter space.
+A tour is providing a continuous animation, showing a sequence of $d$-dimensional (typically $d=1$ or $2$) projections from $p$-dimensional parameter space.
 The tour combines basis selection with basis interpolation.
@@
 
-The notation used in the following for the description of each projection is denoting $\mathbf{X}$ the $m\times n$ dimensional data matrix ($m$ observations in $n$ dimensions) and $\mathbf{A}$ an orthonormal $n \times d$ projection matrix.
+The notation used in the following for the description of each projection is denoting $\mathbf{X}$ the $n\times p$ dimensional data matrix ($n$ observations in $p$ dimensions) and $\mathbf{A}$ an orthonormal $p \times d$ projection matrix.
 The $d$-dimensional projection of the data is thus given by $\mathbf{Y} = \mathbf{X}\cdot\mathbf{A}$, giving the $n\times d$ dimensional projected data matrix to be plotted in each frame of the tour display.
@@
 \end{equation}
-with $x_i, i=1,...,m$ an $n$-dimensional observation in $\mathbf{X}$ and $a_i, i=1,2$ denoting the columns of the projection matrix, $\mathbf{A}=(a_1, a_2)$.
+with $x_i, i=1,...,n$ an $p$-dimensional observation in $\mathbf{X}$ and $a_i, i=1,2$ denoting the columns of the projection matrix, $\mathbf{A}=(a_1, a_2)$.
 Calculating the euclidean norm of $x'$ is straightforward thanks to the orthonormality of $\mathbf{A}$, and gives the components of the required distance vector.
@@
 The distance vector can then be used to display a slice tour by highlighting points for which the orthogonal distance is smaller than a selected cutoff value $h$.
-In general a larger cutoff value and/or much larger sample sizes are required as the original dimensionality ($n$) increases.
-For practical purposes we therefore specify a volume $\epsilon$ and then use as a cutoff $h = \epsilon^{1/n}$.
+In general a larger cutoff value and/or much larger sample sizes are required as the original dimensionality ($p$) increases.
+For practical purposes we therefore specify a volume $\epsilon$ and then use as a cutoff $h = \epsilon^{1/(p-2)}$.
+XXX not yet implemented that way, based on thickness estimates
 
@@
 
-To demonstrate the slice tour we sample points on the surface of a sphere with radius $r=1$ in $n=3$ and $n=5$ dimensions.
+To demonstrate the slice tour we sample points on the surface of a sphere with radius $r=1$ in $p=3$ and $p=5$ dimensions.
 We use the `geozoo` package [@geozoo] to draw 2000 (5000) points on the 3d (5d) sphere.
@@
 $h = 0.1$ in 3d and $h = 0.25$ in 5d.
-Slicing through the origin results in sections where the points inside the slice are approximately on a circle with the full radius.
-Snapshots of the slice tour animation are shown in Figure \ref{fig:centered}, animated gifs are available at XXX.
-Notice that the larger number of points and increased thickness parameter $h$ do not make up for the reduced number of points inside the slice through the 5d hypersphere.
-(projected points are on average closer to the center as we increase the dimension of the hypersphere, is there some convergence similar to central limit theorem at play?)
+Slicing through the origin results in sections where the points inside the slice are approximately on a circle with the full radius, and each view is similar to that shown in Figure~\ref{fig:renderFunction}.
+The animated gifs are available at XXX.
 
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.3\textwidth]{gifs/png/sphere-3-centered-055.png}\hspace{1cm}
-\includegraphics[width=0.3\textwidth]{gifs/png/sphere-5-centered-055.png}}
-\caption{Slice through a 3d (left) and 5d (right) hollow sphere, anchored at the origin.
-Points inside the slice are highlighted with a bullet marker and are found on a circle with the radius of the hypersphere, $r$.}
-\label{fig:centered}
-\end{figure*}
-
-It is also instructive to look at the sliced spheres with an off-center anchor point.
+It is especially instructive to look at the sliced spheres with an off-center anchor point.
 In this case the different views obtained in the slice tour reveal more of the concave structure of the distribution.
@@
 \begin{figure*}[ht]
-\centerline{\includegraphics[width=0.3\textwidth]{gifs/png/sphere-3-anchored-033.png}
-\includegraphics[width=0.3\textwidth]{gifs/png/sphere-3-anchored-036.png}
-\includegraphics[width=0.3\textwidth]{gifs/png/sphere-3-anchored-041.png}}
+\centerline{\includegraphics[width=0.2\textwidth]{gifs/png/sphere-3-anchored-033.png}
+\hspace{5mm}
+\includegraphics[width=0.2\textwidth]{gifs/png/sphere-3-anchored-036.png}
+\hspace{5mm}
+\includegraphics[width=0.2\textwidth]{gifs/png/sphere-3-anchored-041.png}}
 \caption{Different slices through a 3d hollow sphere with $r=1$, with anchor point $(0.7, 0.7, 0.7)$, showing full circle with small radius (left), circle with radius $<r$ (middle) and circle with radius $r$ (right).}
@@
 
-Instead of slicing in the full $n-d=n-2$ dimensional orthogonal space we can also use a subspace in which to slice.
+Instead of slicing in the full $p-d=p-2$ dimensional orthogonal space we can also use a subspace in which to slice.
 As pointed out in @prosection combining projections with sectioning in varying numbers of dimensions can reveal the codimension of an unknown distribution.
@@
 
+In addition slice definitions beyond a simple cutoff on the orthogonal distance should be considered, where different definition can be used to highlight certain aspects of the data.
 We also expect interesting new developments by combining the slice tour and the guided tour into a weighted guided tour with index functions using the distance vector information in addition to the projected data points.

### 2019-09-02 paper.Rmd
Thickness for spheres

@@
   \def\tightlist{}
+  \usepackage{setspace}\doublespacing
 output: rticles::tf_article
@@
 The distance vector can then be used to display a slice tour by highlighting points for which the orthogonal distance is smaller than a selected cutoff value $h$.
-In general a larger cutoff value and/or much larger sample sizes are required as the original dimensionality ($p$) increases.
-For practical purposes we therefore specify a volume $\epsilon$ and then use as a cutoff $h = \epsilon^{1/(p-2)}$.
-XXX not yet implemented that way, based on thickness estimates
+Notice that in 3d, where for each plane we have one well-defined orthogonal direction, this defines a flat slice of hight $2*h$, matching the common conception of a slice.
+When going beyond 3d, the codimension of the plane is $>1$ and using the euclidean distance results in rotation invariant slicing in the orthogonal space.
+As a consequence the "slices" are now spherical in the orthogonal subspace, with radius $h$.
 
@@
 Choosing the slice thickness is a compromise between what feature size can be revealed and how many points fall inside a thin slice.
-As dimensionality increases, the relative number of points that are inside a slice of fixed thickness will decrease.
+As dimensionality $p$ increases, the relative number of points that are inside a slice of fixed thickness $h$ will decrease.
 The exact relation depends heavily on the distribution of the data points.
-Here we present expressions for spectial cases to get an idea of this scaling.
+Here we present expressions for special cases to get an idea of this scaling.
+For practical purposes we specify a volume $\epsilon$ and then use as a cutoff $h = \epsilon^{1/(p-2)}$.
+XXX not yet implemented that way, based on thickness estimates.
 
-We first consider as a simple example a hypercube that is sliced along two of the paremeters (note that a cube is not rotation invariant).
-To estimate the relative number of points in a slice of thickness $2h$ we calculate the relative volume of the slice compared to the full hypercube:
+We consider two types of rotation invariant distributions for our estimates: a hypersphere in $p$ dimensions and a $p$ dimensional normal distribution without correlation.
+
+Let us consider points uniformly distributed in a $p$-dimensional hypersphere.
+The fraction of points inside a slice of thickness $2*h$ can therefore be estimated as the relative volume of the slice compared to the volume of the full hypersphere,
 \begin{equation}
-V_{rel} = \frac{V_{slice}}{V_{cube}} = \frac{(2r)^2 * (2h)^{p-2}}{(2r)^p} = (\frac{h}{r})^{p-2},
+V_{rel} = \frac{1}{2} \frac{h^{p-2}}{r^p} (p r^2 - (p-2) h^2) \approx \frac{1}{2} (\frac{h}{R})^{p-2},
 \end{equation}
-where the cube side lenght is $2r$.
+where the approximation is valid when $h \ll r$, as is typically expected to be the case.
 Thus if we want to keep the relative number of points (i.e.
-$V_{rel}$) constant for slices in different $p$ we calculate $h$ from a fixed volume parameter as $h = \epsilon^{1/(p-2)}$.
+$V_{rel}$) approximately constant for slices in different dimensions $p$, we calculate $h$ from a volume parameter as $h = \epsilon^{1/(p-2)}$.
+XXXX change code to this?
+currently $\epsilon^{1/p}$ is implemented...
 
+XXXX need to redo integration of multivariate normal...
 Considering now slicing through a multivariate normal, we can estimate the number of points in a slice by integrating the probability distribution with boundaries $\pm h$ in $p-2$.

### 2019-09-02 paper.Rmd
Mostly formatting

@@
 bibliography: biblio.bib
+geometry: margin=2.5cm
 abstract: |
@@
   \def\tightlist{}
-  \usepackage{setspace}\doublespacing
+  \usepackage{setspace}
 output: rticles::tf_article
@@
 
+\doublespacing
 # Introduction
@@
 The distance vector can then be used to display a slice tour by highlighting points for which the orthogonal distance is smaller than a selected cutoff value $h$.
-Notice that in 3d, where for each plane we have one well-defined orthogonal direction, this defines a flat slice of hight $2*h$, matching the common conception of a slice.
+Notice that in 3d, where for each plane we have a single orthogonal direction, this defines a flat slice of hight $2*h$, matching the common conception of a slice.
 When going beyond 3d, the codimension of the plane is $>1$ and using the euclidean distance results in rotation invariant slicing in the orthogonal space.
@@
 Points inside the slice are shown as black bullets, points outside the slice are shown as grey dots.
-We can clearly distinguish the full from the hollow objects based on the slice display."}
+We can clearly distinguish the full from the hollow objects based on the slice display.\\label{fig:renderFunction}"}
 library(tidyverse)
@@
 # Examples
-
 XXX choose 6
 
-To demonstrate the slice tour we sample points on the surface of a sphere with radius $r=1$ in $p=3$ and $p=5$ dimensions.
-We use the `geozoo` package [@geozoo] to draw 2000 (5000) points on the 3d (5d) sphere.
-
+## Geometric shapes
+### Hollow sphere
+We sample points on the surface of a sphere with radius $r=1$ in $p=3$ and $p=5$ dimensions.
+Using the `geozoo` package [@geozoo], we draw 2000 (5000) points that are uniformly distributed on the surface of a 3d (5d) sphere.
 For both examples we start by slicing through the origin with the default parameters, in particular $\epsilon = 0.001$, i.e.
@@
 Depending on the viewing angle selected by the tour the slice contains points on the circle with radius $r$ (when the anchor point has a negligible orthogonal component to the viewing plane); a circle with radius $<r$ as the viewing angle is tilted away from the axis connecting the anchor point to the origin; a full circle with small radius as the angle increases; and finally we see an empty slice when the projection plane is orthogonal to the anchor point axis.
+A similar picture is found for the 5d sphere.
+Note however that as dimensionality increases, the larger orthogonal space means that the anchor point will have a large orthogonal component in most views.
 
@@
 
-The same behavior is observed when slicing the 5d hollow sphere with an off-center anchor point, however a large number of sample points is required to get a similarly clear display to that found for the 3d sphere.
-The animation showing the anchored slice tour of the 5d hollow sphere with 5000 sample points is available at XXX.
-In this example the anchor point was fixed to $(0.5,0.5,0.5,0.5,0.5)$.
+### Other geometric shapes
+
+Examples: 4d cube, 4d torus
+
+## Other examples
+
+### Needle in a haystack
+
+### Non-linear boundaries
 

### 2019-09-02 render-examples.R

@@ -29,7 +29,11 @@ render(sphere3, grand_tour(), display_slice(anchor = anchor3), 'png', "gifs/png/
 set.seed(2019)
 render(sphere5, grand_tour(), display_slice(anchor = anchor5), 'png', "gifs/png/sphere-5-anchored-%03d.png", frames = max_f, rescale = FALSE)
 
-
+#testing out different anchor in 5d with large sample
+sphere5 <- geozoo::sphere.hollow(5, 50000)$points
+colnames(sphere5) <- c("x1", "x2", "x3", "x4", "x5")
+anchor5 <- c(1.1, 0, 0, 0, 0)
+animate(sphere5, grand_tour(), display_slice(anchor = anchor5), rescale = FALSE)
 
 ### playing with other exampels
 

### 2019-09-02 paper.Rmd
Small formatting fix
abstract and intro

@@
 abstract: |
-  Projections and sectioning are complementary methods for the visualisation of
-  high-dimensional data. In this note we describe slicing in the orthogonal
+  Taking projections of high-dimensional data is a common analytical and visualisation technique in statistics for working with high-dimensional problems.
+Sectioning, or slicing, through high dimensions is less common, but can be useful for visualising data with concavities, or non-linear structure.
+It is associated with conditional distributions in statistics, and also linked brushing between plots in interactive data visualisation.
+This short technical note describes a simple approach for  slicing in the orthogonal
   space of projections obtained when running a tour, thus presenting the viewer
@@
   implemented in R as an extension to the tourr package, and can be used to
-  explore concave structures in multivariate distributions.
+  explore for concave and non-linear structures in multivariate distributions.
 keywords: |
@@
 <!-- Background -->
-Modern data is typically high-dimensional and thus data visualisation often relies on some form of dimension reduction.
-This entails looking at linear projections of the data, or nonlinear if one considers techniques like multidimensional scaling (MDS) or t-Distributed Stochastic Neighbor Embedding (t-SNE).
-For our purposes we focus on linear projections.
+Data is commonly high-dimensional, and visualisation often relies on some form of dimension reduction.
+This can be taking linear projections, or nonlinear if one considers techniques like multidimensional scaling (MDS) or t-Distributed Stochastic Neighbor Embedding (t-SNE).
+For the purposes here, the focus is on linear projections, in particular as provided by the grand tour [@As85; @BCAH05].
 Interactive and dynamic displays can provide information beyond what can be achieved in a static display.
-For example, the grand tour [@As85; @BCAH05] shows a smooth sequence of interpolated lower dimensional projections and allows the viewer to extrapolate from the low-dimensional shapes to the multidimensional distribution.
+The grand tour shows a smooth sequence of interpolated low dimensional projections, and allows the viewer to extrapolate from the low-dimensional shapes to the multidimensional distribution.
 It is particularly useful for detecting clusters, outliers and non-linear dependence.
@@
 <!-- The disadvantage of projection: opacity, occlusion, even with pseudo-transparency from scatterplots;  motivation -->
-A major limitation of projections is their opacity, even if mitigated somewhat through the pseudo-transparancy present in sparse scatterplots.
-As a result certain features of distributions may easily be hidden in projections, especially in the case of concave structures.
-Considering the example of a simple geometric shape such as a hypershere, we understand that projections cannot distinguish between a full or a hollow sphere.
-Similarly we can think of small scale structures hidden in the center of a multivariate distribution.
-Again projections will not help in findig such features in a distribution.
-Other examples where the use of projections is limited include the visualisation of non-linear boundaries between groups, that will be obscured in each projection.
+A major limitation of projections is their opacity.
+It is mitigated some, when scatterplots are used to render the projected data, through a pseudo-transparancy of sparseness of points.
+This is improved further if points are also drawn using some transparency.
+Some features of multivariate distributions may be visible, but a lot is easily hidden, especially in the case of concave structures.
+Considering the example of a simple geometric shape such as a hypershere, it is difficult to distinguish between a full or a hollow sphere, that is, whether points are uniformly distributed within the sphere or on the surface of a sphere.
+Similarly we can think of small scale structures hidden in the center of a multivariate distribution, that might be considered to be "needles in a haystack", which can be difficult to detect in projections.
+Another example is where one would like to examine non-linear boundaries between groups from a classification model, with the nature of the boundaries being occluded in the projections.
 
@@
 -->
-We can overcome these limitations when adding section information, see e.g.
-@prosection.
-For our example of the hypersphere, we can clearely identify a hollow sphere when looking at a slice throught the distribution, where all points will lie on a circle in a 2d slice.
-Note that there are many ways to define a slice through high-dimensional data.
-In practice, slicing is typically facilitated through linked brushing (e.g.
-@ggobiCSDA2003, @JSSv081i05), where the user is selecting a region in a subspace of the dataspace that is highlighted in a projection of the data points.
-These systems are interactive, and rely on the user to redefine the section many times to explore the data in this way.
-In this note we introduce a simple slicing method for projections where points in the orthogonal space are faded.
-Combining it with the grand tour results in a new dynamic display that can be used to systematically search for features hidden in sections.
+Slicing, or sectioning, is a way that the internal distribution of high-dimensional data can be explored.
+@prosection discusses a technique for combining projections and sections, by slicing in the back-dimensions.
+It is also possible to think of linked brushing (e.g.
+@ggobiCSDA2003, @JSSv081i05) as slicing.
+In this note we discuss another approach to sectioning, where observations in the space orthogonal to the projection are highlighted if they are close to the projection plane, and faded if further afield.
+This is combined with the grand tour to provide a new dynamic display that can be used to systematically search for features hidden in high dimensions.
 
 <!-- paper structure --> 
-The method and its implementation in R are described in Section XXX, and in Section XXX.
-We then apply it to several examples in Section XXX.
-We consider different high-dimensional geometric shapes and XXX, showing how concave or occluded structures can be visualised and explored with a slice tour.
-Possible extensions are discussed in Section XXX.
+The method is described in Section XXX, and it is implemented in the tourr [@tourr] package in R [@rref].
+Section XXX illustrates the use on several high-dimensional geometric shapes and established data sets, showing how concave or occluded structures can be visualised and explored with a slice tour.
+Future work is discussed in Section XXX.
 
@@
 \end{equation}
-
 where the cross term can be expressed as
-
 \begin{equation}

### 2019-09-02 paper.Rmd
Examples added

@@
 # Examples
-XXX choose 6
-
 ## Geometric shapes
@@
 ### Other geometric shapes
+To better understand the slice tour we look at different examples of geometric shapes, see Figure \ref{fig:geoms}.
+For each shape two selected views are shown.
+The first column shows views from the slice tour on a 3d Roman surface.
+The second column shows slices through a 4d torus, revealing different aspects of the shape.
+The last column shows a 6d cube, where the upper plot shows a view along two of the parameter combinations, allowing to clearly identify the rectangular shape.
+The panel below shows that this is not typically the case when looking at a randomly selected slice.
 
-Examples: 4d cube, 4d torus
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.2\textwidth]{gifs/png/roman-surface-centered-010.png}
+\hspace{5mm}
+\includegraphics[width=0.2\textwidth]{gifs/png/torus-4-centered-015.png}
+\hspace{5mm}
+\includegraphics[width=0.2\textwidth]{gifs/png/cube-6-centered-001.png}}
+\centerline{
+\includegraphics[width=0.2\textwidth]{gifs/png/roman-surface-centered-032.png}
+\hspace{5mm}
+\includegraphics[width=0.2\textwidth]{gifs/png/torus-4-centered-033.png}
+\hspace{5mm}
+\includegraphics[width=0.2\textwidth]{gifs/png/cube-6-centered-063.png}
+}
+\caption{Different slices through a Roman surface (left), 4d torus (middle) and a 6d cube (right).}
+\label{fig:geoms}
+\end{figure*}
 

### 2019-09-02 render-examples.R

@@ -29,6 +29,52 @@ render(sphere3, grand_tour(), display_slice(anchor = anchor3), 'png', "gifs/png/
 set.seed(2019)
 render(sphere5, grand_tour(), display_slice(anchor = anchor5), 'png', "gifs/png/sphere-5-anchored-%03d.png", frames = max_f, rescale = FALSE)
 
+torus4 <- geozoo::torus(p=4)$points %>%
+  scale()
+colnames(torus4) <- c("x1", "x2", "x3", "x4")
+set.seed(2019)
+render(torus4, grand_tour(), display_slice(eps=0.2), 'png', "gifs/png/torus-4-centered-%03d.png", frames = max_f, rescale = FALSE)
+
+cube6 <- geozoo::cube.face(6)$points %>%
+  center()
+colnames(cube6) <- c("x1", "x2", "x3", "x4", "x5", "x6")
+set.seed(2019)
+render(cube6, grand_tour(), display_slice(), 'png', "gifs/png/cube-6-centered-%03d.png", frames = max_f, rescale = FALSE)
+
+cubeAnchor <- c(1, 0, 0, 0, 0, 0)
+set.seed(2019)
+render(cube6, grand_tour(), display_slice(anchor = cubeAnchor), 'png', "gifs/png/cube-6-anchored-%03d.png", frames = max_f, rescale = FALSE)
+
+rmS <- geozoo::roman.surface()$points %>%
+  scale()
+colnames(rmS) <- c("x1", "x2", "x3")
+set.seed(2019)
+render(rmS, grand_tour(), display_slice(), 'png', "gifs/png/roman-surface-centered-%03d.png", frames = max_f, rescale = FALSE)
+
+
+library(animation)
+data(pollen)
+pollen <- as.matrix(pollen) %>% scale()
+set.seed(2019)
+render(pollen, grand_tour(), display_slice( half_range=3, eps=0.0005, axes = "bottomleft"), 'png', "gifs/png/pollen-%03d.png", frames = max_f, rescale = FALSE)
+
+set.seed(201909)
+render(pollen, grand_tour(), display_slice( half_range=1, eps=0.005, axes = "bottomleft"), 'png', "gifs/png/pollen-zoom-%03d.png", frames = max_f, rescale = FALSE)
+
+library(RColorBrewer)
+wine_radial <- read_csv("data/wine-svm-radial.csv")
+wine_poly <- read_csv("data/wine-svm-poly.csv")
+clrs <- brewer.pal(3, "Dark2")
[…]

### 2019-09-02 paper.Rmd
Adding in last two examples.

@@
 ### Needle in a haystack
+We use the pollen data as an example for a hidden feature generally occluded in projections.
+This 1986 Exposition's dataset was made by David Coleman of RCA Labs.
+The artificially generated data is comprised of 3848 observations on 5 variables.
+We center and scale each variable of the dataset and run a slice tour with a thin slice ($\epsilon = 0.0005$).
+Selected views are shown in the first two plots in Figure \ref{fig:pollen}, indicating the presence of an interesting feature hidden in the center.
+We next zoom in on the center of the distribution and consider a thicker slice ($\epsilon = 0.005$) to uncover the hidden structure, shown in the last two plots of Figure \ref{fig:pollen}.
+
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.2\textwidth]{gifs/png/pollen-007.png}
+\hspace{5mm}
+\includegraphics[width=0.2\textwidth]{gifs/png/pollen-092.png}
+\hspace{5mm}
+\includegraphics[width=0.2\textwidth]{gifs/png/pollen-zoom-075.png}
+\hspace{5mm}
+\includegraphics[width=0.2\textwidth]{gifs/png/pollen-zoom-018.png}
+}
+\caption{Different slices through the pollen dataset. The first two plots have slice volume $\epsilon = 0.0005$, the last two plots zoom in on the center and have increased volume $\epsilon = 0.005$.}
+\label{fig:pollen}
+\end{figure*}
 
 ### Non-linear boundaries
+XXX say something about this dataset [@sam.11271]
+
+Selected views from a slice tour of the scaled and centered data, and with large slice volume $\epsilon = 0.15$ are shown in Figure \ref{fig:wine}.
+The slices allow to clearly see the non-linearity of the boundary between the green and orange group.
+The first view also shows a non-linear boundary of the purple group, which interestingly does not appear at all in the second slice shown here.
+
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.2\textwidth]{gifs/png/wine-029.png}
+\hspace{5mm}
+\includegraphics[width=0.2\textwidth]{gifs/png/wine-042.png}
+\hspace{5mm}
+\includegraphics[width=0.2\textwidth]{gifs/png/wine-049.png}
+\hspace{5mm}
+\includegraphics[width=0.2\textwidth]{gifs/png/wine-100.png}
+}
+\caption{Different slices through the wine dataset.}
+\label{fig:wine}
+\end{figure*}
 

### 2019-09-03 paper.Rmd
Thickness rewritten

@@
 The distance vector can then be used to display a slice tour by highlighting points for which the orthogonal distance is smaller than a selected cutoff value $h$.
-Notice that in 3d, where for each plane we have a single orthogonal direction, this defines a flat slice of hight $2*h$, matching the common conception of a slice.
+Notice that in 3d, where for each plane we have a single orthogonal direction, this defines a flat slice of hight $2h$, matching the common conception of a slice.
 When going beyond 3d, the codimension of the plane is $>1$ and using the euclidean distance results in rotation invariant slicing in the orthogonal space.
@@
 As dimensionality $p$ increases, the relative number of points that are inside a slice of fixed thickness $h$ will decrease.
-The exact relation depends heavily on the distribution of the data points.
-Here we present expressions for special cases to get an idea of this scaling.
-For practical purposes we specify a volume $\epsilon$ and then use as a cutoff $h = \epsilon^{1/(p-2)}$.
-XXX not yet implemented that way, based on thickness estimates.
+The exact relation depends on the distribution of the data points.
+Here we estimate how it scales with $h$ and $p$ for the example of points uniformly distributed in a hypersphere.
+The example was selected since it represents a rotation invariant uniform distribution in $p$ space.
+As we are interested in hollowness this type of distribution is more relevant than a multivariate normal distribution.
 
-We consider two types of rotation invariant distributions for our estimates: a hypersphere in $p$ dimensions and a $p$ dimensional normal distribution without correlation.
-
-Let us consider points uniformly distributed in a $p$-dimensional hypersphere.
-The fraction of points inside a slice of thickness $2*h$ can therefore be estimated as the relative volume of the slice compared to the volume of the full hypersphere,
+The fraction of points inside a slice of thickness $h$ can be estimated as the relative volume of the slice compared to the volume of the full hypersphere,
 \begin{equation}
@@
 $V_{rel}$) approximately constant for slices in different dimensions $p$, we calculate $h$ from a volume parameter as $h = \epsilon^{1/(p-2)}$.
-XXXX change code to this?
-currently $\epsilon^{1/p}$ is implemented...
-
-XXXX need to redo integration of multivariate normal...
-Considering now slicing through a multivariate normal, we can estimate the number of points in a slice by integrating the probability distribution with boundaries $\pm h$ in $p-2$.
-In this case the relative density scales with the error function as
-\begin{equation}
-\rho_{rel} = \pi^{(2-p)/2} \mathrm{erf}((h/\sqrt 2 \sigma)) ^{p-2}.
-\end{equation}
-
-XXX two plots: vrel/rhorel scaling with h for n for p = 3, 4, 5, 10, and h scaling with p for vrel = 0.001, 0.01, 0.05 and 0.1
 

### 2019-09-03 paper.Rmd
Some rewriting

@@
 <!-- paper structure --> 
-The method is described in Section XXX, and it is implemented in the tourr [@tourr] package in R [@rref].
+The method is described in Section XXX, and it is implemented in the `tourr` [@tourr] package in R [@rref].
 Section XXX illustrates the use on several high-dimensional geometric shapes and established data sets, showing how concave or occluded structures can be visualised and explored with a slice tour.
@@
 We first briefly review the relevant concepts.
-Building on these ideas we introduce a method for displaying slices of parameter space that are defined for each projection matrix.
+Building on these ideas we then introduce a method for displaying slices of parameter space that are defined for each projection matrix.
 
@@
 In a grand tour the basis selection is random (each new basis is chosen from all possible projections), and typically geodesic interpolation is used to generate smooth transitions between viewing planes.
-Different basis selection methods as well as the geodesic interpolation are implemented in the `tourr` package [@tourr] in R [@rref], which also provides several display functions for viewing the tour.
+Different basis selection methods as well as the geodesic interpolation are implemented in the `tourr` package [@tourr], which also provides several display functions for viewing the tour.
 
@@
 
-In addition to the projected data $\mathbf{Y}$ we calculate for each projection the orthogonal distance of the data points from the current projection plane centered at the origin, $v$.
+In addition to the projected data $\mathbf{Y}$ we calculate for each projection the orthogonal distance of the data points from the current projection plane centered at the origin.
 The orthogonal component is given by
@@
 \end{equation}
-with $x_i, i=1,...,n$ an $p$-dimensional observation in $\mathbf{X}$ and $a_i, i=1,2$ denoting the columns of the projection matrix, $\mathbf{A}=(a_1, a_2)$.
+with $x_i, i=1,...,n$ a $p$-dimensional observation in $\mathbf{X}$ and $a_i, i=1,2$ denoting the columns of the projection matrix, $\mathbf{A}=(a_1, a_2)$.
 Calculating the euclidean norm of $x'$ is straightforward thanks to the orthonormality of $\mathbf{A}$, and gives the components of the required distance vector.
@@
 Notice that in 3d, where for each plane we have a single orthogonal direction, this defines a flat slice of hight $2h$, matching the common conception of a slice.
-When going beyond 3d, the codimension of the plane is $>1$ and using the euclidean distance results in rotation invariant slicing in the orthogonal space.
-As a consequence the "slices" are now spherical in the orthogonal subspace, with radius $h$.
+When going beyond 3d, the dimension of the orthogonal space is $>1$.
+Using the euclidean distance results in rotation invariant slicing in the orthogonal space, where a "slice" is spherical in the orthogonal subspace, and has radius $h$.
+
+Figure \ref{fig:renderFunction} shows slices through 4d geometric shapes, illustrating how we can distinguish the full from the hollow shapes.
 
@@
 
-Choosing the slice thickness is a compromise between what feature size can be revealed and how many points fall inside a thin slice.
+Choosing the slice thickness is a compromise between what feature size can be resolved and how many points fall inside a thin slice.
 As dimensionality $p$ increases, the relative number of points that are inside a slice of fixed thickness $h$ will decrease.
@@
 In three dimensions this simply corresponds to moving up or down along the orthogonal direction.
+Note that selecting an off-center anchor point will generally result in fewer points inside most slices.
 
@@
 In addition to usual parameters the user can choose the volume parameter $\epsilon$ by setting `eps` and, if required, the anchor point $\alpha$ is set by the `anchor` argument.
-By default `eps=0.001` and `anchor=NULL`, resulting in slicing through the origin.
-In addition the user can select the marker symbols for point inside and outside the slice.
+By default `eps=0.001` and `anchor=NULL`, resulting in slicing through the mean In addition the user can select the marker symbols for point inside and outside the slice.
 By default points in the slice are highlighted as `pch_slice = 20` and `pch_other = 46`, i.e.
 plotting a bullet for points inside the slice, and a dot for points outside.
+Below we show example code for displaying slices through a hollow 3d sphere.
 
@@
 anchor3 <- rep(0.7, 3)
-animate_slice(sphere3, anchor = anchor3, eps = 0.005, pch_other=26)
+animate_slice(sphere3, anchor = anchor3, eps = 0.2, pch_other=26)
 ```
@@
 Using the `geozoo` package [@geozoo], we draw 2000 (5000) points that are uniformly distributed on the surface of a 3d (5d) sphere.
-For both examples we start by slicing through the origin with the default parameters, in particular $\epsilon = 0.001$, i.e.
-$h = 0.1$ in 3d and $h = 0.25$ in 5d.
+For both examples we start by slicing through the origin with the default parameters, in particular $\epsilon = 0.1$, i.e.
+$h = 0.1$ in 3d and $h = 0.46$ in 5d.
 Slicing through the origin results in sections where the points inside the slice are approximately on a circle with the full radius, and each view is similar to that shown in Figure~\ref{fig:renderFunction}.
-The animated gifs are available at XXX.
+Animated gifs of these examples are available at XXX.
 
@@
 The second column shows slices through a 4d torus, revealing different aspects of the shape.
-The last column shows a 6d cube, where the upper plot shows a view along two of the parameter combinations, allowing to clearly identify the rectangular shape.
+The last column shows a 6d cube, where the upper plot shows a view along two of the original parameters, allowing to clearly identify the rectangular shape.
 The panel below shows that this is not typically the case when looking at a randomly selected slice.
@@
 We use the pollen data as an example for a hidden feature generally occluded in projections.
-This 1986 Exposition's dataset was made by David Coleman of RCA Labs.
-The artificially generated data is comprised of 3848 observations on 5 variables.
-We center and scale each variable of the dataset and run a slice tour with a thin slice ($\epsilon = 0.0005$).
+This 5d 1986 Exposition's dataset was made by David Coleman of RCA Labs.
+[XXX REF] We center and scale each variable of the dataset and run a slice tour with a thin slice ($\epsilon = 0.0005$).
 Selected views are shown in the first two plots in Figure \ref{fig:pollen}, indicating the presence of an interesting feature hidden in the center.
@@
 The slices allow to clearly see the non-linearity of the boundary between the green and orange group.
-The first view also shows a non-linear boundary of the purple group, which interestingly does not appear at all in the second slice shown here.
+The first view also shows a non-linear boundary of the purple group.
+Interestingly this group does not appear at all in the second slice shown here.
 
@@
 
-In this note we introduce a new visualization method for systematic slicing of high-dimensional parameter spaces, based on interpolated projections obtained in a grand tour.
-The application to the hollow sphere demonstrates the potential of this new display to find and explore concave structures in distributions.
+In this note we introduce a new visualization method for dynamic slicing of high-dimensional parameter spaces.
+It is based on interpolated projections obtained in a grand tour and generates an interpolated sequence of projected sections.
+The examples shown in Section XXX demonstrate the potential of this new display to find and explore concave structures, as well as other hidden features.
 
 In applications with large number of dimensions and where only a few samples are available the method is limited.
-In this case thick slices may be needed to get a sufficient number of observations inside the slice, and as a result  small scale concavities cannot be resolved in the slice tour.
+In this case thick slices may be needed to get a sufficient number of observations inside the slice, and as a result  small scale structures cannot be resolved in the slice tour.
 If on the other hand a large number of samples are available one may consider displaying only points inside the slice to keep a clear display.

### 2019-09-05 paper.Rmd
added section labels and links

@@
 Data is commonly high-dimensional, and visualisation often relies on some form of dimension reduction.
-This can be taking linear projections, or nonlinear if one considers techniques like multidimensional scaling (MDS) or t-Distributed Stochastic Neighbor Embedding (t-SNE).
+This can be by taking linear projections, or nonlinear, if one considers techniques like multidimensional scaling (MDS) or t-Distributed Stochastic Neighbor Embedding (t-SNE).
 For the purposes here, the focus is on linear projections, in particular as provided by the grand tour [@As85; @BCAH05].
@@
 A major limitation of projections is their opacity.
-It is mitigated some, when scatterplots are used to render the projected data, through a pseudo-transparancy of sparseness of points.
-This is improved further if points are also drawn using some transparency.
+It is mitigated some, when scatterplots are used to render the projected data, through a pseudo-transparency of sparseness of points.
+This is improved further if points are also drawn using an alpha level that provides transparent dots on most display devices.
 Some features of multivariate distributions may be visible, but a lot is easily hidden, especially in the case of concave structures.
@@
 Similarly we can think of small scale structures hidden in the center of a multivariate distribution, that might be considered to be "needles in a haystack", which can be difficult to detect in projections.
-Another example is where one would like to examine non-linear boundaries between groups from a classification model, with the nature of the boundaries being occluded in the projections.
+Projections also obscure non-linear boundaries as might be constructed from a classification model, or nonlinear model fits in high-dimensions.
 
@@
 Slicing, or sectioning, is a way that the internal distribution of high-dimensional data can be explored.
-@prosection discusses a technique for combining projections and sections, by slicing in the back-dimensions.
+@prosection discusses a technique for combining projections, with sections constructed by slicing in the dimensions orthogonal to the projection.
 It is also possible to think of linked brushing (e.g.
 @ggobiCSDA2003, @JSSv081i05) as slicing.
-In this note we discuss another approach to sectioning, where observations in the space orthogonal to the projection are highlighted if they are close to the projection plane, and faded if further afield.
+In this note, we discuss an approach to sectioning, where observations in the space orthogonal to the projection are highlighted if they are close to a projection plane through the mean of the data, and faded if further afield.
 This is combined with the grand tour to provide a new dynamic display that can be used to systematically search for features hidden in high dimensions.
@@
 <!-- paper structure --> 
-The method is described in Section XXX, and it is implemented in the `tourr` [@tourr] package in R [@rref].
-Section XXX illustrates the use on several high-dimensional geometric shapes and established data sets, showing how concave or occluded structures can be visualised and explored with a slice tour.
-Future work is discussed in Section XXX.
+The new section tour method is described in Section \ref{sec:method}, and it is implemented in the `tourr` [@tourr] package in R [@rref] (Section \ref{sec:implementation}).
+Section \ref{sec:examples} illustrates the use on several high-dimensional geometric shapes and established data sets, showing how concave or occluded structures can be visualised and explored with a slice tour.
+Future work is discussed in Section \ref{sec:discussion}.
 
-# Method
+# Method {#sec:method}
 
@@
 
-# Implementation
+# Implementation {#sec:implementation}
 
@@
 
-# Examples
+# Examples {#sec:examples}
 ## Geometric shapes
@@
 
-# Summary and Outlook
+# Discussion {#sec:discussion}
 

### 2019-09-05 paper.Rmd
working through tour section

@@
 
-The basis underlying our approach is the grand tour.
-We first briefly review the relevant concepts.
-Building on these ideas we then introduce a method for displaying slices of parameter space that are defined for each projection matrix.
+The sectioning is combined with the projections provided by a grand tour. A brief explanation of the tour is provided followed by an explanation of how slicing is done.
 
@@
 
-A tour is providing a continuous animation, showing a sequence of $d$-dimensional (typically $d=1$ or $2$) projections from $p$-dimensional parameter space.
-The tour combines basis selection with basis interpolation.
-In a grand tour the basis selection is random (each new basis is chosen from all possible projections), and typically geodesic interpolation is used to generate smooth transitions between viewing planes.
+A tour provides a continuous sequence of $d$-dimensional (typically $d=1$ or $2$) projections from $p$-dimensional Euclidean space.
+It is constructed by combining a method for basis selection with geodesic interpolation between pairs of bases.
+In a grand tour, the basis selection is random, each new basis is chosen from all possible projections.
+In a guided tour, the bases are chosen based on an index of interestingness.
 Different basis selection methods as well as the geodesic interpolation are implemented in the `tourr` package [@tourr], which also provides several display functions for viewing the tour.
 
-The notation used in the following for the description of each projection is denoting $\mathbf{X}$ the $n\times p$ dimensional data matrix ($n$ observations in $p$ dimensions) and $\mathbf{A}$ an orthonormal $p \times d$ projection matrix.
+For the explanation of the slice tour, the actual mechanics of the tour are not important, and only the notion of a projection plane in high dimensions is needed.
+The notation used for explanations starts with deoting each projection as $\mathbf{X}$ the $n\times p$ dimensional data matrix ($n$ observations in $p$ dimensions) and $\mathbf{A}$ an orthonormal $p \times d$ projection matrix.
 The $d$-dimensional projection of the data is thus given by $\mathbf{Y} = \mathbf{X}\cdot\mathbf{A}$, giving the $n\times d$ dimensional projected data matrix to be plotted in each frame of the tour display.

### 2019-09-12 paper.Rmd
done with 2.2.1

@@
 
-The sectioning is combined with the projections provided by a grand tour. A brief explanation of the tour is provided followed by an explanation of how slicing is done.
-
 ## Tour review
@@
 The notation used for explanations starts with deoting each projection as $\mathbf{X}$ the $n\times p$ dimensional data matrix ($n$ observations in $p$ dimensions) and $\mathbf{A}$ an orthonormal $p \times d$ projection matrix.
-The $d$-dimensional projection of the data is thus given by $\mathbf{Y} = \mathbf{X}\cdot\mathbf{A}$, giving the $n\times d$ dimensional projected data matrix to be plotted in each frame of the tour display.
+The $d$-dimensional projection of the data is thus given by $\mathbf{Y} = \mathbf{X}\cdot\mathbf{A}$, producing the $n\times d$ dimensional projected data matrix to be plotted in each frame of the tour display.
 
-In the following we are interested in the orthogonal distance of scatter points from an anchored projection plane, thus $d=2$.
+To make a section, we are interested in the orthogonal distance of the data points from an anchored projection plane, particularly using $d=2$, but the approach theoretically works for any $d$.
 
@@
 
-In addition to the projected data $\mathbf{Y}$ we calculate for each projection the orthogonal distance of the data points from the current projection plane centered at the origin.
+In addition to the projected data $\mathbf{Y}$, for each projection the orthogonal distance of the data points from the current projection plane centered at the data mean, is calculated.
 The orthogonal component is given by
 \begin{equation}
-x_i' = x_i - (x_i\cdot a_1) a_1 - (x_i\cdot a_2) a_2
+\mathbf{x}_i' = \mathbf{x}_i - (\mathbf{x}_i\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{x}_i\cdot \mathbf{a}_2) \mathbf{a}_2
 \end{equation}
-with $x_i, i=1,...,n$ a $p$-dimensional observation in $\mathbf{X}$ and $a_i, i=1,2$ denoting the columns of the projection matrix, $\mathbf{A}=(a_1, a_2)$.
-Calculating the euclidean norm of $x'$ is straightforward thanks to the orthonormality of $\mathbf{A}$, and gives the components of the required distance vector.
+with $\mathbf{x}_i, i=1,...,n$ a $p$-dimensional observation in $\mathbf{X}$ and $\mathbf{a}_i, i=1,2$ denoting the columns of the projection matrix, $\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)$.
+Calculating the euclidean norm of $\mathbf{x}'$ is straightforward thanks to the orthonormality of $\mathbf{A}$, and gives the components of the required distance vector.
 \begin{equation}
-\tilde{v}_i^2 = ||x_i'||^2 = x_i^2 - (x_i\cdot a_1)^2 - (x_i\cdot a_2)^2.
+\tilde{v}_i^2 = ||\mathbf{x}_i'||^2.
 \end{equation}
-We also see that this is trivially generalizing for orthonormal projection matrices with $d>2$.
+This can be generalized for $d>2$.
 
 The distance vector can then be used to display a slice tour by highlighting points for which the orthogonal distance is smaller than a selected cutoff value $h$.
-Notice that in 3d, where for each plane we have a single orthogonal direction, this defines a flat slice of hight $2h$, matching the common conception of a slice.
-When going beyond 3d, the dimension of the orthogonal space is $>1$.
+In 3D, where for each plane there is only a single orthogonal direction, this defines a flat slice of height $2h$, matching the common conception of a slice.
+When going beyond 3D, the dimension of the orthogonal space is $>1$.
 Using the euclidean distance results in rotation invariant slicing in the orthogonal space, where a "slice" is spherical in the orthogonal subspace, and has radius $h$.
 
-Figure \ref{fig:renderFunction} shows slices through 4d geometric shapes, illustrating how we can distinguish the full from the hollow shapes.
+Figure \ref{fig:renderFunction} shows slices through 4D solid and hollow geometric shapes, illustrating the ease of distinguish the solid from the hollow shapes once slices are shown.
+For each shape, the points are generated either uniformly within the shape, or on the surface.
 

### 2019-09-12 paper.Rmd
done with 2.2.1

@@
 
-In addition to the projected data $\mathbf{Y}$, for each projection the orthogonal distance of the data points from the current projection plane centered at the data mean, is calculated.
-The orthogonal component is given by
+The orthogonal distance of the data points from the current projection plane centered at the data mean, is calculated as,
+
+\begin{equation}
+\tilde{v}_i^2 = ||\mathbf{x}_i'||^2.
+\end{equation}
+
+\noindent where
+
 \begin{equation}
@@
 \end{equation}
-with $\mathbf{x}_i, i=1,...,n$ a $p$-dimensional observation in $\mathbf{X}$ and $\mathbf{a}_i, i=1,2$ denoting the columns of the projection matrix, $\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)$.
+
+\noindent and $\mathbf{x}_i, i=1,...,n$ a $p$-dimensional observation in $\mathbf{X}$ and $\mathbf{a}_i, i=1,2$ denoting the columns of the projection matrix, $\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)$.
 Calculating the euclidean norm of $\mathbf{x}'$ is straightforward thanks to the orthonormality of $\mathbf{A}$, and gives the components of the required distance vector.
-\begin{equation}
-\tilde{v}_i^2 = ||\mathbf{x}_i'||^2.
-\end{equation}
 This can be generalized for $d>2$.
@@
 The distance vector can then be used to display a slice tour by highlighting points for which the orthogonal distance is smaller than a selected cutoff value $h$.
-In 3D, where for each plane there is only a single orthogonal direction, this defines a flat slice of height $2h$, matching the common conception of a slice.
-When going beyond 3D, the dimension of the orthogonal space is $>1$.
+In 3D, where for each plane there is only a single orthogonal direction, this defines a flat slice of height $2h$.
+We could use this single direction to slice systematically from one side of the space to the other.
+It is also possible to think of front and back, some points are closer to the viewer, and some are far, and depth cues could be used.
+When going beyond 3D, the dimension of the orthogonal space is $>1$, and all sense of direction is lost.
+In general, there are $(p-d)$ dimensions orthogonal to the projection plane, so using distance from the points to the plane is the simplest approach to generating slices.
 Using the euclidean distance results in rotation invariant slicing in the orthogonal space, where a "slice" is spherical in the orthogonal subspace, and has radius $h$.
 
-Figure \ref{fig:renderFunction} shows slices through 4D solid and hollow geometric shapes, illustrating the ease of distinguish the solid from the hollow shapes once slices are shown.
+Figure \ref{fig:renderFunction} shows slices through 4D solid and hollow geometric shapes.
 For each shape, the points are generated either uniformly within the shape, or on the surface.
+This illustrates the ease of distinguishing the solid from the hollow, once slices are made.
 

### 2019-09-12 paper.Rmd
added german to author list

@@
     email: dicook@monash.edu
+  - name: German Valencia
+    affil: a
+    email: german.valencia@monash.edu
 affiliation:

### 2019-09-14 paper.Rmd
r, R and epsilon need to be defined

@@
 
-\noindent and $\mathbf{x}_i, i=1,...,n$ a $p$-dimensional observation in $\mathbf{X}$ and $\mathbf{a}_i, i=1,2$ denoting the columns of the projection matrix, $\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)$.
+\noindent and $\mathbf{x}_i, i=1,...,n$ is a $p$-dimensional observation in $\mathbf{X}$ and $\mathbf{a}_i, i=1,2$ denoting the columns of the projection matrix, $\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)$.
 Calculating the euclidean norm of $\mathbf{x}'$ is straightforward thanks to the orthonormality of $\mathbf{A}$, and gives the components of the required distance vector.
@@
 
-Choosing the slice thickness is a compromise between what feature size can be resolved and how many points fall inside a thin slice.
-As dimensionality $p$ increases, the relative number of points that are inside a slice of fixed thickness $h$ will decrease.
+Choosing the slice thickness is a compromise between what feature size can be resolved and the sparseness of the data.
+As $p$ increases, the relative number of points that are inside a slice of fixed thickness $h$ will decrease.
 The exact relation depends on the distribution of the data points.
-Here we estimate how it scales with $h$ and $p$ for the example of points uniformly distributed in a hypersphere.
-The example was selected since it represents a rotation invariant uniform distribution in $p$ space.
-As we are interested in hollowness this type of distribution is more relevant than a multivariate normal distribution.
+To get a best estimate for $h$, given $p$, assume that the points are uniformly distributed in a hypersphere.
+This is a rotation invariant uniform distribution in $p$ space, and because with sllicing we are mostly interested in hollowness this is more relevant than assuming a multivariate normal distribution.
 
@@
 \end{equation}
-where the approximation is valid when $h \ll r$, as is typically expected to be the case.
-Thus if we want to keep the relative number of points (i.e.
-$V_{rel}$) approximately constant for slices in different dimensions $p$, we calculate $h$ from a volume parameter as $h = \epsilon^{1/(p-2)}$.
+where $r$ is XXX, and $R$ is XXX.
+The approximation is valid when $h \ll r$, as is typically expected to be the case.
+To keep the relative number of points (i.e.
+$V_{rel}$) approximately constant for slices in different dimensions $p$, we calculate $h$ from a volume parameter as $h = \epsilon^{1/(p-2)}$, where $epsilon$ is XXX.
 

### 2019-09-14 paper.Rmd
anchoring revised

@@
 To keep the relative number of points (i.e.
-$V_{rel}$) approximately constant for slices in different dimensions $p$, we calculate $h$ from a volume parameter as $h = \epsilon^{1/(p-2)}$, where $epsilon$ is XXX.
+$V_{rel}$) approximately constant for slices in different dimensions $p$, we calculate $h$ from a volume parameter as $h = \epsilon^{1/(p-2)}$, where $epsilon$ is a pre-chosen value indicating a fraction of the overall volume to slice, say 0.001.
 
@@
 
-The above definition implicitly assumes that we slice the distribution using the origin as a reference point.
-This is generally a good starting point assuming the data is centered before applying the slice tour.
-Alternative reference points can however provide additional insight into the concave structure of the distribution.
-We can also use the mean in each parameter as an anchor point if the data is not centered.
+The equations above assume that the projection plane, and thus the the slice, passes through 0, which will be the mean of centered data.
+This (anchor point) is generally a good option for the slice tour, because as the tour progresses the projection plane changes and can catch non-central concavities, too.
+However it is straightforward generalize the equations to use any anchor point, $\alpha$.
 
-We therefore generalize the distance vector to measure the anchored orthogonal distance, i.e. distance with respect to the anchor $\alpha$ used as a reference point.
-
-In general $\alpha$ can be any point in the $n$-dimensional parameter space, but we are only interested in the orthogonal component, thus we use $\alpha' = \alpha - (\alpha\cdot a_1) a_1 - (\alpha\cdot a_2 )a_2$ to define the generalized measure of orthogonal distance as
+In general $\alpha$ can be any point in the $n$-dimensional parameter space, but we are only interested in the orthogonal component, the part of the vector extedning out of the projection plane, $\mathbf{\alpha}' = \mathbf{\alpha} - (\mathbf{\alpha}\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{\alpha}\cdot \mathbf{a}_2 )\mathbf{a}_2$.
+The generalized measure of orthogonal distance is then
 \begin{equation}
-v_i^2 = ||x_i' - \alpha'||^2 = x_i'^2 + \alpha'^2 - 2 x_i'\cdot\alpha'
+v_i^2 = ||\mathbf{x}_i' - \mathbf{\alpha}'||^2 = \mathbf{x}_i'^2 + \mathbf{\alpha}'^2 - 2 \mathbf{x}_i'\cdot\mathbf{\alpha}'
 \end{equation}
@@
 \begin{equation}
-x_i'\cdot\alpha' = x_i\cdot\alpha - (\alpha\cdot a_1) (x_i\cdot a_1) - (\alpha\cdot a_2) (x_i\cdot a_2).
+\mathbf{x}_i'\cdot\mathbf{\alpha}' = \mathbf{x}_i\cdot\mathbf{\alpha} - (\mathbf{\alpha}\cdot \mathbf{a}_1) (\mathbf{x}_i\cdot \mathbf{a}_1) - (\mathbf{\alpha}\cdot \mathbf{a}_2) (\mathbf{x}_i\cdot \mathbf{a}_2).
 \end{equation}
 
-Using the generalized distance measure with a cutoff volume $\epsilon$ then corresponds to moving a slice of fixed thickness to be centered around the anchor point $\alpha$ in the space orthogonal to the current projection plane.
-In three dimensions this simply corresponds to moving up or down along the orthogonal direction.
-Note that selecting an off-center anchor point will generally result in fewer points inside most slices.
+Using the generalized distance measure with a cutoff volume, $\epsilon$, then corresponds to moving a slice of fixed thickness, corresponding to a neighborhood of the projection plane through the anchor point $\mathbf{\alpha}$.
+In 3D, this simply corresponds to moving up or down along the orthogonal direction.
+Note that selecting an off-center anchor point may result in fewer points inside a slice for most data.
 

### 2019-09-14 paper.Rmd
maybe need some diagrams
worked through example sections

@@
   space of projections obtained when running a tour, thus presenting the viewer
-  with an interpolated sequence of projected sections. The method has been
+  with an interpolated sequence of sliced projections. The method has been
   implemented in R as an extension to the tourr package, and can be used to
@@
 
-```{r renderFunction, echo=FALSE, eval=TRUE, warning=FALSE, fig.height=1.5, fig.width = 5.5, fig.cap="Projected slices through 4d geometric shapes.
+```{r renderFunction, echo=FALSE, eval=TRUE, warning=FALSE, fig.height=1.5, fig.width = 5.5, fig.cap="Sliced projections through 4D geometric shapes.
 On the left a full (hollow) hypersphere (a, b), on the right a full (hollow) hypercube (c, d).
@@
 
+*XXX Do we need a diagram?*
+
 ### Anchoring
@@
 
+*XXX Do we need a diagram?*
+
 # Implementation {#sec:implementation}
@@
 plotting a bullet for points inside the slice, and a dot for points outside.
-Below we show example code for displaying slices through a hollow 3d sphere.
+Below we show example code for displaying slices through a hollow 3D sphere.
 
@@
 library(tourr)
-# use geozoo to generate points on a hollow 3d sphere
+# use geozoo to generate points on a hollow 3D sphere
 sphere3 <- geozoo::sphere.hollow(3)$points
@@
 ## Geometric shapes
+
+Using the `geozoo` package [@geozoo] a number of ideal shapes are generated.
+
 ### Hollow sphere
 We sample points on the surface of a sphere with radius $r=1$ in $p=3$ and $p=5$ dimensions.
-Using the `geozoo` package [@geozoo], we draw 2000 (5000) points that are uniformly distributed on the surface of a 3d (5d) sphere.
+Using the `sphere.hollow()` function, we generate two sample data sets by generating 2000 and 5000 points from a uniform distribution on the surface of a 3D and 5D spheres.
 For both examples we start by slicing through the origin with the default parameters, in particular $\epsilon = 0.1$, i.e.
-$h = 0.1$ in 3d and $h = 0.46$ in 5d.
+$h = 0.1$ in 3D and $h = 0.46$ in 5D.
 Slicing through the origin results in sections where the points inside the slice are approximately on a circle with the full radius, and each view is similar to that shown in Figure~\ref{fig:renderFunction}.
@@
 In this case the different views obtained in the slice tour reveal more of the concave structure of the distribution.
-Figure \ref{fig:3anchored} shows example views from a slice tour on a 3d sphere with $r=1$ and anchor point $(0.7, 0.7, 0.7)$.
+Figure \ref{fig:3anchored} shows example views from a slice tour on a 3D sphere with $r=1$ and anchor point $(0.7, 0.7, 0.7)$.
 Notice that the anchor point was chosen to fall outside the sphere.
 Depending on the viewing angle selected by the tour the slice contains points on the circle with radius $r$ (when the anchor point has a negligible orthogonal component to the viewing plane); a circle with radius $<r$ as the viewing angle is tilted away from the axis connecting the anchor point to the origin; a full circle with small radius as the angle increases; and finally we see an empty slice when the projection plane is orthogonal to the anchor point axis.
-A similar picture is found for the 5d sphere.
+A similar picture is found for the 5D sphere.
 Note however that as dimensionality increases, the larger orthogonal space means that the anchor point will have a large orthogonal component in most views.
@@
 \includegraphics[width=0.2\textwidth]{gifs/png/sphere-3-anchored-041.png}}
-\caption{Different slices through a 3d hollow sphere with $r=1$, with anchor point $(0.7, 0.7, 0.7)$, showing full circle with small radius (left), circle with radius $<r$ (middle) and circle with radius $r$ (right).}
+\caption{Different slices through a 3D hollow sphere with $r=1$, with anchor point $(0.7, 0.7, 0.7)$, showing full circle with small radius (left), circle with radius $<r$ (middle) and circle with radius $r$ (right).}
 \label{fig:3anchored}
@@
 For each shape two selected views are shown.
-The first column shows views from the slice tour on a 3d Roman surface.
-The second column shows slices through a 4d torus, revealing different aspects of the shape.
+The first column shows views from the slice tour on a 3D Roman surface.
+The second column shows slices through a 4D torus, revealing different aspects of the shape.
 The last column shows a 6d cube, where the upper plot shows a view along two of the original parameters, allowing to clearly identify the rectangular shape.
@@
 }
-\caption{Different slices through a Roman surface (left), 4d torus (middle) and a 6d cube (right).}
+\caption{Different slices through a Roman surface (left), 4D torus (middle) and a 6d cube (right).}
 \label{fig:geoms}
@@
 We use the pollen data as an example for a hidden feature generally occluded in projections.
-This 5d 1986 Exposition's dataset was made by David Coleman of RCA Labs.
+This 5D 1986 Exposition's dataset was made by David Coleman of RCA Labs.
 [XXX REF] We center and scale each variable of the dataset and run a slice tour with a thin slice ($\epsilon = 0.0005$).
@@
 
+XXX Show projection for each of these as well as slice.
+
 \begin{figure*}[ht]
@@
 In this note we introduce a new visualization method for dynamic slicing of high-dimensional parameter spaces.
-It is based on interpolated projections obtained in a grand tour and generates an interpolated sequence of projected sections.
+It is based on interpolated projections obtained in a grand tour and generates an interpolated sequence of sliced projections.
 The examples shown in Section XXX demonstrate the potential of this new display to find and explore concave structures, as well as other hidden features.

### 2019-09-16 appendix.Rmd
Appendix added.

---
title: "Appendix"
type: Short Technical Note
author:
  - name: Ursula Laa
    affil: a, b
    email: ursula.laa@monash.edu
  - name: Dianne Cook
    affil: b
    email: dicook@monash.edu
  - name: German Valencia
    affil: a
    email: german.valencia@monash.edu
affiliation:
  - num: a
    address: |
      School of Physics and Astronomy, Monash University
  - num: b
    address: |
      Department of Econometrics and Business Statistics, Monash University
bibliography: biblio.bib
geometry: margin=2.5cm
abstract: |
  Taking projections of high-dimensional data is a common analytical and visualisation technique in statistics for working with high-dimensional problems. Sectioning, or slicing, through high dimensions is less common, but can be useful for visualising data with concavities, or non-linear structure. It is associated with conditional distributions in statistics, and also linked brushing between plots in interactive data visualisation. This short technical note describes a simple approach for  slicing in the orthogonal
  space of projections obtained when running a tour, thus presenting the viewer
  with an interpolated sequence of sliced projections. The method has been
  implemented in R as an extension to the tourr package, and can be used to
  explore for concave and non-linear structures in multivariate distributions.
keywords: |
  data visualisation; grand tour; sectioning
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
  \usepackage{setspace}
output: rticles::tf_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

\doublespacing
# Slice volume derivation

We calculate the relative volume of a slice through a $p$ dimensional hypersphere. The volume of a full hypersphere with radius $r$ is

\begin{equation}
V(r, p) = \frac{\pi^{p/2} r^p}{\frac{p}{2} \Gamma(p/2)}.
\end{equation}

It is parametrised by its radius $r$ and its variation with the radius is given as

\begin{equation}
\frac{dV(r, p)}{dr} = 2 \frac{\pi^{p/2} r^{p-1}}{\Gamma(p/2)}.
\end{equation}

Our slice definition is spherical in the $p-2$ dimensional orthogonal space, while capturing the full volume (i.e. area) inside the plane. To calculate the slice volume we therefore integrate $\frac{dV(r, p-2)}{dr}$ multiplied with the area in the plane, from the origin to the slice thickness $h$. Using the parametrisation with the radius $r$, the area in the plane is a circle with radius $\sqrt{R^2 - r^2}$. The slice volume is thus calculated as

\begin{equation}
V_{slice}(R, h, p) =
\int_0^h \frac{dV(r, p)}{dr} V(\sqrt{R^2 - r^2}, 2) dr =
\frac{\pi^{p/2}}{\Gamma(p/2)} \frac{h^{p-2}}{p} [pR^2 - (p-2)h^2].
\end{equation}

The relative volume of the slice is then simply given as

\begin{equation}
V_{rel}(R, h, p) = \frac{V_{slice}(R, h, p)}{V(R, p)} = \frac{h^{p-2}}{2R^p} [pR^2 - (p-2)h^2].
\end{equation}

### 2019-09-16 appendix.Rmd
Small fixed in the appendix.

@@
 
-We calculate the relative volume of a slice through a $p$ dimensional hypersphere. The volume of a full hypersphere with radius $r$ is
-
+We calculate the relative volume of a slice through a $p$ dimensional hypersphere of radius $R$. In general the volume of a full hypersphere with radius $r$ in $q$ dimensions is given as
 \begin{equation}
-V(r, p) = \frac{\pi^{p/2} r^p}{\frac{p}{2} \Gamma(p/2)}.
+V(r, q) = \frac{\pi^{q/2} r^q}{\frac{q}{2} \Gamma(q/2)},
+\end{equation}
+and its variation with the radius is
+\begin{equation}
+\frac{dV(r, q)}{dr} = 2 \frac{\pi^{q/2} r^{q-1}}{\Gamma(q/2)}.
 \end{equation}
 
-It is parametrised by its radius $r$ and its variation with the radius is given as
-
-\begin{equation}
-\frac{dV(r, p)}{dr} = 2 \frac{\pi^{p/2} r^{p-1}}{\Gamma(p/2)}.
-\end{equation}
-
-Our slice definition is spherical in the $p-2$ dimensional orthogonal space, while capturing the full volume (i.e.
+To calculate the slice volume we note that it is spherical in the $p-2$ dimensional orthogonal space on the projection plane, while capturing the full volume (i.e.
 area) inside the plane.
-To calculate the slice volume we therefore integrate $\frac{dV(r, p-2)}{dr}$ multiplied with the area in the plane, from the origin to the slice thickness $h$.
-Using the parametrisation with the radius $r$, the area in the plane is a circle with radius $\sqrt{R^2 - r^2}$.
+To calculate the slice volume we therefore integrate the product of $\frac{dV(r, p-2)}{dr}$ and the area in the plane parametrised by $r$, from the origin to the full slice thickness $h$.
+The area in the plane is a circle with radius $\sqrt{R^2 - r^2}$.
 The slice volume is thus calculated as
-
 \begin{equation}
 V_{slice}(R, h, p) =
-\int_0^h \frac{dV(r, p)}{dr} V(\sqrt{R^2 - r^2}, 2) dr =
+\int_0^h \frac{dV(r, p-2)}{dr} V(\sqrt{R^2 - r^2}, 2) dr =
 \frac{\pi^{p/2}}{\Gamma(p/2)} \frac{h^{p-2}}{p} [pR^2 - (p-2)h^2].
 \end{equation}
-
 The relative volume of the slice is then simply given as
-
 \begin{equation}

### 2019-09-16 paper.Rmd
Notation fixes.

@@
 \begin{equation}
-V_{rel} = \frac{1}{2} \frac{h^{p-2}}{r^p} (p r^2 - (p-2) h^2) \approx \frac{1}{2} (\frac{h}{R})^{p-2},
+V_{rel} = \frac{1}{2} \frac{h^{p-2}}{R^p} (p R^2 - (p-2) h^2) \approx \frac{1}{2} (\frac{h}{R})^{p-2},
 \end{equation}
-where $r$ is XXX, and $R$ is XXX.
-The approximation is valid when $h \ll r$, as is typically expected to be the case.
+where $R$ is the radius of the hypersphere.
+The approximation is valid when $h \ll R$, as is typically expected to be the case.
 To keep the relative number of points (i.e.
@@
 The equations above assume that the projection plane, and thus the the slice, passes through 0, which will be the mean of centered data.
-This (anchor point) is generally a good option for the slice tour, because as the tour progresses the projection plane changes and can catch non-central concavities, too.
-However it is straightforward generalize the equations to use any anchor point, $\alpha$.
+This (center point) is generally a good option for the slice tour, because as the tour progresses the projection plane changes and can catch non-central concavities, too.
+However it is straightforward generalize the equations to use any center point, $\mathbf{c}$.
 
-In general $\alpha$ can be any point in the $n$-dimensional parameter space, but we are only interested in the orthogonal component, the part of the vector extedning out of the projection plane, $\mathbf{\alpha}' = \mathbf{\alpha} - (\mathbf{\alpha}\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{\alpha}\cdot \mathbf{a}_2 )\mathbf{a}_2$.
+In general $\mathbf{c}$ can be any point in the $p$-dimensional parameter space, but we are only interested in the orthogonal component, the part of the vector extedning out of the projection plane, $\mathbf{c}' = \mathbf{c} - (\mathbf{c}\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{c}\cdot \mathbf{a}_2 )\mathbf{a}_2$.
 The generalized measure of orthogonal distance is then
 \begin{equation}
-v_i^2 = ||\mathbf{x}_i' - \mathbf{\alpha}'||^2 = \mathbf{x}_i'^2 + \mathbf{\alpha}'^2 - 2 \mathbf{x}_i'\cdot\mathbf{\alpha}'
+v_i^2 = ||\mathbf{x}_i' - \mathbf{c}'||^2 = \mathbf{x}_i'^2 + \mathbf{c}'^2 - 2 \mathbf{x}_i'\cdot\mathbf{c}'
 \end{equation}
@@
 \begin{equation}
-\mathbf{x}_i'\cdot\mathbf{\alpha}' = \mathbf{x}_i\cdot\mathbf{\alpha} - (\mathbf{\alpha}\cdot \mathbf{a}_1) (\mathbf{x}_i\cdot \mathbf{a}_1) - (\mathbf{\alpha}\cdot \mathbf{a}_2) (\mathbf{x}_i\cdot \mathbf{a}_2).
+\mathbf{x}_i'\cdot\mathbf{c}' = \mathbf{x}_i\cdot\mathbf{c} - (\mathbf{c}\cdot \mathbf{a}_1) (\mathbf{x}_i\cdot \mathbf{a}_1) - (\mathbf{c}\cdot \mathbf{a}_2) (\mathbf{x}_i\cdot \mathbf{a}_2).
 \end{equation}
 
-Using the generalized distance measure with a cutoff volume, $\epsilon$, then corresponds to moving a slice of fixed thickness, corresponding to a neighborhood of the projection plane through the anchor point $\mathbf{\alpha}$.
+Using the generalized distance measure with a cutoff volume, $\epsilon$, then corresponds to moving a slice of fixed thickness, corresponding to a neighborhood of the projection plane through the anchor point $\mathbf{c}$.
 In 3D, this simply corresponds to moving up or down along the orthogonal direction.
@@
 The slice tour has been implemented in R [@rref] as a new display method `display_slice` in the `tourr` package [@tourr].
-In addition to usual parameters the user can choose the volume parameter $\epsilon$ by setting `eps` and, if required, the anchor point $\alpha$ is set by the `anchor` argument.
+In addition to usual parameters the user can choose the volume parameter $\epsilon$ by setting `eps` and, if required, the center point $\mathbf{c}$ is set by the `anchor` argument.
 By default `eps=0.001` and `anchor=NULL`, resulting in slicing through the mean In addition the user can select the marker symbols for point inside and outside the slice.
@@
 ### Hollow sphere
-We sample points on the surface of a sphere with radius $r=1$ in $p=3$ and $p=5$ dimensions.
+We sample points on the surface of a sphere with radius $R=1$ in $p=3$ and $p=5$ dimensions.
 Using the `sphere.hollow()` function, we generate two sample data sets by generating 2000 and 5000 points from a uniform distribution on the surface of a 3D and 5D spheres.
@@
 $h = 0.1$ in 3D and $h = 0.46$ in 5D.
-Slicing through the origin results in sections where the points inside the slice are approximately on a circle with the full radius, and each view is similar to that shown in Figure~\ref{fig:renderFunction}.
+Slicing through the origin results in sections where the points inside the slice are approximately on a circle with the full radius, and each view is similar to that shown in Figure \ref{fig:renderFunction}.
 Animated gifs of these examples are available at XXX.
@@
 In this case the different views obtained in the slice tour reveal more of the concave structure of the distribution.
-Figure \ref{fig:3anchored} shows example views from a slice tour on a 3D sphere with $r=1$ and anchor point $(0.7, 0.7, 0.7)$.
+Figure \ref{fig:3anchored} shows example views from a slice tour on a 3D sphere with $R=1$ and anchor point $(0.7, 0.7, 0.7)$.
 Notice that the anchor point was chosen to fall outside the sphere.
-Depending on the viewing angle selected by the tour the slice contains points on the circle with radius $r$ (when the anchor point has a negligible orthogonal component to the viewing plane); a circle with radius $<r$ as the viewing angle is tilted away from the axis connecting the anchor point to the origin; a full circle with small radius as the angle increases; and finally we see an empty slice when the projection plane is orthogonal to the anchor point axis.
+Depending on the viewing angle selected by the tour the slice contains points on the circle with radius $R$ (when the anchor point has a negligible orthogonal component to the viewing plane); a circle with radius $<R$ as the viewing angle is tilted away from the axis connecting the anchor point to the origin; a full circle with small radius as the angle increases; and finally we see an empty slice when the projection plane is orthogonal to the anchor point axis.
 A similar picture is found for the 5D sphere.
@@
 \includegraphics[width=0.2\textwidth]{gifs/png/sphere-3-anchored-041.png}}
-\caption{Different slices through a 3D hollow sphere with $r=1$, with anchor point $(0.7, 0.7, 0.7)$, showing full circle with small radius (left), circle with radius $<r$ (middle) and circle with radius $r$ (right).}
+\caption{Different slices through a 3D hollow sphere with $R=1$, with anchor point $(0.7, 0.7, 0.7)$, showing full circle with small radius (left), circle with radius $<R$ (middle) and circle with radius $R$ (right).}
 \label{fig:3anchored}

### 2019-09-17 paper.Rmd
Adding additional plots.

@@
 To keep the relative number of points (i.e.
-$V_{rel}$) approximately constant for slices in different dimensions $p$, we calculate $h$ from a volume parameter as $h = \epsilon^{1/(p-2)}$, where $epsilon$ is a pre-chosen value indicating a fraction of the overall volume to slice, say 0.001.
+$V_{rel}$) approximately constant for slices in different dimensions $p$, we calculate $h$ from a volume parameter as $h = \epsilon^{1/(p-2)}$, where $\epsilon$ is a pre-chosen value indicating a fraction of the overall volume to slice, say 0.001.
 
@@
 In this case the different views obtained in the slice tour reveal more of the concave structure of the distribution.
-Figure \ref{fig:3anchored} shows example views from a slice tour on a 3D sphere with $R=1$ and anchor point $(0.7, 0.7, 0.7)$.
+The first row of Figure \ref{fig:3anchored} shows example views from a slice tour on a 3D sphere with $R=1$ and anchor point $(0.7, 0.7, 0.7)$.
 Notice that the anchor point was chosen to fall outside the sphere.
 Depending on the viewing angle selected by the tour the slice contains points on the circle with radius $R$ (when the anchor point has a negligible orthogonal component to the viewing plane); a circle with radius $<R$ as the viewing angle is tilted away from the axis connecting the anchor point to the origin; a full circle with small radius as the angle increases; and finally we see an empty slice when the projection plane is orthogonal to the anchor point axis.
-A similar picture is found for the 5D sphere.
-Note however that as dimensionality increases, the larger orthogonal space means that the anchor point will have a large orthogonal component in most views.
+A similar picture is found for the 5D sphere, see second row of Figure \ref{fig:3anchored}.
+For this example we generate 20k points on a hollow 5D sphere to resolve the features.
+Since $h$ increases with $p$ the resolution is reduced compared to the 3D example.
+Note also that as dimensionality increases, the larger orthogonal space means that the center point will have a large orthogonal component in most views.
 
@@
 \includegraphics[width=0.2\textwidth]{gifs/png/sphere-3-anchored-041.png}}
-\caption{Different slices through a 3D hollow sphere with $R=1$, with anchor point $(0.7, 0.7, 0.7)$, showing full circle with small radius (left), circle with radius $<R$ (middle) and circle with radius $R$ (right).}
+\centerline{\includegraphics[width=0.2\textwidth]{gifs/png/sphere-5-anchored-014.png}
+\hspace{5mm}
+\includegraphics[width=0.2\textwidth]{gifs/png/sphere-5-anchored-010.png}
+\hspace{5mm}
+\includegraphics[width=0.2\textwidth]{gifs/png/sphere-5-anchored-001.png}}
+\caption{Different slices through a 3D (first row) and 5D (second row) hollow sphere with $R=1$, with shifted center point, showing full circle with small radius (left), circle with radius $<R$ (middle) and circle with radius $R$ (right).}
 \label{fig:3anchored}
@@
 
-Selected views from a slice tour of the scaled and centered data, and with large slice volume $\epsilon = 0.15$ are shown in Figure \ref{fig:wine}.
+Selected views from a slice tour of the scaled and centered data, and with large slice volume $\epsilon = 0.15$ are shown in the first row of Figure \ref{fig:wine}.
 The slices allow to clearly see the non-linearity of the boundary between the green and orange group.
@@
 Interestingly this group does not appear at all in the second slice shown here.
-
-XXX Show projection for each of these as well as slice.
+For comparison we show projections onto the same planes in the second row of Figure \ref{fig:wine}, showing how the observed features can be hidden by overplotting.
 
@@
 \hspace{5mm}
-\includegraphics[width=0.2\textwidth]{gifs/png/wine-100.png}
-}
-\caption{Different slices through the wine dataset.}
+\includegraphics[width=0.2\textwidth]{gifs/png/wine-100.png}}
+\centerline{\includegraphics[width=0.2\textwidth]{gifs/png/wine-projected-029.png}
+\hspace{5mm}
+\includegraphics[width=0.2\textwidth]{gifs/png/wine-projected-042.png}
+\hspace{5mm}
+\includegraphics[width=0.2\textwidth]{gifs/png/wine-projected-049.png}
+\hspace{5mm}
+\includegraphics[width=0.2\textwidth]{gifs/png/wine-projected-100.png}}
+\caption{Different slices (first row) and projections (second row) of the wine dataset .}
 \label{fig:wine}

### 2019-09-17 render-examples.R

@@ -20,14 +20,18 @@ render(sphere5, grand_tour(), display_slice(), 'png', "gifs/png/sphere-5-centere
 
 # Animate with off-center anchoring
 anchor3 <- rep(0.7, 3)
-anchor5 <- rep(0.5, 5)
 
 set.seed(2019)
 render(sphere3, grand_tour(), display_slice(anchor = anchor3), 'png', "gifs/png/sphere-3-anchored-%03d.png", frames = max_f, rescale = FALSE)
 
-# Animate with thicker slice to capture more points in each view
+# render 5d sphere with large number of points and off-center anchor
+sphere5 <- geozoo::sphere.hollow(5, 20000)$points
+colnames(sphere5) <- c("x1", "x2", "x3", "x4", "x5")
+anchor5 <- c(1.2, rep(0, 4))
+max_f <- 100 # max frames for rendering the animation
 set.seed(2019)
 render(sphere5, grand_tour(), display_slice(anchor = anchor5), 'png', "gifs/png/sphere-5-anchored-%03d.png", frames = max_f, rescale = FALSE)
+
 
 torus4 <- geozoo::torus(p=4)$points %>%
   scale()
@@ -70,6 +74,9 @@ wine_poly_scaled <- scale(as.matrix(wine_poly[,1:5]))
 set.seed(201909)
 render(wine_poly_scaled, grand_tour(), display_slice(axes = "bottomleft", eps=0.15, col=col), 'png', "gifs/png/wine-%03d.png", frames = max_f, rescale = FALSE)
 
+# render as projections
+set.seed(201909)
+render(wine_poly_scaled, grand_tour(), display_xy(axes = "bottomleft", col=col), 'png', "gifs/png/wine-projected-%03d.png", frames = max_f, rescale = FALSE)
 
 
 ####### end rendered examples

### 2019-09-17 paper.Rmd
Addign diagrams.
pollen data ref

@@
 Data is commonly high-dimensional, and visualisation often relies on some form of dimension reduction.
-This can be by taking linear projections, or nonlinear, if one considers techniques like multidimensional scaling (MDS) or t-Distributed Stochastic Neighbor Embedding (t-SNE).
+This can be done by taking linear projections, or nonlinear if one considers techniques like multidimensional scaling (MDS) (XXXREF) or t-Distributed Stochastic Neighbor Embedding (t-SNE) (XXXREF).
 For the purposes here, the focus is on linear projections, in particular as provided by the grand tour [@As85; @BCAH05].
@@
 We use the pollen data as an example for a hidden feature generally occluded in projections.
-This 5D 1986 Exposition's dataset was made by David Coleman of RCA Labs.
-[XXX REF] We center and scale each variable of the dataset and run a slice tour with a thin slice ($\epsilon = 0.0005$).
-Selected views are shown in the first two plots in Figure \ref{fig:pollen}, indicating the presence of an interesting feature hidden in the center.
-We next zoom in on the center of the distribution and consider a thicker slice ($\epsilon = 0.005$) to uncover the hidden structure, shown in the last two plots of Figure \ref{fig:pollen}.
+This is a classic 5D data set, originally simulated by by David Coleman of RCA Labs, for the Joint Statistics Meetings 1986 Data Expo [@pollen].
+The standardised data is observed in a slice tour with a thin slice ($\epsilon = 0.0005$).
+Selected views are shown in the first two plots in Figure \ref{fig:pollen}.
+They indicate the presence of an interesting feature hidden in the center, which can be identified as the word "EUREKA" by zooming.
 
@@
 }
-\caption{Different slices through the pollen dataset. The first two plots have slice volume $\epsilon = 0.0005$, the last two plots zoom in on the center and have increased volume $\epsilon = 0.005$.}
+\caption{Different slices through the classic pollen dataset.
+The first two plots have slice volume $\epsilon = 0.0005$, the last two plots zoom in on the center and have increased volume $\epsilon = 0.005$.
+The hidden word can be seen in the zoomed slice.}
 \label{fig:pollen}

### 2019-09-17 paper.Rmd
caption to illustration

@@
 
-*XXX Do we need a diagram?*
+\begin{figure*}[ht]
+\includegraphics[width=0.4\textwidth]{diagrams/centered-slice.pdf}}
+\hspace{5mm}
+\centerline{\includegraphics[width=0.4\textwidth]{diagrams/orthogonal-distance.pdf}
+\caption{Illustrations of  slicing, through a 3D sphere (left), and deomnstrating the calculation of the orthogonal distance  (right).
+3D slicing, can be done by sliding the projection plane along the orthogonal direction: centered at the origin (green) and one off-center at $\mathbf{c}$ (blue).
+This intuition does not transfer to higher dimensions, and it is best to use orthogonal distance between point and projection plane for computing the slice.
+Because a projection plane has no specific location, for slicing we can prescribe this, as through the data center, or any other point, $\mathbf{c}$.}
+\label{fig:diagrams}
+\end{figure*}
 

### 2019-09-17 paper.Rmd
fixed error in illustration
placement of fig 1 and fig 2

@@
 Data is commonly high-dimensional, and visualisation often relies on some form of dimension reduction.
-This can be done by taking linear projections, or nonlinear if one considers techniques like multidimensional scaling (MDS) (XXXREF) or t-Distributed Stochastic Neighbor Embedding (t-SNE) (XXXREF).
+This can be done by taking linear projections, or nonlinear if one considers techniques like multidimensional scaling (MDS) [@mds] or t-Distributed Stochastic Neighbor Embedding (t-SNE) [@tsne].
 For the purposes here, the focus is on linear projections, in particular as provided by the grand tour [@As85; @BCAH05].
@@
 This illustrates the ease of distinguishing the solid from the hollow, once slices are made.
+Figure \ref{fig:diagrams} illustrates the slicing method for 3D and higher dimensions.
+
+### Slice thickness
+
+Choosing the slice thickness is a compromise between what feature size can be resolved and the sparseness of the data.
+As $p$ increases, the relative number of points that are inside a slice of fixed thickness $h$ will decrease.
+The exact relation depends on the distribution of the data points.
+To get a best estimate for $h$, given $p$, assume that the points are uniformly distributed in a hypersphere.
+This is a rotation invariant uniform distribution in $p$ space, and because with sllicing we are mostly interested in hollowness this is more relevant than assuming a multivariate normal distribution.
+
+The fraction of points inside a slice of thickness $h$ can be estimated as the relative volume of the slice compared to the volume of the full hypersphere,
+\begin{equation}
+V_{rel} = \frac{1}{2} \frac{h^{p-2}}{R^p} (p R^2 - (p-2) h^2) \approx \frac{1}{2} (\frac{h}{R})^{p-2},
+\end{equation}
+where $R$ is the radius of the hypersphere.
+The approximation is valid when $h \ll R$, as is typically expected to be the case.
+To keep the relative number of points (i.e.
+$V_{rel}$) approximately constant for slices in different dimensions $p$, we calculate $h$ from a volume parameter as $h = \epsilon^{1/(p-2)}$, where $\epsilon$ is a pre-chosen value indicating a fraction of the overall volume to slice, say 0.001.
+
+### Non-central slice
+
+The equations above assume that the projection plane, and thus the the slice, passes through 0, which will be the mean of centered data.
+This (center point) is generally a good option for the slice tour, because as the tour progresses the projection plane changes and can catch non-central concavities, too.
+However it is straightforward generalize the equations to use any center point, $\mathbf{c}$.
+
+In general $\mathbf{c}$ can be any point in the $p$-dimensional parameter space, but we are only interested in the orthogonal component, the part of the vector extedning out of the projection plane, $\mathbf{c}' = \mathbf{c} - (\mathbf{c}\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{c}\cdot \mathbf{a}_2 )\mathbf{a}_2$.
+The generalized measure of orthogonal distance is then
+\begin{equation}
+v_i^2 = ||\mathbf{x}_i' - \mathbf{c}'||^2 = \mathbf{x}_i'^2 + \mathbf{c}'^2 - 2 \mathbf{x}_i'\cdot\mathbf{c}'
+\end{equation}
+where the cross term can be expressed as
+\begin{equation}
+\mathbf{x}_i'\cdot\mathbf{c}' = \mathbf{x}_i\cdot\mathbf{c} - (\mathbf{c}\cdot \mathbf{a}_1) (\mathbf{x}_i\cdot \mathbf{a}_1) - (\mathbf{c}\cdot \mathbf{a}_2) (\mathbf{x}_i\cdot \mathbf{a}_2).
+\end{equation}
+
+Using the generalized distance measure with a cutoff volume, $\epsilon$, then corresponds to moving a slice of fixed thickness, corresponding to a neighborhood of the projection plane through the anchor point $\mathbf{c}$.
+In 3D, this simply corresponds to moving up or down along the orthogonal direction.
+Note that selecting an off-center anchor point may result in fewer points inside a slice for most data.
 
@@
 
-### Slice thickness
-
-Choosing the slice thickness is a compromise between what feature size can be resolved and the sparseness of the data.
-As $p$ increases, the relative number of points that are inside a slice of fixed thickness $h$ will decrease.
-The exact relation depends on the distribution of the data points.
-To get a best estimate for $h$, given $p$, assume that the points are uniformly distributed in a hypersphere.
-This is a rotation invariant uniform distribution in $p$ space, and because with sllicing we are mostly interested in hollowness this is more relevant than assuming a multivariate normal distribution.
-
-The fraction of points inside a slice of thickness $h$ can be estimated as the relative volume of the slice compared to the volume of the full hypersphere,
-\begin{equation}
-V_{rel} = \frac{1}{2} \frac{h^{p-2}}{R^p} (p R^2 - (p-2) h^2) \approx \frac{1}{2} (\frac{h}{R})^{p-2},
-\end{equation}
-where $R$ is the radius of the hypersphere.
-The approximation is valid when $h \ll R$, as is typically expected to be the case.
-To keep the relative number of points (i.e.
-$V_{rel}$) approximately constant for slices in different dimensions $p$, we calculate $h$ from a volume parameter as $h = \epsilon^{1/(p-2)}$, where $\epsilon$ is a pre-chosen value indicating a fraction of the overall volume to slice, say 0.001.
-
 \begin{figure*}[ht]
-\includegraphics[width=0.4\textwidth]{diagrams/centered-slice.pdf}}
+\centerline{\includegraphics[width=0.4\textwidth]{diagrams/centered-slice.pdf}
 \hspace{5mm}
-\centerline{\includegraphics[width=0.4\textwidth]{diagrams/orthogonal-distance.pdf}
+\includegraphics[width=0.4\textwidth]{diagrams/orthogonal-distance.pdf}}
 \caption{Illustrations of slicing, through a 3D sphere (left), and deomnstrating the calculation of the orthogonal distance (right).
@@
 \end{figure*}
-
-### Anchoring
-
-The equations above assume that the projection plane, and thus the the slice, passes through 0, which will be the mean of centered data.
-This (center point) is generally a good option for the slice tour, because as the tour progresses the projection plane changes and can catch non-central concavities, too.
-However it is straightforward generalize the equations to use any center point, $\mathbf{c}$.
-
-In general $\mathbf{c}$ can be any point in the $p$-dimensional parameter space, but we are only interested in the orthogonal component, the part of the vector extedning out of the projection plane, $\mathbf{c}' = \mathbf{c} - (\mathbf{c}\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{c}\cdot \mathbf{a}_2 )\mathbf{a}_2$.
-The generalized measure of orthogonal distance is then
-\begin{equation}
-v_i^2 = ||\mathbf{x}_i' - \mathbf{c}'||^2 = \mathbf{x}_i'^2 + \mathbf{c}'^2 - 2 \mathbf{x}_i'\cdot\mathbf{c}'
-\end{equation}
-where the cross term can be expressed as
-\begin{equation}
-\mathbf{x}_i'\cdot\mathbf{c}' = \mathbf{x}_i\cdot\mathbf{c} - (\mathbf{c}\cdot \mathbf{a}_1) (\mathbf{x}_i\cdot \mathbf{a}_1) - (\mathbf{c}\cdot \mathbf{a}_2) (\mathbf{x}_i\cdot \mathbf{a}_2).
-\end{equation}
-
-Using the generalized distance measure with a cutoff volume, $\epsilon$, then corresponds to moving a slice of fixed thickness, corresponding to a neighborhood of the projection plane through the anchor point $\mathbf{c}$.
-In 3D, this simply corresponds to moving up or down along the orthogonal direction.
-Note that selecting an off-center anchor point may result in fewer points inside a slice for most data.
-
-*XXX Do we need a diagram?*
 

### 2019-09-17 paper.Rmd
fixing equations

@@
 
-The orthogonal distance of the data points from the current projection plane centered at the data mean, is calculated as,
+The orthogonal distance of the data points from the current projection plane is calculated as,
 
@@
 
-\noindent and $\mathbf{x}_i, i=1,...,n$ is a $p$-dimensional observation in $\mathbf{X}$ and $\mathbf{a}_i, i=1,2$ denoting the columns of the projection matrix, $\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)$.
-Calculating the euclidean norm of $\mathbf{x}'$ is straightforward thanks to the orthonormality of $\mathbf{A}$, and gives the components of the required distance vector.
+\noindent and $\mathbf{x}_i, i=1,...,n$ is a $p$-dimensional observation in $\mathbf{X}$ and $\mathbf{a}_k, k=1,2 (=d)$ denoting the columns of the projection matrix, $\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)$.
+$\mathbf{A}$ is assummed to be through 0, because the data has been centered on its mean.
+$\mathbf{x}_i'$ can be considered to be a normal from the projection plane to $\mathbf{x}_i$, and then the norm of this vector gives the orthogonal distance between point and plane.
 This can be generalized for $d>2$.
 
-The distance vector can then be used to display a slice tour by highlighting points for which the orthogonal distance is smaller than a selected cutoff value $h$.
+The distance can then be used to display a slice tour by highlighting points for which the orthogonal distance is smaller than a selected cutoff value $h$.
 In 3D, where for each plane there is only a single orthogonal direction, this defines a flat slice of height $2h$.

### 2019-09-18 coverletter.Rmd
Coverletter added.

---
from: Ursula Laa
fromqualifications: PhD
fromtitle: Research fellow
fromemail: ursula.laa@monash.edu
fromWWW: github.com/uschiLaa
fromphone: +61 4 78 001 984
fromsignature: signature.pdf
toname: Editor
toaddress: Journal of Computational and Graphical Statistics
opening: Dear Dr McCormick,
closing: Sincerely
spacing: 1.4
output: MonashEBSTemplates::letter
---

Please consider our manuscript titled "A slice tour for finding hollowness in high-dimensional data" for publication as a short technical note in the Journal of Computational and Graphical Statistics.

This note describes a new method for the visualisation of high-dimensional data by defining slices in the orthogonal space of 2D projections. It can be combined with a tour for a dynamic display that shows slices on a path of interpolated projections. The method has been implemented in R as a new display option in the tourr package. It is particularly useful to explore for concave and non-linear structures in multivariate distributions. Several examples are included in the note and illustrate the potential for different types of applications.

We believe that it is an ideal fit for a short technical note in the Journal of Computational and Graphical Statistics. It describes a new graphical method that will be useful to people working in multivariate statistics.

Thank you for the consideration of this manuscript.

### 2019-09-21 paper.Rmd
done with paper edits

@@
 keywords: |
-  data visualisation; grand tour; sectioning
+  data visualisation; grand tour; sectioning; statistical computing; statistical graphics; high-dimensional data
 header-includes: |
@@
 \begin{figure*}[ht]
-\centerline{\includegraphics[width=0.2\textwidth]{gifs/png/sphere-3-anchored-033.png}
+\centerline{\includegraphics[width=0.2\textwidth]{gifs/png/sphere-3-anchored/sphere-3-anchored-033.png}
 \hspace{5mm}
-\includegraphics[width=0.2\textwidth]{gifs/png/sphere-3-anchored-036.png}
+\includegraphics[width=0.2\textwidth]{gifs/png/sphere-3-anchored/sphere-3-anchored-036.png}
 \hspace{5mm}
-\includegraphics[width=0.2\textwidth]{gifs/png/sphere-3-anchored-041.png}}
-\centerline{\includegraphics[width=0.2\textwidth]{gifs/png/sphere-5-anchored-014.png}
+\includegraphics[width=0.2\textwidth]{gifs/png/sphere-3-anchored/sphere-3-anchored-041.png}}
+\centerline{\includegraphics[width=0.2\textwidth]{gifs/png/sphere-5-anchored/sphere-5-anchored-014.png}
 \hspace{5mm}
-\includegraphics[width=0.2\textwidth]{gifs/png/sphere-5-anchored-010.png}
+\includegraphics[width=0.2\textwidth]{gifs/png/sphere-5-anchored/sphere-5-anchored-010.png}
 \hspace{5mm}
-\includegraphics[width=0.2\textwidth]{gifs/png/sphere-5-anchored-001.png}}
+\includegraphics[width=0.2\textwidth]{gifs/png/sphere-5-anchored/sphere-5-anchored-001.png}}
 \caption{Different slices through a 3D (first row) and 5D (second row) hollow sphere with $R=1$, with shifted center point, showing full circle with small radius (left), circle with radius $<R$ (middle) and circle with radius $R$ (right).}
@@
 \begin{figure*}[ht]
-\centerline{\includegraphics[width=0.2\textwidth]{gifs/png/roman-surface-centered-010.png}
+\centerline{\includegraphics[width=0.2\textwidth]{gifs/png/roman-surface/roman-surface-centered-010.png}
 \hspace{5mm}
-\includegraphics[width=0.2\textwidth]{gifs/png/torus-4-centered-015.png}
+\includegraphics[width=0.2\textwidth]{gifs/png/torus-4-centered/torus-4-centered-015.png}
 \hspace{5mm}
-\includegraphics[width=0.2\textwidth]{gifs/png/cube-6-centered-001.png}}
+\includegraphics[width=0.2\textwidth]{gifs/png/cube-6-centered/cube-6-centered-001.png}}
 \centerline{
-\includegraphics[width=0.2\textwidth]{gifs/png/roman-surface-centered-032.png}
+\includegraphics[width=0.2\textwidth]{gifs/png/roman-surface/roman-surface-centered-032.png}
 \hspace{5mm}
-\includegraphics[width=0.2\textwidth]{gifs/png/torus-4-centered-033.png}
+\includegraphics[width=0.2\textwidth]{gifs/png/torus-4-centered/torus-4-centered-033.png}
 \hspace{5mm}
-\includegraphics[width=0.2\textwidth]{gifs/png/cube-6-centered-063.png}
+\includegraphics[width=0.2\textwidth]{gifs/png/cube-6-centered/cube-6-centered-063.png}
 }
@@
 \begin{figure*}[ht]
-\centerline{\includegraphics[width=0.2\textwidth]{gifs/png/pollen-007.png}
+\centerline{\includegraphics[width=0.2\textwidth]{gifs/png/pollen/pollen-007.png}
 \hspace{5mm}
-\includegraphics[width=0.2\textwidth]{gifs/png/pollen-092.png}
+\includegraphics[width=0.2\textwidth]{gifs/png/pollen/pollen-092.png}
 \hspace{5mm}
-\includegraphics[width=0.2\textwidth]{gifs/png/pollen-zoom-075.png}
+\includegraphics[width=0.2\textwidth]{gifs/png/pollen-zoom/pollen-zoom-075.png}
 \hspace{5mm}
-\includegraphics[width=0.2\textwidth]{gifs/png/pollen-zoom-018.png}
+\includegraphics[width=0.2\textwidth]{gifs/png/pollen-zoom/pollen-zoom-018.png}
 }
@@
 ### Non-linear boundaries
-XXX say something about this dataset [@sam.11271]
+@sam.11271 described some principles and approaches for visualising models in the data space.
+Much of this is based on examining projections provided by a tour of the model in the high-dimensional space.
+Classification boundaries are explored for the wine data set [@asuncion:2007].
+It is difficult to digest the boundaries fully -- for example, where one group's boundary wraps another for example, if they ar linear or nonlinear, or whether the boundary goes through the space or is only carving out a corner of it.
+The sliced projections makes this easier.
 
-Selected views from a slice tour of the scaled and centered data, and with large slice volume $\epsilon = 0.15$ are shown in the first row of Figure \ref{fig:wine}.
-The slices allow to clearly see the non-linearity of the boundary between the green and orange group.
-The first view also shows a non-linear boundary of the purple group.
-Interestingly this group does not appear at all in the second slice shown here.
-For comparison we show projections onto the same planes in the second row of Figure \ref{fig:wine}, showing how the observed features can be hidden by overplotting.
+Selected views of projections and sliced projections from a radial basis SVM on 3 variables, and a polynomial basis SVM on 5 variables are shown in Figure \ref{fig:wine}.
+A large slice volume $\epsilon = 0.15$ is used.
+The slicing allows exploring the center of the space.
+With 3D it reveals the spherical boundary of the group (green) hidden by the projection.
+In 5D with the polynomial basis, projection might suggest that the boundary between classes is almost linear, but the slicing shows that it to be nonlinear near the center.
 
 \begin{figure*}[ht]
-\centerline{\includegraphics[width=0.2\textwidth]{gifs/png/wine-029.png}
+\centerline{
+\includegraphics[width=0.3\textwidth]{gifs/png/wine-radial-projected/wine-radial-projected-031.png}
 \hspace{5mm}
-\includegraphics[width=0.2\textwidth]{gifs/png/wine-042.png}
+\includegraphics[width=0.3\textwidth]{gifs/png/wine-poly-projected/wine-poly-projected-096.png}
+\includegraphics[width=0.3\textwidth]{gifs/png/wine-poly-projected/wine-poly-projected-042.png}}
+\centerline{
+\includegraphics[width=0.3\textwidth]{gifs/png/wine-radial/wine-radial-031.png}
 \hspace{5mm}
-\includegraphics[width=0.2\textwidth]{gifs/png/wine-049.png}
-\hspace{5mm}
-\includegraphics[width=0.2\textwidth]{gifs/png/wine-100.png}}
-\centerline{\includegraphics[width=0.2\textwidth]{gifs/png/wine-projected-029.png}
-\hspace{5mm}
-\includegraphics[width=0.2\textwidth]{gifs/png/wine-projected-042.png}
-\hspace{5mm}
-\includegraphics[width=0.2\textwidth]{gifs/png/wine-projected-049.png}
-\hspace{5mm}
-\includegraphics[width=0.2\textwidth]{gifs/png/wine-projected-100.png}}
-\caption{Different slices (first row) and projections (second row) of the wine dataset .}
+\includegraphics[width=0.3\textwidth]{gifs/png/wine-poly/wine-poly-096.png}
+\includegraphics[width=0.3\textwidth]{gifs/png/wine-poly/wine-poly-042.png}
+}
+\caption{Exploring classification boundaries of the wine dataset using projection (top row) and sliced projection (bottom row).
+Radial basis SVM (First column) of 3 variables shows how the slicing reveals the spherical shape of the boundary of one group (green) that was hidden in the projection.
+Polynomial basis SVM (second and third columns) of 5 variables.
+The orange group does tend to be wrapped by the green group, and the blue group disappears with slicing, showing that it is on the outer edge of the space.}
 \label{fig:wine}
@@
 
-In this note we introduce a new visualization method for dynamic slicing of high-dimensional parameter spaces.
-It is based on interpolated projections obtained in a grand tour and generates an interpolated sequence of sliced projections.
-The examples shown in Section XXX demonstrate the potential of this new display to find and explore concave structures, as well as other hidden features.
+This paper has introduced a new visualization method for dynamic slicing of high-dimensional spaces.
+It is based on interpolated projections obtained in a (grand) tour and generates an interpolated sequence of sliced projections.
+The examples shown in Section \ref{sec:examples} demonstrate the potential of this new display to find and explore concave structures, as well as other hidden features.
 
-In applications with large number of dimensions and where only a few samples are available the method is limited.
-In this case thick slices may be needed to get a sufficient number of observations inside the slice, and as a result  small scale structures cannot be resolved in the slice tour.
-If on the other hand a large number of samples are available one may consider displaying only points inside the slice to keep a clear display.
+A default slice thickness is provided with the algorithm, that takes dimensionality into account.
+As the data dimension increases, more points in the sample are needed, and a thicker slice may be needed.
+Generally, the tour can be slow to view when there are a large number of samples, and the slicing, where only points inside the slice are drawn, might also be a way to improve the display drawing, with a focus on the important features.
 
-Instead of slicing in the full $p-d=p-2$ dimensional orthogonal space we can also use a subspace in which to slice.
-As pointed out in @prosection combining projections with sectioning in varying numbers of dimensions can reveal the codimension of an unknown distribution.
-One approach would be to interpolate between projection matrices with more than two columns, and use the first two for the display projection.
-The remaining directions then provide an orthogonal subspace that can be used for sectioning.
-A subtlety to consider is that geodesic interpolation is not ideal for such applications, as it ignores rotations between the directions in the projection matrix.
-As alternative interpolation methods are not currently available in the R implementation, we leave such extensions for future work.
+Sliced 2D tours was available in XGobi [@SCB98] but was not documented. This work makes them available in the `tourr` package. It is a simple, but effective, approach to taking slices.
 
-In addition slice definitions beyond a simple cutoff on the orthogonal distance should be considered, where different definition can be used to highlight certain aspects of the data.
-We also expect interesting new developments by combining the slice tour and the guided tour into a weighted guided tour with index functions using the distance vector information in addition to the projected data points.
+The approach by @prosection is more complex, and slices a subspace  of the $p-d=p-2$ dimensional space orthogonal to the projection.
+This generates a more parameters, making it more difficult to navigate.
+More parameters mean more decisions on what to show.
+However, it is one of the next steps to explore different definitions of a slice tour, and the types of structure that might be captured by variations in slicing algorithms.
+
+Lastly, there is a large literature on projection pursuit, and some of this work is available in the projection pursuit guided tour in the `tourr` package.
+The projections shown are more interesting in a guided tour than might be seen using a grand tour.
+The slicing might be introduced into projection pursuit by defining weighted projection pursuit indexes.
+The resulting indexes could be incorporated into a guided slice tour, finding projections where the slice reveals something new.
+
+# Acknowledgements
+
+The paper was written in `rmarkdown` [@rmarkdown] using `knitr` [@knitr]. The source material and animated gifs for this paper are available at
+\url{https://github.com/uschiLaa/paper-slice-tour}.

### 2019-09-21 render-examples.R

@@ -66,17 +66,44 @@ set.seed(201909)
 render(pollen, grand_tour(), display_slice( half_range=1, eps=0.005, axes = "bottomleft"), 'png', "gifs/png/pollen-zoom-%03d.png", frames = max_f, rescale = FALSE)
 
 library(RColorBrewer)
-wine_radial <- read_csv("data/wine-svm-radial.csv")
 wine_poly <- read_csv("data/wine-svm-poly.csv")
 clrs <- brewer.pal(3, "Dark2")
 col <- clrs[as.numeric(as.factor(wine_poly$type))]
 wine_poly_scaled <- scale(as.matrix(wine_poly[,1:5]))
 set.seed(201909)
-render(wine_poly_scaled, grand_tour(), display_slice(axes = "bottomleft", eps=0.15, col=col), 'png', "gifs/png/wine-%03d.png", frames = max_f, rescale = FALSE)
+render(wine_poly_scaled, grand_tour(), 
+       display_slice(axes = "bottomleft", eps=0.15, col=col, cex=4), 
+       'png', "gifs/png/wine-poly/wine-poly-%03d.png", frames = max_f, rescale = FALSE)
 
 # render as projections
 set.seed(201909)
-render(wine_poly_scaled, grand_tour(), display_xy(axes = "bottomleft", col=col), 'png', "gifs/png/wine-projected-%03d.png", frames = max_f, rescale = FALSE)
+render(wine_poly_scaled, 
+       grand_tour(), display_xy(axes = "bottomleft", col=col), 
+       'png', "gifs/png/wine-poly-projected/wine-poly-projected-%03d.png", frames = max_f, rescale = FALSE)
+
+# Radial
+wine_radial <- read_csv("data/wine-svm-radial.csv")
+wine_radial <- wine_radial %>% 
+  group_by(type) %>%
+  sample_frac(0.5)
+
+col <- clrs[as.numeric(as.factor(wine_radial$type))]
+wine_radial_scaled <- scale(as.matrix(wine_radial[,1:3]))
+set.seed(201909)
+render(wine_radial_scaled, grand_tour(), 
+       display_slice(axes = "bottomleft", eps=0.05, col=col), 
+       'png', "gifs/png/wine-radial/wine-radial-%03d.png", frames = max_f, rescale = FALSE)
+set.seed(201909)
+render(wine_radial_scaled, 
+       grand_tour(), display_xy(axes = "bottomleft", col=col), 
[…]

### 2019-09-22 paper.Rmd
Spellcheck

@@
 Data is commonly high-dimensional, and visualisation often relies on some form of dimension reduction.
-This can be done by taking linear projections, or nonlinear if one considers techniques like multidimensional scaling (MDS) [@mds] or t-Distributed Stochastic Neighbor Embedding (t-SNE) [@tsne].
+This can be done by taking linear projections, or nonlinear if one considers techniques like multidimensional scaling (MDS) [@mds] or t-Distributed Stochastic Neighbour Embedding (t-SNE) [@tsne].
 For the purposes here, the focus is on linear projections, in particular as provided by the grand tour [@As85; @BCAH05].
@@
 Considering the example of a simple geometric shape such as a hypershere, it is difficult to distinguish between a full or a hollow sphere, that is, whether points are uniformly distributed within the sphere or on the surface of a sphere.
-Similarly we can think of small scale structures hidden in the center of a multivariate distribution, that might be considered to be "needles in a haystack", which can be difficult to detect in projections.
+Similarly we can think of small scale structures hidden in the centre of a multivariate distribution, that might be considered to be "needles in a haystack", which can be difficult to detect in projections.
 Projections also obscure non-linear boundaries as might be constructed from a classification model, or nonlinear model fits in high-dimensions.
@@
 For the explanation of the slice tour, the actual mechanics of the tour are not important, and only the notion of a projection plane in high dimensions is needed.
-The notation used for explanations starts with deoting each projection as $\mathbf{X}$ the $n\times p$ dimensional data matrix ($n$ observations in $p$ dimensions) and $\mathbf{A}$ an orthonormal $p \times d$ projection matrix.
+The notation used for explanations starts with denoting each projection as $\mathbf{X}$ the $n\times p$ dimensional data matrix ($n$ observations in $p$ dimensions) and $\mathbf{A}$ an orthonormal $p \times d$ projection matrix.
 The $d$-dimensional projection of the data is thus given by $\mathbf{Y} = \mathbf{X}\cdot\mathbf{A}$, producing the $n\times d$ dimensional projected data matrix to be plotted in each frame of the tour display.
@@
 \noindent and $\mathbf{x}_i, i=1,...,n$ is a $p$-dimensional observation in $\mathbf{X}$ and $\mathbf{a}_k, k=1,2 (=d)$ denoting the columns of the projection matrix, $\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)$.
-$\mathbf{A}$ is assummed to be through 0, because the data has been centered on its mean.
+$\mathbf{A}$ is assumed to be through 0, because the data has been centred on its mean.
 $\mathbf{x}_i'$ can be considered to be a normal from the projection plane to $\mathbf{x}_i$, and then the norm of this vector gives the orthogonal distance between point and plane.
@@
 In general, there are $(p-d)$ dimensions orthogonal to the projection plane, so using distance from the points to the plane is the simplest approach to generating slices.
-Using the euclidean distance results in rotation invariant slicing in the orthogonal space, where a "slice" is spherical in the orthogonal subspace, and has radius $h$.
+Using the Euclidean distance results in rotation invariant slicing in the orthogonal space, where a "slice" is spherical in the orthogonal subspace, and has radius $h$.
 
@@
 To get a best estimate for $h$, given $p$, assume that the points are uniformly distributed in a hypersphere.
-This is a rotation invariant uniform distribution in $p$ space, and because with sllicing we are mostly interested in hollowness this is more relevant than assuming a multivariate normal distribution.
+This is a rotation invariant uniform distribution in $p$ space, and because with slicing we are mostly interested in hollowness this is more relevant than assuming a multivariate normal distribution.
 
@@
 
-The equations above assume that the projection plane, and thus the the slice, passes through 0, which will be the mean of centered data.
-This (center point) is generally a good option for the slice tour, because as the tour progresses the projection plane changes and can catch non-central concavities, too.
-However it is straightforward generalize the equations to use any center point, $\mathbf{c}$.
+The equations above assume that the projection plane, and thus the the slice, passes through 0, which will be the mean of centred data.
+This (centre point) is generally a good option for the slice tour, because as the tour progresses the projection plane changes and can catch non-central concavities, too.
+However it is straightforward generalize the equations to use any centre point, $\mathbf{c}$.
 
-In general $\mathbf{c}$ can be any point in the $p$-dimensional parameter space, but we are only interested in the orthogonal component, the part of the vector extedning out of the projection plane, $\mathbf{c}' = \mathbf{c} - (\mathbf{c}\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{c}\cdot \mathbf{a}_2 )\mathbf{a}_2$.
+In general $\mathbf{c}$ can be any point in the $p$-dimensional parameter space, but we are only interested in the orthogonal component, the part of the vector extending out of the projection plane, $\mathbf{c}' = \mathbf{c} - (\mathbf{c}\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{c}\cdot \mathbf{a}_2 )\mathbf{a}_2$.
 The generalized measure of orthogonal distance is then
@@
 
-Using the generalized distance measure with a cutoff volume, $\epsilon$, then corresponds to moving a slice of fixed thickness, corresponding to a neighborhood of the projection plane through the anchor point $\mathbf{c}$.
+Using the generalized distance measure with a cutoff volume, $\epsilon$, then corresponds to moving a slice of fixed thickness, corresponding to a neighbourhood of the projection plane through the anchor point $\mathbf{c}$.
 In 3D, this simply corresponds to moving up or down along the orthogonal direction.
-Note that selecting an off-center anchor point may result in fewer points inside a slice for most data.
+Note that selecting an off-centre anchor point may result in fewer points inside a slice for most data.
 
@@
 \includegraphics[width=0.4\textwidth]{diagrams/orthogonal-distance.pdf}}
-\caption{Illustrations of slicing, through a 3D sphere (left), and deomnstrating the calculation of the orthogonal distance (right).
-3D slicing, can be done by sliding the projection plane along the orthogonal direction: centered at the origin (green) and one off-center at $\mathbf{c}$ (blue).
+\caption{Illustrations of slicing, through a 3D sphere (left), and demonstrating the calculation of the orthogonal distance (right).
+3D slicing, can be done by sliding the projection plane along the orthogonal direction: centred at the origin (green) and one off-centre at $\mathbf{c}$ (blue).
 This intuition does not transfer to higher dimensions, and it is best to use orthogonal distance between point and projection plane for computing the slice.
-Because a projection plane has no specific location, for slicing we can prescribe this, as through the data center, or any other point, $\mathbf{c}$.}
+Because a projection plane has no specific location, for slicing we can prescribe this, as through the data centre, or any other point, $\mathbf{c}$.}
 \label{fig:diagrams}
@@
 The slice tour has been implemented in R [@rref] as a new display method `display_slice` in the `tourr` package [@tourr].
-In addition to usual parameters the user can choose the volume parameter $\epsilon$ by setting `eps` and, if required, the center point $\mathbf{c}$ is set by the `anchor` argument.
+In addition to usual parameters the user can choose the volume parameter $\epsilon$ by setting `eps` and, if required, the centre point $\mathbf{c}$ is set by the `anchor` argument.
 By default `eps=0.001` and `anchor=NULL`, resulting in slicing through the mean In addition the user can select the marker symbols for point inside and outside the slice.
@@
 
-It is especially instructive to look at the sliced spheres with an off-center anchor point.
+It is especially instructive to look at the sliced spheres with an off-centre anchor point.
 In this case the different views obtained in the slice tour reveal more of the concave structure of the distribution.
@@
 Since $h$ increases with $p$ the resolution is reduced compared to the 3D example.
-Note also that as dimensionality increases, the larger orthogonal space means that the center point will have a large orthogonal component in most views.
+Note also that as dimensionality increases, the larger orthogonal space means that the centre point will have a large orthogonal component in most views.
 
@@
 \includegraphics[width=0.2\textwidth]{gifs/png/sphere-5-anchored/sphere-5-anchored-001.png}}
-\caption{Different slices through a 3D (first row) and 5D (second row) hollow sphere with $R=1$, with shifted center point, showing full circle with small radius (left), circle with radius $<R$ (middle) and circle with radius $R$ (right).}
+\caption{Different slices through a 3D (first row) and 5D (second row) hollow sphere with $R=1$, with shifted centre point, showing full circle with small radius (left), circle with radius $<R$ (middle) and circle with radius $R$ (right).}
 \label{fig:3anchored}
@@
 Selected views are shown in the first two plots in Figure \ref{fig:pollen}.
-They indicate the presence of an interesting feature hidden in the center, which can be identified as the word "EUREKA" by zooming.
+They indicate the presence of an interesting feature hidden in the centre, which can be identified as the word "EUREKA" by zooming.
 
@@
 \caption{Different slices through the classic pollen dataset.
-The first two plots have slice volume $\epsilon = 0.0005$, the last two plots zoom in on the center and have increased volume $\epsilon = 0.005$.
+The first two plots have slice volume $\epsilon = 0.0005$, the last two plots zoom in on the centre and have increased volume $\epsilon = 0.005$.
 The hidden word can be seen in the zoomed slice.}
@@
 Classification boundaries are explored for the wine data set [@asuncion:2007].
-It is difficult to digest the boundaries fully -- for example, where one group's boundary wraps another for example, if they ar linear or nonlinear, or whether the boundary goes through the space or is only carving out a corner of it.
+It is difficult to digest the boundaries fully -- for example, where one group's boundary wraps another for example, if they are linear or nonlinear, or whether the boundary goes through the space or is only carving out a corner of it.
 The sliced projections makes this easier.
@@
 A large slice volume $\epsilon = 0.15$ is used.
-The slicing allows exploring the center of the space.
+The slicing allows exploring the centre of the space.
 With 3D it reveals the spherical boundary of the group (green) hidden by the projection.
-In 5D with the polynomial basis, projection might suggest that the boundary between classes is almost linear, but the slicing shows that it to be nonlinear near the center.
+In 5D with the polynomial basis, projection might suggest that the boundary between classes is almost linear, but the slicing shows that it to be nonlinear near the centre.
 

### 2019-09-23 gifs/mkgif.py
Collecting gifs into html page.

@@ -2,19 +2,15 @@ import os
 
 
 p = "./png/"
-allNames = os.listdir(p)
-allNamesUnique = []
-for n in allNames:                                
-  n = n[0:-8]
-  if not n in allNamesUnique: allNamesUnique.append(n)
+names = os.listdir(p)
 
 import moviepy.editor as mpy
 
-for cn in allNamesUnique:                         
-  cNames = []        
-  for nn in allNames:                                 
-    if cn in nn: cNames.append(nn)
+for cn in names:                         
+  cNames = os.listdir(p+cn)
+#  for nn in os.listdir(p+cn):
+#    cNames.append(nn.split(".")[0])
   cNames = sorted(cNames)
-  cPaths = [p+cName for cName in cNames]
+  cPaths = [p+cn+"/"+cName for cName in cNames]
   clip = mpy.ImageSequenceClip(cPaths, fps=6)
   clip.write_gif(cn+".gif", fps=6)

### 2019-09-23 paper.Rmd
Final fixes.

@@
 Slicing, or sectioning, is a way that the internal distribution of high-dimensional data can be explored.
-@prosection discusses a technique for combining projections, with sections constructed by slicing in the dimensions orthogonal to the projection.
+@prosection discusses a technique for combining projections with sections constructed by slicing in the dimensions orthogonal to the projection.
 It is also possible to think of linked brushing (e.g.
@@
 
-To make a section, we are interested in the orthogonal distance of the data points from an anchored projection plane, particularly using $d=2$, but the approach theoretically works for any $d$.
+To make a section, we are interested in the orthogonal distance of the data points from a projection plane, particularly using $d=2$, but the approach theoretically works for any $d$.
 
@@
 \begin{equation}
-\tilde{v}_i^2 = ||\mathbf{x}_i'||^2.
+\tilde{v}_i^2 = ||\mathbf{x}_i'||^2,
 \end{equation}
@@
 When going beyond 3D, the dimension of the orthogonal space is $>1$, and all sense of direction is lost.
-In general, there are $(p-d)$ dimensions orthogonal to the projection plane, so using distance from the points to the plane is the simplest approach to generating slices.
+In general, there are $(p-d)$ dimensions orthogonal to the projection plane, and using distance from the points to the plane is the simplest approach to generating slices.
 Using the Euclidean distance results in rotation invariant slicing in the orthogonal space, where a "slice" is spherical in the orthogonal subspace, and has radius $h$.
@@
 This (centre point) is generally a good option for the slice tour, because as the tour progresses the projection plane changes and can catch non-central concavities, too.
-However it is straightforward generalize the equations to use any centre point, $\mathbf{c}$.
+However it is straightforward to generalize the equations to use any centre point, $\mathbf{c}$.
 
@@
 
-Using the generalized distance measure with a cutoff volume, $\epsilon$, then corresponds to moving a slice of fixed thickness, corresponding to a neighbourhood of the projection plane through the anchor point $\mathbf{c}$.
+Using the generalized distance measure with a cutoff volume, $\epsilon$, then corresponds to moving a slice of fixed thickness, corresponding to a neighbourhood of the projection plane through the centre point $\mathbf{c}$.
 In 3D, this simply corresponds to moving up or down along the orthogonal direction.
-Note that selecting an off-centre anchor point may result in fewer points inside a slice for most data.
+Note that moving $\mathbf{c}$ off-centre will result in fewer points inside a slice for most data.
 
@@
 \caption{Illustrations of slicing, through a 3D sphere (left), and demonstrating the calculation of the orthogonal distance (right).
-3D slicing, can be done by sliding the projection plane along the orthogonal direction: centred at the origin (green) and one off-centre at $\mathbf{c}$ (blue).
+3D slicing can be done by sliding the projection plane along the orthogonal direction: centred at the origin (green) and one off-centre at $\mathbf{c}$ (blue).
 This intuition does not transfer to higher dimensions, and it is best to use orthogonal distance between point and projection plane for computing the slice.
@@
 In addition to usual parameters the user can choose the volume parameter $\epsilon$ by setting `eps` and, if required, the centre point $\mathbf{c}$ is set by the `anchor` argument.
-By default `eps=0.001` and `anchor=NULL`, resulting in slicing through the mean In addition the user can select the marker symbols for point inside and outside the slice.
+By default `eps=0.001` and `anchor=NULL`, resulting in slicing through the mean.
+In addition the user can select the marker symbols for point inside and outside the slice.
 By default points in the slice are highlighted as `pch_slice = 20` and `pch_other = 46`, i.e.
@@
 Slicing through the origin results in sections where the points inside the slice are approximately on a circle with the full radius, and each view is similar to that shown in Figure \ref{fig:renderFunction}.
-Animated gifs of these examples are available at XXX.
 
-It is especially instructive to look at the sliced spheres with an off-centre anchor point.
+It is especially instructive to look at spheres that are sliced off-centre.
 In this case the different views obtained in the slice tour reveal more of the concave structure of the distribution.
-The first row of Figure \ref{fig:3anchored} shows example views from a slice tour on a 3D sphere with $R=1$ and anchor point $(0.7, 0.7, 0.7)$.
-Notice that the anchor point was chosen to fall outside the sphere.
-Depending on the viewing angle selected by the tour the slice contains points on the circle with radius $R$ (when the anchor point has a negligible orthogonal component to the viewing plane); a circle with radius $<R$ as the viewing angle is tilted away from the axis connecting the anchor point to the origin; a full circle with small radius as the angle increases; and finally we see an empty slice when the projection plane is orthogonal to the anchor point axis.
+The first row of Figure \ref{fig:3anchored} shows example views from a slice tour on a 3D sphere with $R=1$ and centre point $(0.7, 0.7, 0.7)$.
+Notice that this was chosen to fall outside the sphere.
+Depending on the viewing angle selected by the tour the slice contains points on the circle with radius $R$ (when the centre point has a negligible orthogonal component to the viewing plane); a circle with radius $<R$ as the viewing angle is tilted away from the axis connecting the centre point to the origin; a full circle with small radius as the angle increases; and finally we see an empty slice when the projection plane is orthogonal to the centre point axis.
 A similar picture is found for the 5D sphere, see second row of Figure \ref{fig:3anchored}.
@@
 We use the pollen data as an example for a hidden feature generally occluded in projections.
-This is a classic 5D data set, originally simulated by by David Coleman of RCA Labs, for the Joint Statistics Meetings 1986 Data Expo [@pollen].
+This is a classic 5D data set, originally simulated by David Coleman of RCA Labs, for the Joint Statistics Meetings 1986 Data Expo [@pollen].
 The standardised data is observed in a slice tour with a thin slice ($\epsilon = 0.0005$).
@@
 Classification boundaries are explored for the wine data set [@asuncion:2007].
-It is difficult to digest the boundaries fully -- for example, where one group's boundary wraps another for example, if they are linear or nonlinear, or whether the boundary goes through the space or is only carving out a corner of it.
+It is difficult to digest the boundaries fully -- for example, where one group's boundary wraps another, if they are linear or nonlinear, or whether the boundary goes through the space or is only carving out a corner of it.
 The sliced projections makes this easier.
@@
 Selected views of projections and sliced projections from a radial basis SVM on 3 variables, and a polynomial basis SVM on 5 variables are shown in Figure \ref{fig:wine}.
-A large slice volume $\epsilon = 0.15$ is used.
 The slicing allows exploring the centre of the space.
@@
 \caption{Exploring classification boundaries of the wine dataset using projection (top row) and sliced projection (bottom row).
-Radial basis SVM (First column) of 3 variables shows how the slicing reveals the spherical shape of the boundary of one group (green) that was hidden in the projection.
+Radial basis SVM (first column) of 3 variables shows how the slicing reveals the spherical shape of the boundary of one group (green) that was hidden in the projection.
 Polynomial basis SVM (second and third columns) of 5 variables.
@@
 
-Sliced 2D tours was available in XGobi [@SCB98] but was not documented. This work makes them available in the `tourr` package. It is a simple, but effective, approach to taking slices.
+Sliced 2D tours were available in XGobi [@SCB98] but were not documented. This work makes them available in the `tourr` package. It is a simple, but effective, approach to taking slices.
 
@@
 
-The paper was written in `rmarkdown` [@rmarkdown] using `knitr` [@knitr]. The source material and animated gifs for this paper are available at
-\url{https://github.com/uschiLaa/paper-slice-tour}.
+The paper was written in `rmarkdown` [@rmarkdown] using `knitr` [@knitr].
+The source material for this paper are available at \url{https://github.com/uschiLaa/paper-slice-tour} and animated gifs can be found at \url{https://uschilaa.github.io/paper-slice-tour/}.

### 2019-09-23 relative_vol_area.R
Removing old file.

@@ -1,64 +1,0 @@
-
-
-relVolCube <- function(p){
-  Vectorize(
-    function(x){
-      (x)^(p-2)
-    }
-  )
-}
-
-erf <- function(x) 2 * pnorm(x * sqrt(2)) - 1
-erfinv <- function (x) qnorm((1 + x)/2)/sqrt(2)
-
-relVolNormal <- function(p){
-  Vectorize(
-    function(x){
-      pi^((2-p)/2) * (erf(x/sqrt(2)))^(p-2)
-    }
-  )
-}
-
-ggplot(data.frame(x=c(0.01, 0.2)), aes(x=x)) +
-  stat_function(fun=relVolCube(3), geom="line", color="black") +
-  stat_function(fun=relVolCube(4), geom="line", color="blue") +
-  stat_function(fun=relVolCube(5), geom="line", color="green") +
-  stat_function(fun=relVolCube(6), geom="line", color="red") +
-  stat_function(fun=relVolNormal(3), geom="line", color="black", linetype="dashed") +
-  stat_function(fun=relVolNormal(4), geom="line", color="blue", linetype="dashed") +
-  stat_function(fun=relVolNormal(5), geom="line", color="green", linetype="dashed") +
-  stat_function(fun=relVolNormal(6), geom="line", color="red", linetype="dashed") +
-  xlab("h/r") +
-  ylab("relative volume/area") +
-  scale_y_log10()
-
-
-hrelForVrel <- function(vrel){
-  Vectorize(
-    function(p){
-      vrel^(1/(p-2))
-    }
-  )
-}
-
-hrelForVolNormal <- function(vrel){
-  Vectorize(
-    function(p){
-      sqrt(2) * erfinv(vrel^(1/(p-2)) * sqrt(pi))
-    }
-  )
-}
-
-ggplot(data.frame(x=c(3, 50)), aes(x=x)) +
-  stat_function(fun=hrelForVrel(0.001), geom="line", color="black") +
-  stat_function(fun=hrelForVrel(0.01), geom="line", color="blue") +
-  stat_function(fun=hrelForVrel(0.05), geom="line", color="green") +
-  stat_function(fun=hrelForVrel(0.1), geom="line", color="red") +
-  stat_function(fun=hrelForVolNormal(0.001), geom="line", color="black", linetype="dashed") +
-  stat_function(fun=hrelForVolNormal(0.01), geom="line", color="blue", linetype="dashed") +
[…]

### 2019-09-23 coverletter.Rmd
cover letter changed

@@
 
-We believe that it is an ideal fit for a short technical note in the Journal of Computational and Graphical Statistics.
-It describes a new graphical method that will be useful to people working in multivariate statistics.
+We believe that it is a good fit as a short technical note in the Journal of Computational and Graphical Statistics.
+Tew graphical method that will be useful to people working in multivariate analysis.
 

### 2019-09-23 paper.Rmd
New default for epsilon reflected in paper.

@@
 To keep the relative number of points (i.e.
-$V_{rel}$) approximately constant for slices in different dimensions $p$, we calculate $h$ from a volume parameter as $h = \epsilon^{1/(p-2)}$, where $\epsilon$ is a pre-chosen value indicating a fraction of the overall volume to slice, say 0.001.
+$V_{rel}$) approximately constant for slices in different dimensions $p$, we calculate $h$ from a volume parameter as $h = \epsilon^{1/(p-2)}$, where $\epsilon$ is a pre-chosen value indicating a fraction of the overall volume to slice, say 0.1.
 
@@
 In addition to usual parameters the user can choose the volume parameter $\epsilon$ by setting `eps` and, if required, the centre point $\mathbf{c}$ is set by the `anchor` argument.
-By default `eps=0.001` and `anchor=NULL`, resulting in slicing through the mean.
+By default `eps=0.1` and `anchor=NULL`, resulting in slicing through the mean.
 In addition the user can select the marker symbols for point inside and outside the slice.

### 2019-09-23 appendix.Rmd
Appendix format changed.

@@
 title: "Appendix"
-type: Short Technical Note
-author:
-  - name: Ursula Laa
-    affil: a, b
-    email: ursula.laa@monash.edu
-  - name: Dianne Cook
-    affil: b
-    email: dicook@monash.edu
-  - name: German Valencia
-    affil: a
-    email: german.valencia@monash.edu
-affiliation:
-  - num: a
-    address: |
-      School of Physics and Astronomy, Monash University
-  - num: b
-    address: |
-      Department of Econometrics and Business Statistics, Monash University
-bibliography: biblio.bib
-geometry: margin=2.5cm
-abstract: |
-  Taking projections of high-dimensional data is a common analytical and visualisation technique in statistics for working with high-dimensional problems.
-Sectioning, or slicing, through high dimensions is less common, but can be useful for visualising data with concavities, or non-linear structure.
-It is associated with conditional distributions in statistics, and also linked brushing between plots in interactive data visualisation.
-This short technical note describes a simple approach for  slicing in the orthogonal
-  space of projections obtained when running a tour, thus presenting the viewer
-  with an interpolated sequence of sliced projections. The method has been
-  implemented in R as an extension to the tourr package, and can be used to
-  explore for concave and non-linear structures in multivariate distributions.
-keywords: |
-  data visualisation; grand tour; sectioning
-header-includes: |
-  \usepackage{hyperref}
-  \usepackage[utf8]{inputenc}
-  \def\tightlist{}
-  \usepackage{setspace}
-output: rticles::tf_article
+output: pdf_document
 ---
@@
 
-\doublespacing
 # Slice volume derivation

### 2019-09-23 render-examples.R
Cleaning up render examples, adding gifski for all gifs.

@@ -101,89 +101,11 @@ render(wine_radial_scaled,
 
 # Render to animated gif
 library(gifski)
-fls <- list.files("gifs/png/wine-radial/")
-gifski(fls , "wine-radial.gif")
+all_dir <- list.dirs(path = "gifs/png", recursive = FALSE)
+for (d in all_dir){
+  name <- paste0("gifs/", strsplit(d, "/")[[1]][3], ".gif")
+  fls <- list.files(d, full.names = TRUE)
+  gifski(fls , name)
+}
 
-
-
-####### end rendered examples
-
-
-#testing out different anchor in 5d with large sample
-sphere5 <- geozoo::sphere.hollow(5, 50000)$points
-colnames(sphere5) <- c("x1", "x2", "x3", "x4", "x5")
-anchor5 <- c(1.1, 0, 0, 0, 0)
-animate(sphere5, grand_tour(), display_slice(anchor = anchor5), rescale = FALSE)
-
-### playing with other exampels
-
-ex1 <- geozoo::boy.surface()$points
-colnames(ex1) <- c("x1", "x2", "x3")
-animate_slice(ex1)
-
-ex2 <- geozoo::cross.cap()$points
-colnames(ex2) <- c("x1", "x2", "x3")
-animate_slice(ex2)
-
-# nice example, only sometimes do we see "corners and edges"
-ex3 <- geozoo::cube.face(6)$points %>%
-  center()
-colnames(ex3) <- c("x1", "x2", "x3", "x4", "x5", "x6")
-animate_slice(ex3)
-
-ex4 <- geozoo::dini.surface()$points %>%
-  scale()
-colnames(ex4) <- c("x1", "x2", "x3")
-animate_slice(ex4)
-
-ex5 <- geozoo::enneper.surface()$points %>%
-  scale()
-colnames(ex5) <- c("x1", "x2", "x3")
-animate_slice(ex5)
-
-# not so nice as example
-ex6 <- geozoo::klein.fig.eight()$points %>%
-  scale()
-colnames(ex6) <- c("x1", "x2", "x3", "x4")
-animate_slice(ex6)
-
-ex7 <- geozoo::mobius()$points %>%
-  scale()
-colnames(ex7) <- c("x1", "x2", "x3")
-animate_slice(ex7)
-
-ex8 <- geozoo::roman.surface()$points %>%
-  scale()
-colnames(ex8) <- c("x1", "x2", "x3")
-animate_slice(ex8)
-
[…]

### 2019-09-23 paper.Rmd
Reference appendix in paper.
Fixing acknowledgements.

@@
 
+The authors gratefully acknowledge the support of the Australian Research Council.
 The paper was written in `rmarkdown` [@rmarkdown] using `knitr` [@knitr].
-The source material for this paper are available at \url{https://github.com/uschiLaa/paper-slice-tour} and animated gifs can be found at \url{https://uschilaa.github.io/paper-slice-tour/}.
+The source material and animated gifs for this paper are available at \url{https://github.com/uschiLaa/paper-slice-tour}.
+An appendix deriving the relative slice volume for the hypersphere is included as supplemental material.

### 2020-04-16 paper.Rmd
Keeping tex file for submission.
Preparing for resubmission.

@@
 output: rticles::tf_article
+keep_tex: true
 ---
@@
 A major limitation of projections is their opacity.
-It is mitigated some, when scatterplots are used to render the projected data, through a pseudo-transparency of sparseness of points.
+It is mitigated somewhat when scatterplots are used to render the projected data, through a pseudo-transparency of sparseness of points.
 This is improved further if points are also drawn using an alpha level that provides transparent dots on most display devices.
 Some features of multivariate distributions may be visible, but a lot is easily hidden, especially in the case of concave structures.
-Considering the example of a simple geometric shape such as a hypershere, it is difficult to distinguish between a full or a hollow sphere, that is, whether points are uniformly distributed within the sphere or on the surface of a sphere.
-Similarly we can think of small scale structures hidden in the centre of a multivariate distribution, that might be considered to be "needles in a haystack", which can be difficult to detect in projections.
+Consider the example of a simple geometric shape such as a hypersphere.
+It is difficult to distinguish between a full or a hollow sphere, that is, whether points are uniformly distributed within the sphere or on the surface of a sphere.\footnote{\textcolor{red}{In standard terms of geometry we are comparing a disk (or ball), $D^p = \{ x \in \mathbb{R}^p : |x| \leq 1\}$ with a standard sphere, $S^{p-1} = \{ x \in \mathbb{R}^p : |x| = 1\}$.}} Similarly we can think of small scale structures hidden in the centre of a multivariate distribution, that might be considered to be "needles in a haystack", which can be difficult to detect in projections.
 Projections also obscure non-linear boundaries as might be constructed from a classification model, or nonlinear model fits in high-dimensions.
@@
 Slicing, or sectioning, is a way that the internal distribution of high-dimensional data can be explored.
-@prosection discusses a technique for combining projections with sections constructed by slicing in the dimensions orthogonal to the projection.
+\textcolor{red}{Here we use "slice" and "section" interchangably, to refer to a portion of the full parameter space spanned by the data distribution.
+The section is defined by limiting the parameter ranges, thus selecting only a sub-space, and in general will be determined by a set of inequality conditions.
+Using a linear condition for going from 3D to 2D we can think of flat slices through the space.} @prosection discusses a technique for combining projections with sections constructed by slicing in the dimensions orthogonal to the projection.
 It is also possible to think of linked brushing (e.g.
@@
 It is constructed by combining a method for basis selection with geodesic interpolation between pairs of bases.
-In a grand tour, the basis selection is random, each new basis is chosen from all possible projections.
+In a grand tour, the basis selection is random; each new basis is chosen from all possible projections.
 In a guided tour, the bases are chosen based on an index of interestingness.
@@
 For the explanation of the slice tour, the actual mechanics of the tour are not important, and only the notion of a projection plane in high dimensions is needed.
-The notation used for explanations starts with denoting each projection as $\mathbf{X}$ the $n\times p$ dimensional data matrix ($n$ observations in $p$ dimensions) and $\mathbf{A}$ an orthonormal $p \times d$ projection matrix.
+The data consists of $n$ observations in $p$-dimensions ($n$ points in $\mathbb{R}^p$) and is thus represented as an $n\times p$- dimensional data matrix $\mathbf{X}$.
+We denote by $\mathbf{A}$, a $p \times d$-dimensional (projection) matrix, the columns of which span a $d$-dimensional subspace of $\mathbb{R}^p$.
 The $d$-dimensional projection of the data is thus given by $\mathbf{Y} = \mathbf{X}\cdot\mathbf{A}$, producing the $n\times d$ dimensional projected data matrix to be plotted in each frame of the tour display.
@@
 The orthogonal distance of the data points from the current projection plane is calculated as, 
-
 \begin{equation}
@@
 \end{equation}
-
 \noindent where 
-
 \begin{equation}
@@
 \end{equation}
-
 \noindent and $\mathbf{x}_i, i=1,...,n$ is a $p$-dimensional observation in $\mathbf{X}$ and $\mathbf{a}_k, k=1,2 (=d)$ denoting the columns of the projection matrix, $\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)$.
-$\mathbf{A}$ is assumed to be through 0, because the data has been centred on its mean.
-$\mathbf{x}_i'$ can be considered to be a normal from the projection plane to $\mathbf{x}_i$, and then the norm of this vector gives the orthogonal distance between point and plane.
+\textcolor{red}{The plane defined by $\mathbf{A}$ passes through 0, suggesting that the data should be centred on its mean.
+Alternatively, we can use a rigid translation to a parallel affine plane that passes through the data mean, as discussed in Section \ref{sec:offcentre}.} $\mathbf{x}_i'$ can be considered to be a normal from the projection plane to $\mathbf{x}_i$, and then the norm of this vector gives the orthogonal distance between point and plane.
 This can be generalized for $d>2$.
 
-The distance can then be used to display a slice tour by highlighting points for which the orthogonal distance is smaller than a selected cutoff value $h$.
+The distance can then be used to display a slice tour by highlighting \textcolor{red}{the projected} points for which the orthogonal distance is smaller than a selected cutoff value $h$.
 In 3D, where for each plane there is only a single orthogonal direction, this defines a flat slice of height $2h$.
-We could use this single direction to slice systematically from one side of the space to the other.
-It is also possible to think of front and back, some points are closer to the viewer, and some are far, and depth cues could be used.
-When going beyond 3D, the dimension of the orthogonal space is $>1$, and all sense of direction is lost.
-In general, there are $(p-d)$ dimensions orthogonal to the projection plane, and using distance from the points to the plane is the simplest approach to generating slices.
+We could use this single direction to slice systematically from one side of the space to the other, as illustrated in Figure \ref{fig:diagrams}.
+It is also possible to think of front and back; some points are closer to the viewer, and some are far, and depth cues could be used.
+\textcolor{red}{When the orthogonal dimension $p-d \geq 2$, there is no longer a unique orthogonal direction in which to translate the projection plane.
+In general,} using distance from the points to the plane is the simplest approach to generating slices.
 Using the Euclidean distance results in rotation invariant slicing in the orthogonal space, where a "slice" is spherical in the orthogonal subspace, and has radius $h$.
@@
 \begin{equation}
-V_{rel} = \frac{1}{2} \frac{h^{p-2}}{R^p} (p R^2 - (p-2) h^2) \approx \frac{1}{2} (\frac{h}{R})^{p-2},
+V_{rel} = \frac{1}{2} \frac{h^{p-2}}{R^p} \left(p R^2 - (p-2) h^2\right) \approx \frac{1}{2} \left(\frac{h}{R}\right)^{p-2},
 \end{equation}
@@
 
-### Non-central slice
+### Non-central slice {#sec:offcentre}
 
-The equations above assume that the projection plane, and thus the the slice, passes through 0, which will be the mean of centred data.
+The equations above assume that the projection plane, and thus the slice, passes through 0, which will be the mean of centred data.
 This (centre point) is generally a good option for the slice tour, because as the tour progresses the projection plane changes and can catch non-central concavities, too.
@@
 Note that moving $\mathbf{c}$ off-centre will result in fewer points inside a slice for most data.
+
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=0.4\textwidth]{diagrams/centered-slice.pdf}
+\hspace{5mm}
+\includegraphics[width=0.4\textwidth]{diagrams/orthogonal-distance.pdf}}
+\caption{Illustrations of slicing, through a 3D sphere (left), and demonstrating the calculation of the orthogonal distance (right).
+3D slicing can be done by sliding the projection plane along the orthogonal direction: centred at the origin (green) and one off-centre at $\mathbf{c}$ (blue).
+This intuition does not transfer to higher dimensions, and it is best to use orthogonal distance between point and projection plane for computing the slice.
+Because a projection plane has no specific location, for slicing we can prescribe this, as through the data centre, or any other point, $\mathbf{c}$.}
+\label{fig:diagrams}
+\end{figure*}
 
@@
 
-\begin{figure*}[ht]
-\centerline{\includegraphics[width=0.4\textwidth]{diagrams/centered-slice.pdf}
-\hspace{5mm}
-\includegraphics[width=0.4\textwidth]{diagrams/orthogonal-distance.pdf}}
-\caption{Illustrations of slicing, through a 3D sphere (left), and demonstrating the calculation of the orthogonal distance (right).
-3D slicing can be done by sliding the projection plane along the orthogonal direction: centred at the origin (green) and one off-centre at $\mathbf{c}$ (blue).
-This intuition does not transfer to higher dimensions, and it is best to use orthogonal distance between point and projection plane for computing the slice.
-Because a projection plane has no specific location, for slicing we can prescribe this, as through the data centre, or any other point, $\mathbf{c}$.}
-\label{fig:diagrams}
-\end{figure*}
-
 # Implementation {#sec:implementation}
@@
 plotting a bullet for points inside the slice, and a dot for points outside.
-Below we show example code for displaying slices through a hollow 3D sphere.
+Below we show example code for displaying slices through a hollow 3D sphere, similar to the example in Figure \ref{fig:renderFunction}.
 
@@
 Notice that this was chosen to fall outside the sphere.
-Depending on the viewing angle selected by the tour the slice contains points on the circle with radius $R$ (when the centre point has a negligible orthogonal component to the viewing plane); a circle with radius $<R$ as the viewing angle is tilted away from the axis connecting the centre point to the origin; a full circle with small radius as the angle increases; and finally we see an empty slice when the projection plane is orthogonal to the centre point axis.
+Depending on the viewing angle selected by the tour the slice contains points on the circle with radius $R$ when the centre point has a negligible orthogonal component to the viewing plane (shown on the right); a circle with radius $<R$ as the viewing angle is tilted away from the axis connecting the centre point to the origin (shown in the middle); a full circle with small radius as the angle increases (shown on the left); and finally we see an empty slice when the projection plane is orthogonal to the centre point axis.
 A similar picture is found for the 5D sphere, see second row of Figure \ref{fig:3anchored}.
@@
 \includegraphics[width=0.2\textwidth]{gifs/png/sphere-5-anchored/sphere-5-anchored-001.png}}
-\caption{Different slices through a 3D (first row) and 5D (second row) hollow sphere with $R=1$, with shifted centre point, showing full circle with small radius (left), circle with radius $<R$ (middle) and circle with radius $R$ (right).}
+\caption{Different slices through a 3D (first row) and 5D (second row) hollow sphere with $R=1$, with shifted centre point, showing a full circle with small radius (left), a circle with radius $<R$ (middle) and a circle with radius $R$ (right).}
 \label{fig:3anchored}
@@
 The second column shows slices through a 4D torus, revealing different aspects of the shape.
-The last column shows a 6d cube, where the upper plot shows a view along two of the original parameters, allowing to clearly identify the rectangular shape.
+The last column shows a 6D cube, where the upper plot shows a view along two of the original parameters, allowing to clearly identify the rectangular shape.
 The panel below shows that this is not typically the case when looking at a randomly selected slice.
@@
 The slicing allows exploring the centre of the space.
-With 3D it reveals the spherical boundary of the group (green) hidden by the projection.
-In 5D with the polynomial basis, projection might suggest that the boundary between classes is almost linear, but the slicing shows that it to be nonlinear near the centre.
+With 3D it reveals the spherical boundary of the group (green) hidden by the projection, as seen by comparing the top and bottom left plots of Figure \ref{fig:wine}.
+In 5D with the polynomial basis, projection might suggest that the boundary between classes is almost linear (top middle and right plots of Figure \ref{fig:wine}), but the slicing shows it to be nonlinear near the centre (bottom middle and right plots of Figure \ref{fig:wine}).
 
 \begin{figure*}[ht]
+\flushleft{\hspace{5mm} Projection, radial basis \hspace{25mm}  Projection, polynomial basis}
 \centerline{
@@
 \includegraphics[width=0.3\textwidth]{gifs/png/wine-poly-projected/wine-poly-projected-042.png}}
+\flushleft{\hspace{10mm} Slice, radial basis \hspace{30mm}  Slice, polynomial basis}
 \centerline{
@@
 The examples shown in Section \ref{sec:examples} demonstrate the potential of this new display to find and explore concave structures, as well as other hidden features.
+\textcolor{red}{Slice tour animations for all examples are available at \url{https://github.com/uschiLaa/paper-slice-tour}.}
 
@@
 The approach by @prosection is more complex, and slices a subspace  of the $p-d=p-2$ dimensional space orthogonal to the projection.
-This generates a more parameters, making it more difficult to navigate.
+This generates more parameters, making it more difficult to navigate.
 More parameters mean more decisions on what to show.
@@
 An appendix deriving the relative slice volume for the hypersphere is included as supplemental material.
+\textcolor{red}{We thank Andreas Buja for his contributions to the initial stage of this project.}

### 2020-04-16 render-examples.R

@@ -12,17 +12,17 @@ colnames(sphere5) <- c("x1", "x2", "x3", "x4", "x5")
 # Animate with the slice display using the default parameters
 max_f <- 100 # max frames for rendering the animation
 set.seed(2019)
-render(sphere3, grand_tour(), display_slice(), 'png', "gifs/png/sphere-3-centered-%03d.png", frames = max_f, rescale = FALSE)
+render(sphere3, grand_tour(), display_slice(axes="off"), 'png', "gifs/png/sphere-3-centered-%03d.png", frames = max_f, rescale = FALSE)
 
 set.seed(2019)
-render(sphere5, grand_tour(), display_slice(), 'png', "gifs/png/sphere-5-centered-%03d.png", frames = max_f, rescale = FALSE)
+render(sphere5, grand_tour(), display_slice(axes="off"), 'png', "gifs/png/sphere-5-centered-%03d.png", frames = max_f, rescale = FALSE)
 
 
 # Animate with off-center anchoring
 anchor3 <- rep(0.7, 3)
 
 set.seed(2019)
-render(sphere3, grand_tour(), display_slice(anchor = anchor3), 'png', "gifs/png/sphere-3-anchored-%03d.png", frames = max_f, rescale = FALSE)
+render(sphere3, grand_tour(), display_slice(anchor = anchor3, axes="off"), 'png', "gifs/png/sphere-3-anchored-%03d.png", frames = max_f, rescale = FALSE)
 
 # render 5d sphere with large number of points and off-center anchor
 sphere5 <- geozoo::sphere.hollow(5, 20000)$points
@@ -30,30 +30,30 @@ colnames(sphere5) <- c("x1", "x2", "x3", "x4", "x5")
 anchor5 <- c(1.2, rep(0, 4))
 max_f <- 100 # max frames for rendering the animation
 set.seed(2019)
-render(sphere5, grand_tour(), display_slice(anchor = anchor5), 'png', "gifs/png/sphere-5-anchored-%03d.png", frames = max_f, rescale = FALSE)
+render(sphere5, grand_tour(), display_slice(anchor = anchor5, axes="off", eps=0.01), 'png', "gifs/png/sphere-5-anchored-%03d.png", frames = max_f, rescale = FALSE)
 
 
 torus4 <- geozoo::torus(p=4)$points %>%
   scale()
 colnames(torus4) <- c("x1", "x2", "x3", "x4")
[…]

### 2020-04-17 paper.Rmd
Working on revision, renaming epsilon.

@@
 Consider the example of a simple geometric shape such as a hypersphere.
-It is difficult to distinguish between a full or a hollow sphere, that is, whether points are uniformly distributed within the sphere or on the surface of a sphere.\footnote{\textcolor{red}{In standard terms of geometry we are comparing a disk (or ball), $D^p = \{ x \in \mathbb{R}^p : |x| \leq 1\}$ with a standard sphere, $S^{p-1} = \{ x \in \mathbb{R}^p : |x| = 1\}$.}} Similarly we can think of small scale structures hidden in the centre of a multivariate distribution, that might be considered to be "needles in a haystack", which can be difficult to detect in projections.
+It is difficult to distinguish between a full or a hollow sphere, that is, whether points are uniformly distributed within the sphere or on the surface of a sphere.
+\textcolor{red}{In standard terms of geometry we are comparing a disk (or ball), $D^p = \{ x \in \mathbb{R}^p : |x| \leq 1\}$ with a standard sphere, $S^{p-1} = \{ x \in \mathbb{R}^p : |x| = 1\}$.} Similarly we can think of small scale structures hidden in the centre of a multivariate distribution, that might be considered to be "needles in a haystack", which can be difficult to detect in projections.
 Projections also obscure non-linear boundaries as might be constructed from a classification model, or nonlinear model fits in high-dimensions.
@@
 Slicing, or sectioning, is a way that the internal distribution of high-dimensional data can be explored.
-\textcolor{red}{Here we use "slice" and "section" interchangably, to refer to a portion of the full parameter space spanned by the data distribution.
-The section is defined by limiting the parameter ranges, thus selecting only a sub-space, and in general will be determined by a set of inequality conditions.
-Using a linear condition for going from 3D to 2D we can think of flat slices through the space.} @prosection discusses a technique for combining projections with sections constructed by slicing in the dimensions orthogonal to the projection.
+\textcolor{red}{Mathematically, a section of a data distribution is obtained as the intersection of the observations with a thickened version of an affine subspace of the data space [@prosections].
+This subspace can be defined as a set of inequality conditions which in general can depend on any function of the input variables.} @prosection discusses a technique for combining projections with sections constructed by slicing in the dimensions orthogonal to the projection.
 It is also possible to think of linked brushing (e.g.
@@
 \noindent and $\mathbf{x}_i, i=1,...,n$ is a $p$-dimensional observation in $\mathbf{X}$ and $\mathbf{a}_k, k=1,2 (=d)$ denoting the columns of the projection matrix, $\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)$.
-\textcolor{red}{The plane defined by $\mathbf{A}$ passes through 0, suggesting that the data should be centred on its mean.
+\textcolor{red}{A priori, the projection matrix $\mathbf{A}$ defines a plane that is centered at the the origin.
+Generally we want to centre the plane at the data mean, and this is achieved by centring the data distribution on its mean.
 Alternatively, we can use a rigid translation to a parallel affine plane that passes through the data mean, as discussed in Section \ref{sec:offcentre}.} $\mathbf{x}_i'$ can be considered to be a normal from the projection plane to $\mathbf{x}_i$, and then the norm of this vector gives the orthogonal distance between point and plane.
@@
 The distance can then be used to display a slice tour by highlighting \textcolor{red}{the projected} points for which the orthogonal distance is smaller than a selected cutoff value $h$.
-In 3D, where for each plane there is only a single orthogonal direction, this defines a flat slice of height $2h$.
+\textcolor{red}{While all projected points may be drawn in the display, point size and/or color can be used to highlight the points inside the slice.
+This is illustrated in Figure \ref{fig:renderFunction}.} In 3D, where for each plane there is only a single orthogonal direction, this defines a flat slice of height $2h$.
 We could use this single direction to slice systematically from one side of the space to the other, as illustrated in Figure \ref{fig:diagrams}.
 It is also possible to think of front and back; some points are closer to the viewer, and some are far, and depth cues could be used.
-\textcolor{red}{When the orthogonal dimension $p-d \geq 2$, there is no longer a unique orthogonal direction in which to translate the projection plane.
-In general,} using distance from the points to the plane is the simplest approach to generating slices.
+\textcolor{red}{However, only $p-d =1$ results in a unique orthogonal direction along which the projection plane can be translated, and in general using distance} from the points to the plane is the simplest approach to generating slices.
 Using the Euclidean distance results in rotation invariant slicing in the orthogonal space, where a "slice" is spherical in the orthogonal subspace, and has radius $h$.
@@
 To keep the relative number of points (i.e.
-$V_{rel}$) approximately constant for slices in different dimensions $p$, we calculate $h$ from a volume parameter as $h = \epsilon^{1/(p-2)}$, where $\epsilon$ is a pre-chosen value indicating a fraction of the overall volume to slice, say 0.1.
+$V_{rel}$) approximately constant for slices in different dimensions $p$, we calculate $h$ from a volume parameter as $h = v_{rel}^{1/(p-2)}$, where $v_{rel}$ is a pre-chosen value indicating a fraction of the overall volume to slice, say 0.1 \textcolor{red}{(and an approximation to the relative volume of the slice in a hypersphere)}.
 
@@
 
-Using the generalized distance measure with a cutoff volume, $\epsilon$, then corresponds to moving a slice of fixed thickness, corresponding to a neighbourhood of the projection plane through the centre point $\mathbf{c}$.
+Using the generalized distance measure with a cutoff volume, $v_{rel}$, then corresponds to moving a slice of fixed thickness, corresponding to a neighbourhood of the projection plane through the centre point $\mathbf{c}$.
 In 3D, this simply corresponds to moving up or down along the orthogonal direction.
@@
 The slice tour has been implemented in R [@rref] as a new display method `display_slice` in the `tourr` package [@tourr].
-In addition to usual parameters the user can choose the volume parameter $\epsilon$ by setting `eps` and, if required, the centre point $\mathbf{c}$ is set by the `anchor` argument.
-By default `eps=0.1` and `anchor=NULL`, resulting in slicing through the mean.
+In addition to usual parameters the user can choose the volume parameter $v_{rel}$ by setting `v_rel` and, if required, the centre point $\mathbf{c}$ is set by the `anchor` argument.
+By default `v_rel=0.1` and `anchor=NULL`, resulting in slicing through the mean.
 In addition the user can select the marker symbols for point inside and outside the slice.
@@
 anchor3 <- rep(0.7, 3)
-animate_slice(sphere3, anchor = anchor3, eps = 0.2, pch_other=26)
+animate_slice(sphere3, anchor = anchor3, v_rel = 0.2, pch_other=26)
 ```
@@
 Using the `geozoo` package [@geozoo] a number of ideal shapes are generated.
+\textcolor{red}{Here we show selected slices, slice tour animations for all examples are available at \url{https://github.com/uschiLaa/paper-slice-tour}.}
 
@@
 Using the `sphere.hollow()` function, we generate two sample data sets by generating 2000 and 5000 points from a uniform distribution on the surface of a 3D and 5D spheres.
-For both examples we start by slicing through the origin with the default parameters, in particular $\epsilon = 0.1$, i.e.
+For both examples we start by slicing through the origin with the default parameters, in particular $v_{rel} = 0.1$, i.e.
 $h = 0.1$ in 3D and $h = 0.46$ in 5D.
@@
 This is a classic 5D data set, originally simulated by David Coleman of RCA Labs, for the Joint Statistics Meetings 1986 Data Expo [@pollen].
-The standardised data is observed in a slice tour with a thin slice ($\epsilon = 0.0005$).
+The standardised data is observed in a slice tour with a thin slice ($v_{rel} = 0.0005$).
 Selected views are shown in the first two plots in Figure \ref{fig:pollen}.
@@
 \caption{Different slices through the classic pollen dataset.
-The first two plots have slice volume $\epsilon = 0.0005$, the last two plots zoom in on the centre and have increased volume $\epsilon = 0.005$.
+The first two plots have slice volume $v_{rel} = 0.0005$, the last two plots zoom in on the centre and have increased volume $v_{rel} = 0.005$.
 The hidden word can be seen in the zoomed slice.}
@@
 An appendix deriving the relative slice volume for the hypersphere is included as supplemental material.
-\textcolor{red}{We thank Andreas Buja for his contributions to the initial stage of this project.}
+\textcolor{red}{We thank Andreas Buja for his contributions during the initial stage of this project.}

### 2020-04-17 paper.Rmd
done with proffing the changes

@@
 keywords: |
-  data visualisation; grand tour; sectioning; statistical computing; statistical graphics; high-dimensional data
+  data visualisation; grand tour; statistical computing; statistical graphics; multivariate data; dynamic graphics
 header-includes: |
@@
 -->
-Slicing, or sectioning, is a way that the internal distribution of high-dimensional data can be explored.
-\textcolor{red}{Mathematically, a section of a data distribution is obtained as the intersection of the observations with a thickened version of an affine subspace of the data space [@prosections].
+\textcolor{red}{Slicing, or sectioning, is a way to explore for these types of structures.
+Mathematically, a section of a data distribution is obtained as the intersection of the observations with a thickened version of an affine subspace of the data space} [@prosection] \textcolor{red}{.
 This subspace can be defined as a set of inequality conditions which in general can depend on any function of the input variables.} @prosection discusses a technique for combining projections with sections constructed by slicing in the dimensions orthogonal to the projection.
@@
 \noindent and $\mathbf{x}_i, i=1,...,n$ is a $p$-dimensional observation in $\mathbf{X}$ and $\mathbf{a}_k, k=1,2 (=d)$ denoting the columns of the projection matrix, $\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)$.
-\textcolor{red}{A priori, the projection matrix $\mathbf{A}$ defines a plane that is centered at the the origin.
-Generally we want to centre the plane at the data mean, and this is achieved by centring the data distribution on its mean.
-Alternatively, we can use a rigid translation to a parallel affine plane that passes through the data mean, as discussed in Section \ref{sec:offcentre}.} $\mathbf{x}_i'$ can be considered to be a normal from the projection plane to $\mathbf{x}_i$, and then the norm of this vector gives the orthogonal distance between point and plane.
+\textcolor{red}{A priori, the projection matrix $\mathbf{A}$ defines a plane that is centered at the origin.
+Generally we want to centre the plane at the data mean, and this is achieved by centering the data distribution at the origin by subtracting its mean.
+Alternatively, we can use a rigid translation to a parallel affine plane that passes through the data mean, as discussed in Section \ref{sec:offcentre}.} The quantity $\mathbf{x}_i'$ can be considered to be a normal from the projection plane to $\mathbf{x}_i$, and then the norm of this vector gives the orthogonal distance between point and plane.
 This can be generalized for $d>2$.
@@
 Using the `geozoo` package [@geozoo] a number of ideal shapes are generated.
-\textcolor{red}{Here we show selected slices, slice tour animations for all examples are available at \url{https://github.com/uschiLaa/paper-slice-tour}.}
+\textcolor{red}{Selected slices views are shown in the figures.
+The slice tour animations for all examples are available at \url{https://github.com/uschiLaa/paper-slice-tour}.}
 
@@
 The examples shown in Section \ref{sec:examples} demonstrate the potential of this new display to find and explore concave structures, as well as other hidden features.
-\textcolor{red}{Slice tour animations for all examples are available at \url{https://github.com/uschiLaa/paper-slice-tour}.}
 
@@
 
-# Acknowledgements
+# Acknowledgements {-}
 
@@
 The paper was written in `rmarkdown` [@rmarkdown] using `knitr` [@knitr].
+\textcolor{red}{We thank Andreas Buja for his contributions during the initial stage of this project.}
+
+# Supplementary material {-}
+
 The source material and animated gifs for this paper are available at \url{https://github.com/uschiLaa/paper-slice-tour}.
-An appendix deriving the relative slice volume for the hypersphere is included as supplemental material.
-\textcolor{red}{We thank Andreas Buja for his contributions during the initial stage of this project.}
+An appendix contains the derivation of the relative slice volume for the hypersphere.

### 2020-04-19 paper.Rmd
Australian spelling.

@@
 For the explanation of the slice tour, the actual mechanics of the tour are not important, and only the notion of a projection plane in high dimensions is needed.
-The data consists of $n$ observations in $p$-dimensions ($n$ points in $\mathbb{R}^p$) and is thus represented as an $n\times p$- dimensional data matrix $\mathbf{X}$.
-We denote by $\mathbf{A}$, a $p \times d$-dimensional (projection) matrix, the columns of which span a $d$-dimensional subspace of $\mathbb{R}^p$.
+The data consists of $n$ observations in $p$-dimensions ($n$ points in $\mathbb{R}^p$) and is thus represented as an $n\times p$ dimensional data matrix $\mathbf{X}$.
+We denote by $\mathbf{A}$, a $p \times d$ dimensional (projection) matrix, the columns of which span a $d$-dimensional subspace of $\mathbb{R}^p$.
 The $d$-dimensional projection of the data is thus given by $\mathbf{Y} = \mathbf{X}\cdot\mathbf{A}$, producing the $n\times d$ dimensional projected data matrix to be plotted in each frame of the tour display.
@@
 \noindent and $\mathbf{x}_i, i=1,...,n$ is a $p$-dimensional observation in $\mathbf{X}$ and $\mathbf{a}_k, k=1,2 (=d)$ denoting the columns of the projection matrix, $\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)$.
-\textcolor{red}{A priori, the projection matrix $\mathbf{A}$ defines a plane that is centered at the origin.
-Generally we want to centre the plane at the data mean, and this is achieved by centering the data distribution at the origin by subtracting its mean.
+\textcolor{red}{A priori, the projection matrix $\mathbf{A}$ defines a plane that is centred at the origin.
+Generally we want to centre the plane at the data mean, and this is achieved by centring the data distribution at the origin by subtracting its mean.
 Alternatively, we can use a rigid translation to a parallel affine plane that passes through the data mean, as discussed in Section \ref{sec:offcentre}.} The quantity $\mathbf{x}_i'$ can be considered to be a normal from the projection plane to $\mathbf{x}_i$, and then the norm of this vector gives the orthogonal distance between point and plane.
@@
 The distance can then be used to display a slice tour by highlighting \textcolor{red}{the projected} points for which the orthogonal distance is smaller than a selected cutoff value $h$.
-\textcolor{red}{While all projected points may be drawn in the display, point size and/or color can be used to highlight the points inside the slice.
+\textcolor{red}{While all projected points may be drawn in the display, point size and/or colour can be used to highlight the points inside the slice.
 This is illustrated in Figure \ref{fig:renderFunction}.} In 3D, where for each plane there is only a single orthogonal direction, this defines a flat slice of height $2h$.
